% !TEX root = ../main.tex
\chapter{Introduction}
\label{ch:intro}

Estimating tail risk is challlenging because the events of greatest practical
interest are precisely those for which the data are scarce. In financial,
insurance and operational-risk applications, the target may lie beyond the
range where empirical averages or order statistics alone give reliable
answers. Extreme-value methods address this by combining intermediate tail
information with a heavy-tailed model, trading a small amount of structural
assumption for the ability to extrapolate to rarer loss levels. This thesis
works in that regime, with a focus on extreme expectiles as tail-risk
functionals.

Expectiles, introduced by \textcite{NeweyPowell1987}, are asymmetric
least-squares analogues of quantiles. At high levels they respond to both the
size and probability of large losses, and they have attractive risk-measure
properties, including coherence in the relevant high-level range and
elicitability under standard integrability conditions
\parencite{BelliniKlarMullerRosazzaGianin2014,Ziegel2016,Gneiting2011}.
For heavy-tailed distributions with finite mean, extreme expectiles are also
asymptotically proportional to extreme quantiles. In the notation used
throughout the thesis,
\[
  \expectile{\tau}
  \sim
  \psifn(\gamma)Q(\tau),
  \qquad
  \psifn(\gamma)=(\gamma^{-1}-1)^{-\gamma},
  \qquad \tau\uparrow1,
\]
where \(Q\) is the marginal quantile function and \(\gamma\) is the
extreme-value index. This bridge is the point at which expectile inference can
be connected to the more developed theory of extreme-quantile extrapolation.

The second ingredient is pooling. In many statistical settings the available
sample is naturally distributed across several machines, institutions or data
sources. If these sources share the same upper-tail target, each local sample
contains useful information but none is necessarily ideal on its own.
\textcite{DaouiaPadoanStupfler2021} develop an optimal weighted-pooling
framework for tail-index and extreme-quantile inference, including geometric
pooling of local Weissman estimators. Their results identify variance- and
AMSE-optimal weights for the pooled extreme-quantile problem and provide
estimated-weight and plug-in inference machinery under source conditions. The
question in this thesis is how much of that structure survives after the target
is changed from an extreme quantile to an extreme expectile.

The estimator studied here applies the expectile--quantile bridge after the
pooled extreme-quantile extrapolation. For a target level
\(\tau_n=1-p_n\), define
\[
  \widehat\xi_{\tau_n}^{\mathrm{pool},\star}(\bm\nu,\bm\omega)
  =
  \psifn\!\big(\widehat\gamma_n(\bm\nu)\big)\,
  \widehat q_n^\star(\tau_n\mid\bm\omega).
\]
The vector \(\bm\omega\) weights the geometrically pooled Weissman estimator
\(\widehat q_n^\star(\tau_n\mid\bm\omega)\). The vector \(\bm\nu\) weights
only the pooled Hill estimator appearing in the outer bridge factor. This
separation is important. The diagonal choice \(\bm\nu=\bm\omega\) is a
natural prototype, but it is not forced by the construction and is not treated
as an optimality principle.

The main theoretical contribution is to analyse this estimator from an exact
log decomposition. The decomposition separates three terms: the
pooled-Weissman quantile error, the error from plugging a pooled Hill estimator
into \(\psifn\), and the population error in replacing the expectile by its
asymptotic quantile bridge. In the finite-\(m\) iid common-marginal setting,
under the same extreme-Weissman scale and an explicit bridge-rate condition,
the latter two terms are lower order. The resulting first-order limit is
therefore inherited from the pooled-Weissman component. Consequently, the
first-order bias, variance and AMSE criteria depend on \(\bm\omega\), while
\(\bm\nu\) is first-order unidentified. The practical bridge-Hill convention
used later is the deterministic threshold-allocation vector
\(\bm\nu_n^k=(k_1/k,\ldots,k_m/k)^\top\), not an estimated or optimised
bridge-Hill weight.

This first-order separation also determines the inference route. Estimated
weights are introduced only in the pooled-Weissman role: when the
Daouia--Padoan--Stupfler source theory supplies a consistent estimated
\(\widehat{\bm\omega}_n\), together with consistent plug-in centring and
variance objects, these objects transfer through the expectile decomposition.
This yields a plug-in-centred studentised statistic and a multiplicative
log-scale interval for the extreme expectile.

The finite-sample study is deliberately narrow and follows the theory's
scope. It keeps the bridge-Hill vector fixed at
\(\bm\nu_n^k\) and varies only the pooled-Weissman weights. In the simulated
iid common-marginal designs, Daouia--Padoan--Stupfler variance weights are a
stable practical baseline and track the centralised benchmark closely for
point estimation. AMSE weights are almost indistinguishable from variance
weights in the main equal-threshold grid, but the heterogeneous-threshold
diagnostic shows that they can become informative, with some stability costs.
The log-scale intervals under-cover in several finite-sample rows, so they
are reported as diagnostics of the first-order plug-in route rather than as a
claim of strong finite-sample calibration.

The rest of the thesis is organised as follows. \Cref{ch:background} recalls
the heavy-tail, expectile and optimal-pooling results used as source inputs.
\Cref{ch:pooled-extreme} develops the pooled extreme-expectile estimator, its
exact decomposition, the deterministic first-order theorem, the weight
consequences and the estimated-weight inference transfer. \Cref{ch:finite-sample}
reports the finite-sample study for the promoted estimator and inference
route.
