% !TEX root = ../main.tex
\chapter{Background}
\label{ch:background}

\section*{Outline of the chapter}

This chapter collects the technical ingredients on which the thesis
builds: regular variation and the second-order condition; the Hill and
Weissman estimators for the tail index and extreme quantiles;
expectiles and their coherence and elicitability properties as risk
measures; single-sample iid extreme-expectile inference, with the
plug-in central limit theorem that the contribution chapters take as
their workhorse; and the optimal pooling framework of
\textcite{DaouiaPadoanStupfler2021}. Nothing in this chapter is
original; the contributions begin in \Cref{ch:pool-int}.

\section{Regular variation and the second-order condition}
\label{sec:bg:rv}

Throughout the thesis we work with a real-valued random loss $X$ that
has continuous distribution function $F$ and survival function
$\overline F = 1 - F$. Given an iid sample $X_{1}, \ldots, X_{n}$ from
$F$, the corresponding order statistics are denoted by
$X_{1:n} \le X_{2:n} \le \cdots \le X_{n:n}$.

Heavy-tailed asymptotics are governed by the language of regular
variation; the reference is \textcite[Appendix~B.1]{deHaanFerreira2006}.

\begin{definition}[Regular variation]
\label{def:bg:rv}
A Lebesgue measurable function $f : (0,\infty) \to \R$ that is eventually
positive is \emph{regularly varying at infinity with index} $\alpha \in \R$, written
$f \in \mathrm{RV}_{\alpha}$, if for every $y > 0$,
\[
  \lim_{t \to \infty} \frac{f(t y)}{f(t)} \;=\; y^{\alpha}.
\]
The functions in $\mathrm{RV}_{0}$ are called \emph{slowly varying}.
\end{definition}

The first-order condition under which all of our asymptotics are taken
asserts that $F$ has a Pareto-type tail.

\begin{condition}[First-order condition]
\label{cond:bg:first-order}
There exists $\gamma > 0$ such that
$\overline F \in \mathrm{RV}_{-1/\gamma}$, or equivalently the
\emph{tail quantile function}
\[
  U(t) \;\coloneqq\; \inf\!\Big\{ x \in \R \,:\, \tfrac{1}{\overline F(x)} \ge t \Big\}
  \,,\qquad t>1,
\]
satisfies $U \in \mathrm{RV}_{\gamma}$.
\end{condition}

Throughout, for a continuous distribution function $F$, we write
\[
  \Quant_X(\tau)
  \;\coloneqq\;
  F^{\leftarrow}(\tau)
  =
  \inf\{x\in\R:F(x)\ge \tau\},
  \qquad 0<\tau<1.
\]
With the tail quantile $U(t)=F^{\leftarrow}(1-1/t)$, $t>1$, this gives
\[
  \Quant_X(\tau)=U((1-\tau)^{-1}),
  \qquad
  \Quant_X(1-p)=U(1/p).
\]

\Cref{cond:bg:first-order} is equivalent to saying that $X$ lies in the
Fr\'echet max-domain of attraction with extreme-value index
$\gamma$ \parencite[Theorem~1.2.1]{deHaanFerreira2006}. The parameter
$\gamma > 0$ is the \emph{extreme-value index} of $F$; it controls how
slowly $\overline F$ decays at infinity. Standard heavy-tailed
distributions encountered in risk management satisfy
\Cref{cond:bg:first-order}, including the Pareto, Burr, Student-$t$,
Fr\'echet, and $F$-distribution families
\parencite[\S2.6]{deHaanFerreira2006}.

For the asymptotic-normality statements that drive the inferential
theory we need a quantitative refinement of \Cref{cond:bg:first-order}
controlling the rate of convergence in \Cref{def:bg:rv}. The standard
refinement is the second-order condition.

\begin{condition}[Second-order condition $\mathcal C_{2}(\gamma, \rho, A)$]
\label{cond:bg:second-order}
There exist $\gamma > 0$, a second-order parameter $\rho \le 0$, and a
measurable auxiliary function $A$ of constant sign with $A(t) \to 0$ as
$t \to \infty$, such that for every $y > 0$,
\[
  \lim_{t \to \infty} \frac{1}{A(t)}
    \!\left[ \frac{U(t y)}{U(t)} - y^{\gamma}\right]
  \;=\; y^{\gamma}\, \frac{y^{\rho} - 1}{\rho},
\]
where the right-hand side is read as $y^{\gamma} \log y$ when
$\rho = 0$.
\end{condition}

The function $|A|$ is itself regularly varying with index $\rho$
\parencite[Theorem~2.3.9]{deHaanFerreira2006}. Since the auxiliary
function $A$ has constant sign eventually, all signed ratios
$A(t y)/A(t)$ used below have the same limit $y^\rho$. Smaller $|\rho|$
means slower convergence and therefore larger second-order bias of the
tail-index estimators we shall encounter in
\Cref{sec:bg:tail-index-quantile}.

\begin{condition}[Lower-tail moment condition]
\label{cond:bg:lower-tail-moment}
For some $\delta>0$, the negative part
$X^{-}=\max(-X,0)$ satisfies
\[
  \E (X^{-})^{2+\delta}<\infty .
\]
\end{condition}

\textcite{DaouiaGirardStupfler2018} write the same assumption as
$\E|Y_-|^{2+\delta}<\infty$ with $Y_-=\min(Y,0)$; here $X^-$ denotes
the nonnegative negative part.

For the expectile CLTs used below we shall work in the range
$\gamma \in (0, 1/2)$ together with an explicit lower-tail moment
condition. The restriction $\gamma < 1/2$ controls the positive tail
and ensures $\E (X^{+})^{2} < \infty$; since $X$ is real-valued, it does
not by itself control the negative part. \Cref{cond:bg:lower-tail-moment}
is the corresponding left-tail assumption used in the extreme-expectile
limit theorems.

\section{Tail-index and extreme-quantile estimation}
\label{sec:bg:tail-index-quantile}

Under the first-order condition, the extreme-value index $\gamma > 0$
is the primary parameter determining the behaviour of $F$ in the tail.
We collect here the two classical estimators that the thesis builds
on: the Hill estimator of $\gamma$ and the Weissman extrapolation for
extreme quantiles. Both are standard; the textbook reference is
\textcite[Ch.~3--4]{deHaanFerreira2006}.

\subsection{The Hill estimator}
\label{sec:bg:hill}

Given an iid sample $X_{1}, \ldots, X_{n}$ from $F$ and an integer
$k = k_{n} \in \{1, \ldots, n-1\}$, the \emph{Hill estimator}
\parencite{Hill1975} of
$\gamma$ based on the $k + 1$ top order statistics is
\begin{equation}
\label{eq:bg:hill}
  \hat\gamma_{n}(k) \;=\; \frac{1}{k} \sum_{i=1}^{k} \log\!\left(
    \frac{X_{n-i+1\,:\,n}}{X_{n-k\,:\,n}} \right).
\end{equation}
Since $U(t) \to \infty$ under \Cref{cond:bg:first-order}, the event
$X_{n-k:n} > 0$ has probability tending to one for intermediate
$k$. The log-based tail estimators in this section are understood on
that event; equivalently one may assume that the upper tail is
eventually supported on $(0,\infty)$, or, for tail-index estimation
only, apply a deterministic shift, which leaves the extreme-value index
unchanged. For quantile or expectile estimation the shift must be
undone after extrapolation; the large-quantile ratio asymptotics are
unaffected because the upper endpoint is infinite.
The interpretation is direct: if the upper tail of $F$ were exactly
Pareto with index $1/\gamma$, the logarithms of excess ratios above a
fixed tail threshold would be exponential with mean $\gamma$; in
Hill's order-statistic derivation, this appears through the scaled
consecutive log-spacings
$i\{\log X_{n-i+1:n}-\log X_{n-i:n}\}$, which are independent
exponential variables with mean $\gamma$ in the exact Pareto case and
whose average is \eqref{eq:bg:hill}.
\Cref{cond:bg:first-order} asserts that the upper tail is
\emph{approximately} Pareto. The consistency statement below is
\textcite[Theorem~3.2.2]{deHaanFerreira2006}; the normal limit is
\textcite[Theorem~3.2.5]{deHaanFerreira2006} when $\rho<0$, with the
$\rho=0$ endpoint in the same $\mathcal C_{2}$ convention covered by
the $m=1$ specialisation of \textcite[Theorem~1]{DaouiaPadoanStupfler2021}.

\begin{theorem}
\label{thm:bg:hill}
Assume $\mathcal C_{2}(\gamma, \rho, A)$ with $\gamma > 0$. Let
$k = k_{n} \to \infty$ and $k / n \to 0$ as $n \to \infty$. Then
$\hat\gamma_{n}(k) \toprob \gamma$. If, in addition,
$\sqrt{k}\, A(n/k) \to \lambda \in \R$, then
\[
  \sqrt{k}\,\big( \hat\gamma_{n}(k) - \gamma \big)
  \;\todist\;
  \mathcal N\!\left( \frac{\lambda}{1 - \rho},\; \gamma^{2} \right).
\]
\end{theorem}

The choice of $k$ controls the well-known bias--variance trade-off:
too small a $k$ inflates the variance, while too large a $k$ injects
second-order bias through the term $\lambda / (1 - \rho)$. Adaptive
choice of $k$ is a research topic in its own right; we do not pursue
it in this thesis and treat $k$ as fixed by the data analyst.

\subsection{Weissman extrapolation to extreme quantiles}
\label{sec:bg:weissman}

The \emph{Weissman extrapolation} \parencite{Weissman1978} extrapolates
from the intermediate-order statistic $X_{n - k:n}$ --- which
estimates the quantile $\Quant_{X}(1 - k/n)$ at the intermediate level
$1 - k/n$ --- to a quantile at a very-extreme level $1 - p$ with
$p = p_{n} \to 0$ as $n \to \infty$. Plugging the Hill estimator into
the first-order identity $U(t y) / U(t) \to y^{\gamma}$ gives the
Weissman estimator
\begin{equation}
\label{eq:bg:weissman}
  \hatWei_{n}(1 - p \mid k)
  \;=\; \left( \frac{k}{n p} \right)^{\hat\gamma_{n}(k)}
        X_{n - k\,:\,n}.
\end{equation}
For $p=c/n$ and fixed $k$, \eqref{eq:bg:weissman} reduces to the
large-quantile estimator proposed by \textcite[\S4]{Weissman1978}
in the Fr\'echet case, written there with tail parameter
$\alpha=1/\gamma$ and descending order statistics.
The theorem below covers the far-tail regime $np=o(k)$. The especially
severe very-extreme case has $np=O(1)$, where the expected number of
observations beyond $\Quant_X(1-p)$ remains bounded; when $p<1/n$, the
target is beyond the usual sample-maximum scale.
It is the Hill-estimator specialisation of
\textcite[Theorem~4.3.8]{deHaanFerreira2006}; since
$\log k=o(\sqrt{k})$, the source condition
$\log(np_n)=o(\sqrt{k})$ is equivalently written here as
$\log(k/(np_n))=o(\sqrt{k})$.

\begin{theorem}
\label{thm:bg:weissman}
Assume $\mathcal C_{2}(\gamma, \rho, A)$ with $\gamma > 0$ and
$\rho < 0$. Let $k = k_{n} \to \infty$, $k / n \to 0$, $p = p_{n} \to 0$
with $k / (np) \to \infty$ and $\sqrt{k} / \log(k / (np)) \to \infty$.
Suppose $\sqrt{k}\, A(n/k) \to \lambda \in \R$. Then
\[
  \frac{\sqrt{k}}{\log\!\big(k/(np)\big)}
  \left( \frac{\hatWei_{n}(1 - p \mid k)}{\Quant_{X}(1 - p)} - 1 \right)
  \;\todist\;
  \mathcal N\!\left( \frac{\lambda}{1 - \rho},\; \gamma^{2} \right).
\]
\end{theorem}

The rate $\sqrt{k} / \log(k / (np))$ reflects the cost of
extrapolation: the further into the tail one extrapolates, the larger
the log-denominator and the slower the rate. The asymptotic variance
is identical to that of the Hill estimator (\Cref{thm:bg:hill}); the
extrapolation contributes only through the multiplicative log-scale
factor. This is the structural fact that
\textcite[\S2.3]{DaouiaPadoanStupfler2021} exploits to motivate the
weighted-\emph{geometric} pooling of Weissman estimators --- rather
than the arithmetic pooling natural for centred quantities --- and
that we shall in turn exploit in \Cref{ch:pool-ext} for the
extrapolation of pooled expectiles.

\section{Expectiles}
\label{sec:bg:expectiles}

This section collects the definition and the basic properties of
expectiles that we shall use in the rest of the thesis. Expectiles
were introduced by \textcite{NeweyPowell1987} as an
asymmetric-least-squares analogue of quantiles. Their role as risk
measures, in particular their coherence properties, was clarified by
\textcite{BelliniKlarMullerRosazzaGianin2014} and
\textcite{Ziegel2016}; their elicitability and scoring-function
representations are treated in \textcite{Gneiting2011} and
\textcite{EhmGneitingJordanKrueger2016}.

\subsection{Definition}

Let $X$ be a real-valued random variable with $\E |X| < \infty$, and
fix a level $\tau \in (0, 1)$. The $\tau$-\emph{expectile} of $X$ is
defined as
\begin{equation}
\label{eq:bg:expectile-def}
  \expectile{\tau}(X)
  \;\coloneqq\;
  \argmin_{\theta \in \R}\;
  \E\!\left[\, \eta_{\tau}(X - \theta) - \eta_{\tau}(X) \,\right],
  \qquad
  \eta_{\tau}(u) \coloneqq \abs{\tau - \one\{u \le 0\}}\, u^{2}.
\end{equation}
The asymmetric squared check function $\eta_{\tau}$ penalises positive
and negative deviations from $\theta$ with the asymmetric weights
$\tau$ and $1 - \tau$, respectively. Recovering the mean in the
symmetric case is immediate: $\eta_{1/2}(u) = u^{2} / 2$, so
$\expectile{1/2}(X) = \E X$.
The subtraction of $\eta_\tau(X)$ in
\eqref{eq:bg:expectile-def} is a centering device that makes the
criterion finite under the first-moment condition $\E|X|<\infty$; it
does not affect the minimiser in $\theta$.

The convex objective \eqref{eq:bg:expectile-def} has a unique
minimiser, and its first-order condition reduces to the implicit
equation
\begin{equation}
\label{eq:bg:expectile-foc}
  \tau\, \E\!\big[(X - \expectile{\tau}(X))_{+}\big]
  \;=\;
  (1 - \tau)\, \E\!\big[(\expectile{\tau}(X) - X)_{+}\big],
\end{equation}
where $u_{+} = \max(u, 0)$. Thus $\expectile{\tau}(X)$ is the unique
point where the upper and lower first partial moments, with asymmetric
weights, balance. For a strictly increasing continuous distribution,
$\Quant_X(\tau)$ is equivalently characterised by
$F(\Quant_X(\tau))=\tau$. This can be written in the formally balanced
form
\[
  \tau\, \Prob(X > \Quant_X(\tau))
  =
  (1 - \tau)\, \Prob(X < \Quant_X(\tau)),
\]
but, unlike \eqref{eq:bg:expectile-foc}, this identity expresses a
probability balance rather than a balance of tail magnitudes.

\subsection{Basic properties}

The following properties are immediate from
\eqref{eq:bg:expectile-def} or \eqref{eq:bg:expectile-foc}; we collect
them without proof and refer to \textcite[\S2]{NeweyPowell1987} and
\textcite[Sections~2--3]{BelliniKlarMullerRosazzaGianin2014} for
details.

\begin{proposition}
\label{prop:bg:expectile-properties}
Let $X$ and $Y$ be real-valued random variables with $\E |X|, \E |Y|
< \infty$. For each $\tau \in (0, 1)$:
\begin{enumerate}[label=(\roman*)]
\item \textbf{Existence and uniqueness.} $\expectile{\tau}(X)$ exists
      and is uniquely defined.
\item \textbf{Law invariance.} If $X$ and $Y$ have the same
      distribution, then $\expectile{\tau}(X)=\expectile{\tau}(Y)$.
\item \textbf{Mean recovery.} $\expectile{1/2}(X) = \E X$.
\item \textbf{Location equivariance.}
      $\expectile{\tau}(X + c) = \expectile{\tau}(X) + c$ for
      $c \in \R$.
\item \textbf{Positive homogeneity.}
      $\expectile{\tau}(\lambda X) = \lambda\, \expectile{\tau}(X)$
      for $\lambda \ge 0$.
\item \textbf{Sign reflection.}
      $\expectile{\tau}(-X) = -\expectile{1 - \tau}(X)$.
\item \textbf{Monotonicity in $\tau$.} The map
      $\tau \mapsto \expectile{\tau}(X)$ is nondecreasing on
      $(0, 1)$, and is strictly increasing when $X$ is nondegenerate,
      with the boundary limits
      $\expectile{\tau}(X) \to \mathrm{ess}\inf(X)$ as $\tau \to 0$
      and $\expectile{\tau}(X) \to \mathrm{ess}\sup(X)$ as
      $\tau \to 1$ (infinite limits allowed).
\item \textbf{Monotonicity in the argument.} If $X \le Y$ almost
      surely, then $\expectile{\tau}(X) \le \expectile{\tau}(Y)$.
\end{enumerate}
\end{proposition}

For losses, expectiles are translation equivariant, positively
homogeneous, and monotone. For positions, the induced risk functional
$\rho_\tau(Y)=\expectile{\tau}(-Y)$ is cash-additive with the usual sign
convention $\rho_\tau(Y+c)=\rho_\tau(Y)-c$ and is monotone decreasing in
positions. The remaining coherence axiom in
\textcite{ArtznerDelbaenEberHeath1999} is subadditivity, to which we
turn next.

\subsection{Coherence and elicitability}

In \Cref{sec:bg:rv,sec:bg:tail-index-quantile,sec:bg:extreme-expectile}
and in the contribution chapters, $X$ denotes a loss and high risk is
represented by the upper expectile $\expectile{\tau}(X)$ with
$\tau \uparrow 1$. The coherence literature is often written for
financial positions, where positive values are gains and larger values
are better. In this subsection only, write $Y$ for such a position and
define the capital requirement
\[
  \rho_{\tau}(Y) \;=\; \expectile{\tau}(-Y).
\]
Subadditivity of $\rho_{\tau}$ requires
$\rho_{\tau}(Y + Z) \le \rho_{\tau}(Y) + \rho_{\tau}(Z)$ for all
positions $Y, Z$.

\begin{theorem}[\textcite{BelliniKlarMullerRosazzaGianin2014}]
\label{thm:bg:expectile-coherence}
The functional $\rho_{\tau}(Y) = \expectile{\tau}(-Y)$ is a coherent
risk measure if and only if $\tau \ge 1/2$.
\end{theorem}

Recall that a real-valued functional $T$ on a class of distributions
is \emph{elicitable} \parencite{Gneiting2011} if there exists a
strictly consistent scoring function $S : \R \times \R \to [0,\infty)$
whose expected score is uniquely minimised by reporting $T(P)$ for
every $P$ in the class. \textcite[\S3.3]{Gneiting2011} shows
that the $\tau$-expectile is elicitable on classes with finite first
moment; the familiar asymmetric squared score
$S_{\tau}(\theta, x)=\eta_{\tau}(x-\theta)$ is strictly consistent on
classes with finite second moment, while the general characterisation
uses Bregman-type scores with the corresponding integrability
conditions. Thus competing expectile forecasts can be compared by
empirical averages of strictly consistent scores. Taken together with
\Cref{thm:bg:expectile-coherence}, this places expectiles in a
privileged position among risk measures. On suitable convex classes of
distributions and under the usual monetary-risk-measure convention,
Bellini et al.\ show that, within the class of M-quantiles, the
coherent risk measures are exactly the expectiles, while
\textcite{Ziegel2016} characterises elicitable law-invariant coherent
risk measures as certain expectiles, with the expectation appearing as
the special case.

\subsection{Choquet representation}

\textcite{EhmGneitingJordanKrueger2016} place expectiles within a
broader theory of consistent scoring functions. They show that regular
scoring functions consistent for expectiles admit a mixture
representation over elementary scores, analogous in spirit to the
corresponding representation for quantile scores. After a normalisation
that removes trivial dilations of the scoring class, these elementary
scores are extreme points, yielding the advertised Choquet-type
representation. This is relevant for forecast comparison and
backtesting, but it is not used in the asymptotic theory below; all
subsequent arguments rely only on the expectile first-order condition
and the extreme expectile--quantile equivalence.

\section{Extreme expectile inference: the iid heavy-tailed baseline}
\label{sec:bg:extreme-expectile}

We now restrict to a heavy-tailed loss $X$ satisfying
\Cref{cond:bg:first-order} with extreme-value index $\gamma \in (0,1)$,
and consider inference for $\expectile{\tau}$ at levels $\tau$
approaching one. We distinguish, throughout, between two asymptotic
regimes:
\begin{itemize}
\item \emph{intermediate-level}: $\tau_{n} \to 1$ with
      $n(1 - \tau_{n}) \to \infty$. There are still many observations
      above $\Quant_{X}(\tau_{n})$, and central limit theorems hold at
      the rate $\sqrt{n(1 - \tau_{n})}$.
\item \emph{extreme-level}: $\tau_{n}^{\prime} \to 1$ with
      $n(1 - \tau_{n}^{\prime}) \to c \in [0, \infty)$. There are at
      most finitely many observations above
      $\Quant_{X}(\tau_{n}^{\prime})$, and direct estimation is
      infeasible; extrapolation is required.
\end{itemize}
The results in this section are due, in the iid setting, to
\textcite{DaouiaGirardStupfler2018,DaouiaGirardStupfler2019,DaouiaGirardStupfler2020};
the time-series extension that we have used to organise the
exposition is \textcite{DavisonPadoanStupfler2023}, and we shall
borrow Proposition~C.6 of that paper's supplement to formulate the
abstract plug-in lemma that the contribution chapters use.

\subsection{The expectile--quantile asymptotic equivalence}

The single most important fact for extreme expectile inference is
that, in the heavy-tailed regime, the expectile is asymptotically
proportional to the quantile of the same level.

\begin{proposition}[Asymptotic equivalence;
                    \textcite{BelliniKlarMullerRosazzaGianin2014},
                    \textcite{DaouiaGirardStupfler2019}, Corollary~1 with $p=2$]
\label{prop:bg:expectile-quantile}
Suppose $X$ satisfies \Cref{cond:bg:first-order} with index
$\gamma \in (0, 1)$ and $\E X^{-} < \infty$. Then
\begin{equation}
\label{eq:bg:expectile-quantile-equivalence}
  \frac{\expectile{\tau}(X)}{\Quant_{X}(\tau)}
  \;\xrightarrow{\tau \uparrow 1}\;
  \psifn(\gamma) \;\coloneqq\; \big(\gamma^{-1} - 1\big)^{-\gamma}.
\end{equation}
\end{proposition}

The factor $\psifn(\gamma)$ is the linchpin of all extreme-expectile
inference: it converts asymptotic statements about extreme quantiles
into asymptotic statements about extreme expectiles, and vice versa.
The first-order equivalence in
\Cref{prop:bg:expectile-quantile} requires $\gamma < 1$ together with
the integrability of the lower tail; the stronger range
$\gamma < 1/2$ enters the CLT-level inference below.
The function $\psifn$ is continuous on $(0, 1)$, with
$\psifn(\gamma) \to 1$ as $\gamma \downarrow 0$ and
$\psifn(\gamma) \to \infty$ as $\gamma \uparrow 1$. It is \emph{not}
monotone: its logarithmic derivative
\begin{equation}
\label{eq:bg:m-function}
  m(\gamma)
  \;\coloneqq\;
  \frac{d}{d\gamma}\log\psifn(\gamma)
  \;=\;
  \frac{1}{1 - \gamma} - \log\big(\gamma^{-1} - 1\big)
\end{equation}
changes sign at the value $\gamma^{\star} \approx 0.2178$ solving
$m(\gamma^{\star}) = 0$, so $\psifn$ decreases on $(0, \gamma^{\star})$
to a minimum $\psifn(\gamma^{\star}) \approx 0.757$ and increases on
$(\gamma^{\star}, 1)$; in particular it returns to $\psifn = 1$ at
$\gamma = 1/2$. Only for tail indices beyond $\gamma^{\star}$ does a
heavier tail move the expectile closer to, and for $\gamma>1/2$ above,
the same-level quantile.

Under the second-order condition $\mathcal C_{2}(\gamma, \rho, A)$ with
$\gamma \in (0, 1/2)$ and \Cref{cond:bg:lower-tail-moment}, and assuming
that $F$ is strictly increasing in a neighbourhood of its upper endpoint,
the first-order equivalence \eqref{eq:bg:expectile-quantile-equivalence}
sharpens to the second-order expansion
\parencite[Proposition~1(i)]{DaouiaGirardStupfler2020}
\begin{equation}
\label{eq:bg:c-gamma-rho}
  \begin{aligned}
  \frac{\expectile{\tau}(X)}{\Quant_{X}(\tau)}
  &=
  \psifn(\gamma)\Big[
      1 + c(\gamma, \rho)\,A\big((1 - \tau)^{-1}\big)
      + \gamma(\gamma^{-1} - 1)^{\gamma}
        \frac{\E X}{\Quant_{X}(\tau)}
  \\
  &\hspace{5.6em}
      + o\big(A((1 - \tau)^{-1})\big)
      + o\big(\Quant_{X}(\tau)^{-1}\big)
    \Big],
  \\
  c(\gamma, \rho)
  &\coloneqq
  \frac{(\gamma^{-1} - 1)^{-\rho}}{1 - \gamma - \rho}
  + \frac{(\gamma^{-1} - 1)^{-\rho} - 1}{\rho}.
  \end{aligned}
\end{equation}
The displayed form is the version used in the thesis: it combines the
second-order tail expansion with the lower-tail moment condition
sufficient to control the non-tail contribution to the expectile
equation. At the endpoint $\rho=0$, the second term in $c(\gamma,\rho)$
is read by continuity as $\rho\to0$, so that
$c(\gamma,0)=m(\gamma)$. The constant $c(\gamma, \rho)$ measures the
deterministic gap between the extreme expectile and the same-level
quantile beyond the leading factor $\psifn(\gamma)$. The term
$\gamma(\gamma^{-1}-1)^{\gamma}\E X/\Quant_{X}(\tau)$ is the explicit
first-moment contribution in the expansion of
\textcite{DaouiaGirardStupfler2020}; it is often abbreviated below as an
$O(\Quant_{X}(\tau)^{-1})$ contribution, because the lower-tail moment
assumption ensures that it is well controlled. Under the moment condition
$\sqrt{n(1 - \tau_{n})}\,/\Quant_{X}(\tau_{n}) \to 0$ adopted below it
is $\sqrt{n(1 - \tau_{n})}$-negligible, so the $A$-driven term alone
governs the asymptotic bias of the quantile-based estimator. This
source is the version used here because it is stated under the
$\rho\le0$ convention of \Cref{cond:bg:second-order}; the earlier
\textcite[Corollary~1]{DaouiaGirardStupfler2018} gives the same
constant under its stricter $\rho<0$ formulation.

\subsection{Two families of estimators at intermediate levels}

Two natural estimators of $\expectile{\tau_{n}}$ at an intermediate
level $\tau_{n}$ arise from the asymmetric-LS definition
\eqref{eq:bg:expectile-def} on the one hand, and from the asymptotic
equivalence \eqref{eq:bg:expectile-quantile-equivalence} on the other.

\paragraph{LAWS direct estimator.}
The \emph{asymmetric-least-squares} (LAWS) estimator replaces the
population minimisation \eqref{eq:bg:expectile-def} with its empirical
analogue:
\begin{equation}
\label{eq:bg:laws}
  \tildexpectile{\tau_{n}}
  \;=\;
  \argmin_{\theta \in \R}\,
  \sum_{i=1}^{n} \eta_{\tau_{n}}(X_{i} - \theta).
\end{equation}
The estimator $\tildexpectile{\tau_{n}}$ can be computed with standard
expectile routines; \textcite[\S3.1.2]{DaouiaGirardStupfler2018} use
the \texttt{expectile} function of the \texttt{expectreg} package.

\begin{theorem}[Intermediate LAWS CLT;
                \textcite{DaouiaGirardStupfler2018}, Theorem~2]
\label{thm:bg:laws-iid}
Assume \Cref{cond:bg:first-order} with
$\gamma \in (0,1/2)$, and assume
\Cref{cond:bg:lower-tail-moment}. Let
$\tau_{n} \to 1$ with $n(1 - \tau_{n}) \to \infty$. Then
\[
  \sqrt{n(1 - \tau_{n})}\!
  \left( \frac{\tildexpectile{\tau_{n}}}{\expectile{\tau_{n}}} - 1
  \right)
  \;\todist\;
  \mathcal N\!\left(0,\; \frac{2 \gamma^{3}}{1 - 2\gamma}\right).
\]
\end{theorem}

\paragraph{Quantile-based plug-in estimator.}
A second family substitutes estimators of $\gamma$ and of the
intermediate quantile $\Quant_{X}(\tau_{n})$ into the
asymptotic-equivalence formula
\eqref{eq:bg:expectile-quantile-equivalence}:
\begin{equation}
\label{eq:bg:qb}
  \hatexpectile{\tau_{n}}
  \;=\;
  \psifn\!\big(\hat\gamma_{n}\big)\, \hatQuant(\tau_{n}),
\end{equation}
where $\hatQuant(\tau_{n}) = X_{n - \lfloor n(1 - \tau_{n}) \rfloor : n}$
is the empirical quantile (an intermediate-order statistic) and
$\hat\gamma_{n}$ is a tail-index estimator admitting the required
$\sqrt{n(1 - \tau_{n})}$ limit theory. The natural choice is the Hill
estimator of \eqref{eq:bg:hill} with effective sample size
$k = \lfloor n(1 - \tau_{n})\rfloor$.

\begin{theorem}[Intermediate QB CLT;
                \textcite{DaouiaGirardStupfler2020}, Theorem~4 with $\beta=1$]
\label{thm:bg:qb-iid}
Assume $\mathcal C_{2}(\gamma, \rho, A)$ with
$\gamma \in (0,1/2)$ and \Cref{cond:bg:lower-tail-moment}. Assume also
that $F$ is strictly increasing in a neighbourhood of its upper endpoint.
Let
$\tau_n \to 1$ with $n(1-\tau_n)\to\infty$ and suppose
\[
  \sqrt{n(1-\tau_n)}\, A\big((1-\tau_n)^{-1}\big)\to \lambda\in\R,
  \qquad
  \frac{\sqrt{n(1-\tau_n)}}{\Quant_X(\tau_n)}\to 0 .
\]
With $\hat\gamma_{n}$ the Hill estimator at effective sample size
$k = \lfloor n(1 - \tau_{n}) \rfloor$,
\[
  \sqrt{n(1 - \tau_{n})}\!
  \left(\frac{\hatexpectile{\tau_{n}}}{\expectile{\tau_{n}}} - 1
  \right)
  \;\todist\;
  \mathcal N\!\left(\Big(\frac{m(\gamma)}{1 - \rho}
                          - c(\gamma, \rho)\Big)\lambda,\;
                    \gamma^{2}\big(1 + m(\gamma)^{2}\big)\right),
\]
with $m(\gamma)$ as in \eqref{eq:bg:m-function} and $c(\gamma, \rho)$
the second-order expectile--quantile constant of
\eqref{eq:bg:c-gamma-rho}. The asymptotic bias has two pieces: the
Hill bias $\lambda / (1 - \rho)$ propagated through the logarithmic
derivative $m(\gamma)$ of $\psifn$, and the deterministic
expectile--quantile gap $-c(\gamma, \rho)\,\lambda$. The empirical
quantile $\hatQuant(\tau_{n})$ contributes no asymptotic mean term at
the $\sqrt{n(1-\tau_n)}$ scale when centered at
$\Quant_X(\tau_n)$; the condition
$\sqrt{n(1 - \tau_{n})}\,/\Quant_{X}(\tau_{n}) \to 0$ inherited from
the theorem ensures that the moment-term contribution to the
expectile--quantile gap is negligible.
\end{theorem}

This is the $\lambda_2=0$ specialisation of the more general
intermediate QB CLT of \textcite{DaouiaGirardStupfler2020}, where
\[
  \lambda_2
  =
  \lim_{n\to\infty}
  \frac{\sqrt{n(1-\tau_n)}}{\Quant_X(\tau_n)}.
\]
In the general statement, the asymptotic mean contains an additional
term of order $\lambda_2$ depending on the first moment of $X$. The
present thesis imposes
$\sqrt{n(1-\tau_n)}/\Quant_X(\tau_n)\to0$ because this is the regime
needed later to isolate the second-order tail bias.
\textcite[Corollary~2]{DaouiaGirardStupfler2018} is the original iid
QB statement under the stricter $\rho<0$ convention; the 2020 tail
expectile process result is the version that matches
\Cref{cond:bg:second-order} at the intermediate level.

\paragraph{Comparison.}
The LAWS estimator has asymptotic variance
$2 \gamma^{3} / (1 - 2 \gamma)$, which is independent of the choice
of tail-index estimator. The QB estimator's variance
$\gamma^{2}(1 + m(\gamma)^{2})$ depends on the tail-index estimator
used and reflects the contribution of $\hat\gamma_{n}$ via the term
$m(\gamma)^{2}$. Numerical comparison over $\gamma \in (0, 1/2)$
reveals that no single estimator dominates the other across the whole
range: the two variances cross once, at
$\gamma \approx 0.2623$. For lighter heavy tails, i.e.\ small
$\gamma$, the LAWS variance is smaller. For sufficiently heavy tails,
and especially as $\gamma \uparrow 1/2$, the QB variance is smaller
because the LAWS variance diverges. The thesis works with the QB
construction throughout not because it uniformly dominates LAWS, but
because it composes directly with the pooled $(\hat\gamma,\hat Q)$
inputs of the pooling framework. This comparison concerns direct
intermediate-level expectile estimation; in the very-extreme
extrapolation regime below, the first-order limit is instead driven by
the tail-index estimator.

\subsection{Weissman extrapolation to very-extreme expectile levels}
\label{sec:bg:weissman-expectile}

For levels $\tau_{n}^{\prime}$ so close to one that
$n(1 - \tau_{n}^{\prime}) \to c \in [0, \infty)$, neither
$\tildexpectile{\tau_{n}^{\prime}}$ nor
$\hatexpectile{\tau_{n}^{\prime}}$ provides a consistent direct
estimator of $\expectile{\tau_n'}$.
Extrapolation from an intermediate level $\tau_{n}$ to the extreme
level $\tau_{n}^{\prime}$ is required.

The asymptotic equivalence
\eqref{eq:bg:expectile-quantile-equivalence} and the first-order
identity $U(t y) / U(t) \to y^{\gamma}$ together yield the heuristic
\[
  \frac{\expectile{\tau_{n}^{\prime}}}{\expectile{\tau_{n}}}
  \;\sim\;
  \frac{\Quant_{X}(\tau_{n}^{\prime})}{\Quant_{X}(\tau_{n})}
  \;\sim\;
  \left(\frac{1 - \tau_{n}^{\prime}}{1 - \tau_{n}}\right)^{\!-\gamma},
\]
which suggests the \emph{Weissman-type} extrapolated estimator
\begin{equation}
\label{eq:bg:weissman-expectile}
  \overline{\xi}^{\,\star}_{\tau_{n}^{\prime}}
  \;=\;
  \left(\frac{1 - \tau_{n}^{\prime}}{1 - \tau_{n}}\right)^{\!-\hat\gamma_{n}}\!
  \overline{\xi}_{\tau_{n}},
\end{equation}
where $\overline{\xi}_{\tau_{n}}$ is any consistent intermediate-level
estimator satisfying a $\sqrt{n(1-\tau_n)}$-CLT, such as
$\tildexpectile{\tau_{n}}$ or $\hatexpectile{\tau_{n}}$ under the
conditions above, and $\hat\gamma_{n}$ is a
$\sqrt{n(1-\tau_n)}$-consistent tail-index estimator. We write
$\tilde{\xi}^{\,\star}$ and $\hat{\xi}^{\,\star}$ for the
LAWS-anchored and QB-anchored extrapolations, respectively.

\begin{theorem}[Extreme-level CLT;
                \textcite{DaouiaGirardStupfler2018}, Corollaries~3--4]
\label{thm:bg:weissman-expectile-iid}
Assume $\mathcal C_{2}(\gamma,\rho,A)$ with
$\gamma\in(0,1/2)$, $\rho<0$, and
\Cref{cond:bg:lower-tail-moment}. Assume also that $F$ is strictly
increasing in a neighbourhood of its upper endpoint. Let
$\tau_n\to1$ satisfy $n(1-\tau_n)\to\infty$ and let
$\tau_{n}^{\prime} \to 1$ satisfy
$n(1-\tau'_n)\to c\in[0,\infty)$,
$(1 - \tau_{n}^{\prime})/(1 - \tau_{n}) \to 0$ and
$\sqrt{n(1 - \tau_{n})}\,/\log((1 - \tau_{n})/(1 - \tau_{n}^{\prime}))
\to \infty$. Suppose moreover that
\[
  \sqrt{n(1-\tau_n)}\,A((1-\tau_n)^{-1})\to\lambda\in\R,
  \qquad
  \frac{\sqrt{n(1-\tau_n)}}{\Quant_X(\tau_n)}\to 0 .
\]
If $\hat\gamma_n$ is the Hill estimator and
$\overline\xi_{\tau_n}$ is either the LAWS or QB intermediate
estimator above, then
\[
  \frac{\sqrt{n(1 - \tau_{n})}}
       {\log\!\big((1 - \tau_{n})/(1 - \tau_{n}^{\prime})\big)}
  \!\left(\frac{\overline{\xi}^{\,\star}_{\tau_{n}^{\prime}}}
                {\expectile{\tau_{n}^{\prime}}} - 1\right)
  \;\todist\;
  \mathcal N\!\left(\frac{\lambda}{1 - \rho},\; \gamma^{2}\right).
\]
\end{theorem}

Two remarks are in order. First, the rate
$\sqrt{n(1 - \tau_{n})}\, / \log((1 - \tau_{n})/(1 - \tau_{n}^{\prime}))$
is the classical Weissman rate of \Cref{thm:bg:weissman}: the further
into the tail one extrapolates, the larger the log-denominator and the
slower the rate of convergence. Second, the asymptotic variance is
$\gamma^{2}$ \emph{regardless of whether the anchor estimator is LAWS
or QB}: the contribution of $\overline{\xi}_{\tau_{n}}$ is dominated,
at the extreme level, by that of $\hat\gamma_{n}$ entering the
multiplicative power. This is the central structural feature that
makes extreme-expectile inference tractable; at the extreme level,
asymptotic inference for $\expectile{\tau_{n}^{\prime}}$ reduces to
asymptotic inference for the tail index.

\subsection{The plug-in central limit theorem}
\label{sec:bg:plug-in}

\Cref{thm:bg:laws-iid,thm:bg:qb-iid,thm:bg:weissman-expectile-iid}
together specify a complete asymptotic theory for extreme expectiles
in the iid heavy-tailed setting. For the pooled inference developed
in \Cref{ch:pool-int,ch:pool-ext}, however, we require a more abstract
formulation that decouples the asymptotic distribution of the inputs
$(\hat\gamma_{n}, \overline{\xi}_{\tau_{n}})$ from the way in which
they are obtained. We use the following high-level consequence of
\emph{Proposition~C.6} in the supplement to
\textcite{DavisonPadoanStupfler2023}. It packages the deterministic
expectile-ratio control as an explicit assumption and writes the
normalising rate as a generic sequence $a_n$, which is the form needed
when the intermediate estimator is itself produced by pooling.

\begin{proposition}[Plug-in CLT for extrapolated expectiles;
                    cf.\ \textcite{DavisonPadoanStupfler2023}, Prop.~C.6]
\label{prop:bg:plug-in-clt}
Assume $\mathcal C_2(\gamma,\rho,A)$ with
$\gamma\in(0,1/2)$ and $\rho<0$, and assume
\Cref{cond:bg:lower-tail-moment}.
Let $a_n\to\infty$, let $\hat\gamma_n$ be an estimator of $\gamma$,
and let $\widehat\xi_{\tau_n}^{\,I}$ be an intermediate-level
estimator of $\expectile{\tau_n}$ satisfying the joint asymptotic
representation
\begin{equation}
\label{eq:bg:joint-input}
  a_n
    \begin{pmatrix}
      \hat\gamma_{n} - \gamma \\
      \widehat\xi_{\tau_n}^{\,I} / \expectile{\tau_{n}} - 1
    \end{pmatrix}
  \;\todist\;
  \begin{pmatrix} \Gamma \\ \Delta \end{pmatrix},
\end{equation}
for some random vector $(\Gamma, \Delta)^{\top}$. Let
$\tau_n,\tau'_n\to1$ with $n(1-\tau_n)\to\infty$,
$n(1-\tau'_n)\to c\in[0,\infty)$,
$(1 - \tau_{n}^{\prime})/(1 - \tau_{n}) \to 0$, put
\[
  \ell_n
  \;\coloneqq\;
  \log\!\left(\frac{1-\tau_n}{1-\tau'_n}\right),
\]
and suppose
\[
  \frac{a_n}{\ell_n}\to\infty,\qquad
  a_n A((1-\tau_n)^{-1})\to \lambda_1\in\R,\qquad
  \frac{a_n}{\Quant_X(\tau_n)}\to \lambda_2\in\R .
\]
Assume moreover that $\Prob(\widehat\xi_{\tau_n}^{\,I}>0)\to1$ and
that the deterministic extrapolation error satisfies
\[
  \frac{a_n}{\ell_n}
  \left\{
    \log\expectile{\tau'_n}
    - \log\expectile{\tau_n}
    - \gamma\ell_n
  \right\}
  \;\longrightarrow\; 0.
\]
Then the Weissman-extrapolated expectile
\[
  \widehat\xi_{\tau'_n}^{\,\star}
  =
  \left(\frac{1-\tau'_n}{1-\tau_n}\right)^{-\hat\gamma_n}
  \widehat\xi_{\tau_n}^{\,I}
\]
satisfies
\[
  \frac{a_n}
       {\log\!\big((1 - \tau_{n})/(1 - \tau_{n}^{\prime})\big)}
  \!\left(\frac{\widehat\xi_{\tau'_n}^{\,\star}}
                {\expectile{\tau_{n}^{\prime}}} - 1\right)
  \;\todist\;
  \Gamma.
\]
\end{proposition}

\begin{remark}[Deterministic ratio control]
\label{rem:bg:plugin-ratio-control}
The deterministic log-ratio condition in
\Cref{prop:bg:plug-in-clt} is the high-level form of the extrapolation
bias control. It is precisely the negligibility
\[
  \frac{a_n}{\ell_n}
  \left\{
    \log\expectile{\tau'_n}
    - \log\expectile{\tau_n}
    - \gamma\ell_n
  \right\}
  \;\longrightarrow\; 0.
\]
In the original supplement of
\textcite{DavisonPadoanStupfler2023}, this control is part of the
high-level condition $C^\ast$. In the iid heavy-tailed setting of this
thesis, under the strict upper-tail monotonicity used for the DGS
second-order expansion, it follows from the second-order quantile
expansion and \eqref{eq:bg:c-gamma-rho}: the $A$-term contributes
$O(a_n A((1-\tau_n)^{-1})/\ell_n)=o(1)$ and the lower-tail remainder
contributes $O(a_n/\{\Quant_X(\tau_n)\ell_n\})=o(1)$.
\end{remark}

\Cref{prop:bg:plug-in-clt} is the technical workhorse of the thesis.
The original supplement of \textcite{DavisonPadoanStupfler2023}
states the result under broader high-level assumptions; the version
above is deliberately restricted to the heavy-tailed iid regime used
throughout this thesis; \Cref{rem:bg:plugin-ratio-control} records the
deterministic step that links this concrete regime to the abstract
formulation.
It says that the asymptotic distribution of the extrapolated expectile
at very-extreme levels is governed \emph{only} by the asymptotic
distribution of the tail-index estimator: the contribution of the
intermediate-level expectile estimator $\widehat\xi_{\tau_n}^{\,I}$
washes out through the log-rate. In \Cref{ch:pool-ext} we shall feed
the pooled inputs
$(\hat\gamma_{n}(\bm\omega), \overline{\xi}^{\mathrm{pool}}_{\tau_{n}})$
of \Cref{ch:pool-int} through this proposition to obtain the CLT for
the pooled extrapolated expectile in essentially one step.

\section{The optimal pooling framework}
\label{sec:bg:pooling}

We now turn to the pooling framework of
\textcite{DaouiaPadoanStupfler2021}, which the rest of the thesis
extends from quantiles to expectiles. We refer throughout to the final
Bernoulli article version listed in the bibliography; the earlier arXiv
preprint circulated under the same arXiv identifier but differs in some
display notation. The framework concerns inference from $m$ samples for
the tail index and, under an additional
tail-homoskedasticity condition, for the extreme quantile. The samples
may have
heterogeneous sizes, heterogeneous second-order properties, and
non-trivial cross-sample tail dependence.

\subsection{Setup}

Throughout \Cref{sec:bg:pooling}, unless explicitly stated otherwise,
$m$ is fixed while $n=\sum_{j=1}^{m}n_j\to\infty$. This is the
finite-$m$ regime imported into \Cref{ch:pool-int,ch:pool-ext}.
Let $\bm X_i=(X_{i,1},\ldots,X_{i,m})^\top$, $i\ge1$, be iid copies of a
generic $m$-dimensional random vector
$\bm X = (X_{1}, \ldots, X_{m})^{\top}$. Margin $j$ is observed for
$i=1,\ldots,n_j$, so if $n_j\ne n_\ell$ the two margins overlap only for
the first $\min(n_j,n_\ell)$ vector observations. The $j$th sample
therefore comprises
$n_{j}$ observations $X_{1, j}, X_{2, j}, \ldots, X_{n_{j}, j}$ with
continuous, right-heavy-tailed marginal distribution function $F_{j}$,
tail quantile function $U_{j} = (1 / \overline F_{j})^{\leftarrow}$,
and extreme-value index $\gamma_{j} > 0$. The order statistics of the
$j$th sample are denoted
$X_{1 : n_{j}, j} \le \cdots \le X_{n_{j} : n_{j}, j}$. The total
sample size is $n = \sum_{j=1}^{m} n_{j}$.

Tail inference in sample $j$ uses its top $k_{j} + 1$ order statistics,
where the effective sample size $k_{j} = k_{j}(n)$ satisfies
$k_{j} \to \infty$ and $k_{j} / n_{j} \to 0$ as $n \to \infty$. The
total effective sample size is $k = \sum_{j=1}^{m} k_{j}$. Assume the
asymptotic-proportionality conditions
\begin{equation}
\label{eq:bg:pool-rates}
  \frac{n_{1}}{n_{j}} \;\to\; b_{j} \in (0, \infty),
  \qquad
  \frac{k_{1}}{k_{j}} \;\to\; c_{j} \in (0, \infty),
\end{equation}
with $b_{1} = c_{1} = 1$. The marginal $F_{j}$ satisfies
$\mathcal C_{2}(\gamma_{j}, \rho_{j}, A_{j})$
(\Cref{cond:bg:second-order}). The thesis works under the
\emph{common-tail-index} assumption
\[
  (H_{\gamma}) \qquad
  \gamma_{1} \;=\; \cdots \;=\; \gamma_{m} \;=\; \gamma,
\]
which is the natural setup when the $m$ samples are believed to come
from the same underlying tail behaviour.
All common-$\gamma$ vector limits below are expressed on the aggregate
scale $\sqrt{k}$; the constants $c_j$ translate the margin-specific
$\sqrt{k_j}$ rates onto this common scale.

The cross-sample asymptotic dependence is captured by a pairwise
\emph{tail-copula} structure
\parencite[\S2.1]{DaouiaPadoanStupfler2021}.

\paragraph{Condition $\mathcal J(\bm R)$ (tail-copula condition).}
For each pair $(j, \ell)$ with $j \ne \ell$, there exists a function
$\Rjl : [0, \infty]^{2} \setminus \{(\infty, \infty)\} \to [0, \infty)$
such that, for every
$(x_{j}, x_{\ell}) \in [0, \infty]^{2} \setminus \{(\infty,\infty)\}$,
\[
  \lim_{s \to \infty}
    s\, \Prob\!\big(\,
      1 - F_{j}(X_{j}) \le x_{j}/s,\;
      1 - F_{\ell}(X_{\ell}) \le x_{\ell}/s\,\big)
  \;=\; \Rjl(x_{j}, x_{\ell}).
\]

The function $\Rjl$ is the bivariate tail copula of
$(X_{j}, X_{\ell})$ \parencite[\S8.2]{deHaanFerreira2006}; it measures
the rate at which the joint upper-tail probability of
$(X_{j}, X_{\ell})$ decays in comparison with the marginals. The
case $\Rjl \equiv 0$ corresponds to \emph{asymptotic tail
independence} of $X_{j}$ and $X_{\ell}$, which is in particular
implied by independence of the samples themselves --- the setting of
this thesis.
This formulation covers the general pooling framework in which
cross-sample extremal dependence may be represented through a tail
copula. In the distributed-inference case used later in the thesis,
the samples are independent marginal subsamples and all pairwise
tail-copula terms vanish.

\subsection{The pooled Hill estimator}

Each sample contributes a Hill estimator $\hat\gamma_{j}(k_{j})$ as
in \eqref{eq:bg:hill}; we collect these into the random vector
\[
  \hat{\bm\gamma}_{n}
  \;=\;
  \big(\hat\gamma_{1}(k_{1}), \ldots, \hat\gamma_{m}(k_{m})\big)^{\top}.
\]
Given a weight vector $\bm\omega \in \R^{m}$ with $\bm\omega^{\top}
\bm 1 = 1$, the \emph{pooled Hill estimator} is the weighted arithmetic
combination
\begin{equation}
\label{eq:bg:pooled-hill}
  \hat\gamma_{n}(\bm\omega)
  \;\coloneqq\;
  \bm\omega^{\top}\hat{\bm\gamma}_{n}
  \;=\;
  \sum_{j=1}^{m} \omega_{j}\,\hat\gamma_{j}(k_{j}).
\end{equation}
The constraint $\bm\omega^{\top}\bm 1 = 1$ ensures consistency of
$\hat\gamma_{n}(\bm\omega)$ for the common $\gamma$ under
$(H_{\gamma})$. The naive choice $\bm\omega = m^{-1}\bm 1$
recovers a simple arithmetic average; the choice $\bm\omega \propto
(k_{1}, \ldots, k_{m})$ recovers effective-sample-size weighting. The
goal of the optimal-pooling literature is to identify weight choices
that improve on these heuristics in terms of asymptotic variance or
AMSE.

The fundamental asymptotic result is the joint CLT for the marginal
Hill estimators.

\begin{theorem}[\textcite{DaouiaPadoanStupfler2021}, Theorem~1]
\label{thm:bg:pool-hill}
Assume $\mathcal C_{2}(\gamma_{j}, \rho_{j}, A_{j})$ for every $j$,
the tail-copula condition $\mathcal J(\bm R)$, the
proportionality conditions \eqref{eq:bg:pool-rates}, and
$\sqrt{k_{j}}\, A_{j}(n_{j}/k_{j}) \to \lambda_{j} \in \R$ for every
$j$. Then
\[
  \big(\sqrt{k_{1}}(\hat\gamma_{1}(k_{1}) - \gamma_{1}),
       \ldots,
       \sqrt{k_{m}}(\hat\gamma_{m}(k_{m}) - \gamma_{m})\big)^{\top}
  \;\todist\;
  \mathcal N(\bm B, \bm V),
\]
with $\bm B = \big(\lambda_{j}/(1 - \rho_{j})\big)_{j=1}^{m}$ and
$\bm V$ the symmetric matrix whose entries are
\[
  V_{j, \ell}
  \;=\;
  \begin{cases}
    \gamma_{j}^{2}, & j = \ell, \\[2pt]
    \displaystyle
    \frac{\gamma_{j}\gamma_{\ell}}
         {\max(b_j,b_\ell)\sqrt{c_jc_\ell}}\,
      \Rjl(b_jc_\ell,\, b_\ell c_j),
        & j < \ell.
  \end{cases}
\]
The matrix is symmetric, so $V_{\ell,j}=V_{j,\ell}$ for $\ell>j$.
Suppose, in addition, the common-$\gamma$ condition
$(H_{\gamma})$. Then the same marginal estimators, now viewed on the
common $\sqrt{k}$ scale, satisfy
\[
  \sqrt{k}\,\big(\hat{\bm\gamma}_{n}-\gamma\bm 1\big)
  \;\todist\;
  \mathcal N(\Bc,\Vc),
\]
where
\[
  (\Bc)_j
  =
  \sqrt{c_j\sum_{i=1}^{m} c_i^{-1}}\,
  \frac{\lambda_j}{1-\rho_j}
  \qquad j=1,\ldots,m.
\]
The covariance matrix $\Vc$ is given by
\[
  (\Vc)_{j,\ell}
  =
  \left(\sum_{i = 1}^{m} c_i^{-1}\right)
  \begin{cases}
    \gamma^{2}c_j, & j=\ell, \\[2pt]
    \displaystyle
    \frac{\gamma^{2}}{\max(b_j,b_\ell)}\,
      \Rjl(b_jc_\ell,b_\ell c_j),
      & j<\ell,
  \end{cases}
\]
with $(\Vc)_{\ell,j}=(\Vc)_{j,\ell}$ for $\ell>j$.
Consequently, for any $\bm\omega$ with $\bm\omega^{\top}\bm 1 = 1$,
\[
  \sqrt{k}\,\big(\hat\gamma_{n}(\bm\omega) - \gamma\big)
  \;\todist\;
  \mathcal N\!\big(\bm\omega^{\top}\Bc,\;
                    \bm\omega^{\top}\Vc\bm\omega\big).
\]
\end{theorem}

In the case of \emph{independent} samples ($\Rjl \equiv 0$ for every
$j \ne \ell$), the off-diagonal entries of $\Vc$ vanish identically
and $\Vc$ becomes the diagonal matrix
\begin{equation}
\label{eq:bg:V_c-diagonal}
  \Vc
  \;=\;
  \gamma^{2}\!\left(\sum_{i = 1}^{m} c_{i}^{-1}\right)
    \mathrm{diag}(c_{1}, \ldots, c_{m}).
\end{equation}
\Cref{eq:bg:V_c-diagonal} is the form of $\Vc$ that the thesis will
exploit in \Cref{ch:pool-int,ch:pool-ext}.

\subsection{Variance- and AMSE-optimal weights}

Given \Cref{thm:bg:pool-hill}, two natural choices of $\bm\omega$
emerge as solutions to convex quadratic programmes.
Unless explicitly stated otherwise, weights are allowed to be real and
are constrained only by $\bm\omega^\top\bm 1=1$; the optimal weights
therefore need not be convex weights.

\begin{proposition}[Optimal weights;
                    \textcite{DaouiaPadoanStupfler2021}, Theorem~1 and \S2.2]
\label{prop:bg:optimal-weights}
Suppose $\Vc$ is positive definite. Under the common-$\gamma$
condition $(H_{\gamma})$:
\begin{enumerate}[label=(\roman*)]
\item The variance-optimal weights, minimising
      $\bm\omega^{\top}\Vc\bm\omega$ subject to
      $\bm\omega^{\top}\bm 1 = 1$, are
      \[
        \bm\omega^{\mathrm{Var}}
        \;=\;
        \frac{\Vc^{-1}\bm 1}{\bm 1^{\top}\Vc^{-1}\bm 1},
      \]
      with corresponding asymptotic variance
      $\bm\omega^{\mathrm{Var},\top}\Vc\bm\omega^{\mathrm{Var}}
        = (\bm 1^{\top}\Vc^{-1}\bm 1)^{-1}$.
\item The AMSE-optimal weights, minimising
      $(\bm\omega^{\top}\Bc)^{2} + \bm\omega^{\top}\Vc\bm\omega$
      subject to $\bm\omega^{\top}\bm 1 = 1$, are
      \[
        \bm\omega^{\mathrm{AMSE}}
        \;=\;
        \frac{(1 + \Bc^{\top}\Vc^{-1}\Bc)\,\Vc^{-1}\bm 1
              - (\bm 1^{\top}\Vc^{-1}\Bc)\,\Vc^{-1}\Bc}
             {(1 + \Bc^{\top}\Vc^{-1}\Bc)(\bm 1^{\top}\Vc^{-1}\bm 1)
              - (\bm 1^{\top}\Vc^{-1}\Bc)^{2}}.
      \]
\end{enumerate}
\end{proposition}

In the independent-samples case, $\Vc$ is diagonal
(\Cref{eq:bg:V_c-diagonal}) and the variance-optimal weights reduce
to
\begin{equation}
\label{eq:bg:V-opt-distrib}
  \omega_{j}^{\mathrm{Var}}
  \;=\;
  \Bigg(\sum_{i = 1}^{m} c_{i}^{-1}\Bigg)^{\!\!-1}\!
  c_{j}^{-1}
  \;\sim\;
  \frac{k_{j}}{k},
\end{equation}
where the equivalence uses $c_{i}^{-1} = \lim k_{i}/k_{1}$. Thus the
samples contribute to the pooled tail index in proportion to their
effective sample size, the same rule that one would expect from the
unfeasible Hill estimator built from the entire combined sample.

In practice $\Vc$ and $\Bc$ must be replaced by consistent estimators
$\hatVc$ and $\hatBc$ to render the optimal weights data-driven. The
estimators are built from the marginal $\hat\gamma_{j}$, estimators of
the second-order quantities entering $A_j$ and $\rho_j$, and a
plug-in estimator $\widehat{\Rjl}$ of the bivariate tail copula
\parencite[\S2.2]{DaouiaPadoanStupfler2021}. The composite estimator
$\hat\gamma_{n}(\hat{\bm\omega})$ is then $\sqrt{k}$-asymptotically
equivalent to $\hat\gamma_{n}(\bm\omega)$ provided
$\hat{\bm\omega}^{\top}\bm 1=1$ exactly and
$\hat{\bm\omega} \toprob \bm\omega$. If exact normalisation is not
imposed, one must instead require
$\sqrt{k}\,\abs{\hat{\bm\omega}^{\top}\bm 1-1}\toprob0$.
In applications, the main additional difficulty is not the quadratic
minimisation itself but the stable estimation of $\Vc$, especially its
off-diagonal tail-copula entries.

\subsection{Weighted geometric pooling of extreme quantiles}

For estimation of an extreme quantile $q(1 - p)$ with $p = p(n) \to
0$, each sample contributes a Weissman estimator
\[
  \hat q_{j}^{\star}(1 - p \mid k_{j})
  \;=\;
  \left(\frac{k_{j}}{n_{j}\, p}\right)^{\!\hat\gamma_{j}(k_{j})}\!
  X_{n_{j} - k_{j} : n_{j},\,j}.
\]
The aggregation of these per-sample Weissman estimators must be
adapted to their multiplicative-power structure
(\Cref{eq:bg:weissman}), and is done by \emph{weighted geometric}
--- rather than arithmetic --- averaging. Under the additional
assumption that the samples have asymptotically equivalent extreme
quantiles, formalised as the asymptotic-equivalence condition
\begin{condition}[Tail homoskedasticity $(\mathcal H)$]
\label{cond:bg:tail-homoskedasticity}
For every $j \ne \ell$, $U_{j}(t) / U_{\ell}(t) \to 1$ as $t \to
\infty$.
\end{condition}
The terminology reflects the fact that the marginal tail scales
are asymptotically equal. Equivalently,
$\Quant_j(1-p)/\Quant_\ell(1-p)\to1$ as $p\downarrow0$. The notation in
\textcite{DaouiaPadoanStupfler2021} writes this population extreme
quantile as $q_j(1-p)$; here we keep the thesis-wide notation
$\Quant_j(1-p)$.
The pooled extreme-quantile estimator of
\textcite[\S2.3]{DaouiaPadoanStupfler2021} is
\begin{equation}
\label{eq:bg:pooled-weissman}
  \hatWei_{n}(1 - p \mid \bm\omega)
  \;\coloneqq\;
  \prod_{j = 1}^{m}
    \big[\hat q_{j}^{\star}(1 - p \mid k_{j})\big]^{\omega_{j}}.
\end{equation}
The reason for the geometric form is visible from the log-scale
identity
$\log \hat q_{j}^{\star}(1 - p \mid k_{j})
  = \log(k_{j}/(n_{j} p))\,\hat\gamma_{j}(k_{j})
    + \log X_{n_{j} - k_{j} : n_{j},\,j}$:
a weighted sum on the log scale puts the per-sample
$\hat\gamma_{j}(k_{j})$ on the standard linear scale, which is what
the joint CLT of \Cref{thm:bg:pool-hill} addresses.
Under the proportionality conditions \eqref{eq:bg:pool-rates},
$\log(k_j/(n_jp))=\log(k/(np))+O(1)$, so the per-sample logarithmic
normalisers are asymptotically equivalent to the common normaliser in
the diverging-log regime.

\begin{theorem}[\textcite{DaouiaPadoanStupfler2021}, Theorem~2]
\label{thm:bg:pool-weissman}
Assume the hypotheses of \Cref{thm:bg:pool-hill}, the common-$\gamma$
condition $(H_{\gamma})$, the tail-homoskedasticity
condition $(\mathcal H)$, and, in addition, $\rho_j<0$ for every
$j=1,\ldots,m$. Let $p = p(n) \to 0$ with
$k/(np) \to \infty$ and $\sqrt{k}/\log(k/(np)) \to \infty$. Then for
any $\bm\omega$ with $\bm\omega^{\top}\bm 1 = 1$ and any
$j \in \{1, \ldots, m\}$,
\[
  \frac{\sqrt{k}}{\log\!\big(k/(np)\big)}
  \!\left(\frac{\hatWei_{n}(1 - p \mid \bm\omega)}{\Quant_{j}(1 - p)}
            - 1\right)
  \;\todist\;
  \mathcal N\!\big(\bm\omega^{\top}\Bc,\;
                    \bm\omega^{\top}\Vc\bm\omega\big).
\]
\end{theorem}

The strict condition $\rho_j<0$ is part of the Weissman extrapolation
theorem and is stronger than the $\rho_j\le0$ condition needed for the
marginal Hill CLT.
Condition $(\mathcal H)$ itself is intentionally only the first-order
tail-homoskedasticity condition of \textcite{DaouiaPadoanStupfler2021}.
The recentering by any marginal $\Quant_j(1-p)$ in
\Cref{thm:bg:pool-weissman} is valid because the strict second-order
assumption $\rho_j<0$ supplies the missing rate: by
\textcite[Supplement, Lemma~B.2]{DaouiaPadoanStupfler2021}, for any
$j,\ell$,
\[
  \log\Quant_j(1-p)-\log\Quant_\ell(1-p)
  =
  O\!\big(\abs{A_j(1/p)}\big)
  +
  O\!\big(\abs{A_\ell(1/p)}\big)
  =
  o(k^{-1/2}),
\]
along the Weissman sequence. After multiplication by
$\sqrt{k}/\log(k/(np))$, this deterministic centering difference is
negligible.

The asymptotic distribution of the pooled Weissman estimator is
\emph{identical}, up to the log-rate factor, to that of the pooled
Hill estimator. Two structural consequences follow. First, the
variance- and AMSE-optimal weights of \Cref{prop:bg:optimal-weights}
are also optimal for the pooled extreme-quantile estimator. Second,
the entire information needed for inference at very-extreme quantile
levels is contained in the joint distribution of the per-sample Hill
estimators; the extreme-quantile inputs contribute only through the
deterministic log-scale factor. This structural fact is the analogue,
in the multi-sample setting, of the dominance of the tail-index
estimator at the extreme level that we have already encountered in
\Cref{thm:bg:weissman-expectile-iid,prop:bg:plug-in-clt}.
In the geometric quantile pool, real weights remain algebraically
well-defined because the Weissman estimators are positive with
probability tending to one under the positivity convention of
\Cref{sec:bg:hill}.

\subsection{Tail-homogeneity testing and confidence intervals}

The covariance structure $\Vc$ also yields a Wald-type chi-square
test for the common-$\gamma$ hypothesis $(H_{\gamma})$,
sometimes called \emph{tail homogeneity} across samples.

\begin{theorem}[Tail-homogeneity test;
                \textcite{DaouiaPadoanStupfler2021}, Corollary~3]
\label{thm:bg:tail-homogeneity-test}
Assume the hypotheses of \Cref{thm:bg:pool-hill}, the null
$H_{0}: \gamma_{1} = \cdots = \gamma_{m}$, and the bias-free
calibration $\lambda_j=0$ for every $j$. The statistic
\[
  \Lambda_{n}
  \;=\;
  k\,(\hat{\bm\gamma}_{n} - \bar\mu_{n}\bm 1)^{\top}\hatVc^{-1}
       (\hat{\bm\gamma}_{n} - \bar\mu_{n}\bm 1),
  \qquad
  \bar\mu_{n}
  \;=\;
  \frac{\bm 1^{\top}\hatVc^{-1}\hat{\bm\gamma}_{n}}
       {\bm 1^{\top}\hatVc^{-1}\bm 1},
\]
where $\hatVc$ is a consistent estimator of the covariance matrix of
$\sqrt{k}(\hat{\bm\gamma}_n-\gamma\bm 1)$ under $H_0$.
Then $\Lambda_{n} \todist \chi^{2}_{m - 1}$ under $H_{0}$ and
$\Lambda_{n} \toprob \infty$ under fixed alternatives. A test of
asymptotic size $\alpha$ rejects $H_{0}$ when
$\Lambda_{n} > \chi^{2}_{m - 1,\, 1 - \alpha}$.
\end{theorem}

Paper~3 states the bias-free version above. The same Gaussian GLS
calculation also shows that, if the asymptotic bias is retained rather
than removed by undersmoothing or bias correction, the deviance has the
noncentral limit
\[
  \Lambda_n \todist \chi^2_{m-1}(\delta_\gamma),
  \qquad
  \delta_\gamma
  =
  \Bc^\top\Vc^{-1}\Bc
  -
  \frac{(\bm 1^\top\Vc^{-1}\Bc)^2}{\bm 1^\top\Vc^{-1}\bm 1}.
\]
Thus the central chi-square limit holds whenever
$\Bc\in\mathrm{span}\{\bm 1\}$, in particular under
$\lambda_j=0$ for every $j$. Under contiguous alternatives of order
$k^{-1/2}$, the limit is likewise noncentral rather than divergent.

Under tail homogeneity, \Cref{prop:bg:optimal-weights} furnishes
asymptotic confidence intervals for the common $\gamma$ on the
$\sqrt{k}$ scale; log-scale confidence intervals for the extrapolated
extreme quantile follow from \Cref{thm:bg:pool-weissman}. These
variance-only intervals are centered correctly under undersmoothing or
after valid bias correction; otherwise the asymptotic bias term must be
included in the centering.

\subsection{Distributed inference as a special case}
\label{sec:bg:pool-distrib}

The distributed-inference setting is the special case in which the
$m$ samples are obtained as iid subdivisions of one common underlying
iid sample: data points are iid both within and across samples, with
a common marginal distribution. The framework simplifies considerably
and matches the assumptions of this thesis.

Under distributed-inference assumptions:
\begin{itemize}
\item the common-$\gamma$ condition $(H_{\gamma})$ and the
      tail-homoskedasticity condition $(\mathcal H)$ are trivially
      satisfied;
\item the tail copula $\Rjl$ between any pair of distinct samples
      vanishes identically (independence implies tail independence);
\item $\Vc$ in \Cref{thm:bg:pool-hill} becomes the diagonal matrix
      \eqref{eq:bg:V_c-diagonal};
\item the variance-optimal weights of \Cref{prop:bg:optimal-weights}
      reduce to $\omega_{j}^{\mathrm{Var}} \propto k_{j}$ as in
      \eqref{eq:bg:V-opt-distrib}.
\end{itemize}

\begin{corollary}[Distributed oracle equivalence;
                  \textcite{DaouiaPadoanStupfler2021},
                  Corollary~6, Theorem~3 and Corollary~8]
\label{cor:bg:pool-distrib-oracle}
Under the distributed-inference assumptions of Corollary~5 in
\textcite{DaouiaPadoanStupfler2021}, suppose the sample fractions are
asymptotically equal in the stronger sense
$k_{j} / n_{j} = (k/n)(1 + O(k^{-1/2}))$ for every $j$. Then, under the
conditions of their Theorem~3, the variance-optimally-pooled Hill
estimator
$\hat\gamma_{n}(\bm\omega^{\mathrm{Var}})$ is $\sqrt{k}$-asymptotically
equivalent to the unfeasible benchmark Hill estimator
$\hat\gamma_{n}^{(\mathrm{Hill})}(k)$ built from the entire combined
sample. If, in addition, $\rho<0$ and $p=p(n)$ satisfies
$k/(np)\to\infty$ and $\sqrt{k}/\log(k/(np))\to\infty$, then the
variance-optimally-pooled Weissman estimator
$\hatWei_{n}(1 - p \mid \bm\omega^{\mathrm{Var}})$ is asymptotically
equivalent to the unfeasible benchmark Weissman estimator built from
the combined sample.
\end{corollary}

\Cref{cor:bg:pool-distrib-oracle} is the central efficiency result of
the optimal pooling framework: under the distributed-inference
assumptions, equal sample fractions, and variance-optimal
weighting, the pooled estimators pay no first-order asymptotic
efficiency price relative to the infeasible pooled-sample benchmark.
It is this property that we shall seek to preserve in the
pooled-expectile estimators of \Cref{ch:pool-int,ch:pool-ext}.
