% !TEX root = ../main.tex
\chapter[Pooled intermediate expectiles]{Pooled inference for extreme expectiles at intermediate levels}
\label{ch:pool-int}

\section*{Overview}

This chapter introduces the pooled extreme-expectile estimator at
intermediate levels and develops its asymptotic theory. The construction
plugs the pooled tail-index estimator $\hat\gamma_n(\bm\omega_\gamma)$
and the pooled weighted-geometric Weissman extreme-quantile estimator
$\hatWei_n(\,\cdot\,\mid\bm\omega_q)$ of
\textcite{DaouiaPadoanStupfler2021} into the expectile--quantile
asymptotic equivalence
$\expectile{\tau} \approx \psifn(\gamma)\,\Quant_X(\tau)$. The
delta method applied through $\psifn$ yields a joint Gaussian limit,
from which we read off the variance- and AMSE-optimal pooling weights
tailored to the pooled expectile target.

\section{Setup and notation}
\label{sec:pool-int:setup}

The contribution of this chapter rests on the pooling framework of
\Cref{sec:bg:pooling}, specialised to independent samples with a
common marginal distribution --- the common-distribution
distributed setting of \Cref{sec:bg:pool-distrib}. This is
deliberately stronger than assuming only a common tail index: it gives
one common population target $\expectile{\tau_n}$ and one common
quantile function $\Quant_X$, whereas heterogeneous margins would
require sample-specific targets and additional tail-scale alignment
conditions before pooled expectile inference could be interpreted.
Thus the chapter is intentionally narrower than the full
heterogeneous-tail-copula scope of \textcite{DaouiaPadoanStupfler2021}:
the restriction is what makes the pooled object a single common
expectile rather than a collection of sample-specific targets.

The equal effective sample fraction condition
$k_{j}/n_{j} = (k/n)(1 + o(1))$
is \emph{not} part of the standing setup; it is invoked only in
\Cref{sec:pool-int:weights,sec:pool-int:CI,sec:pool-int:test} where it
simplifies the variance-optimal weights and the limiting bias
structure. When oracle equivalence to the unfeasible pooled-sample
benchmark is cited, we use the stronger
$k_{j}/n_{j} = (k/n)(1 + O(k^{-1/2}))$ condition of
\Cref{cor:bg:pool-distrib-oracle}.

We restate below only those assumptions used throughout. The
pooled-Hill CLT of \Cref{thm:bg:pool-hill} is imported directly; the weighted-geometric
Weissman construction of \Cref{sec:bg:pooling} is kept as the
quantile primitive, while its finite-log intermediate limit is derived
inside this chapter rather than read off from the diverging-log
Weissman CLT \Cref{thm:bg:pool-weissman}.

\paragraph{The $m$ samples and the pooling primitives.}
We have $m$ \emph{independent} samples
$\{X_{i, j}\}_{i = 1, \ldots, n_{j}}$, $j = 1, \ldots, m$, with common
marginal distribution function $F$ and common extreme-value index
$\gamma$. Let $n \coloneqq \sum_{j = 1}^{m} n_j$. All asymptotics are
taken along a triangular array with fixed $m$, $n_j \to \infty$,
$k_j \to \infty$, and $k_j/n_j \to 0$ for every $j$; all CLTs below
are therefore finite-$m$ statements. Independence of the samples
implies asymptotic tail independence, hence the pairwise tail-copula functions in
Condition $\mathcal J(\bm R)$ vanish:
\begin{equation}
\label{eq:pool-int:R-zero}
  \Rjl \;\equiv\; 0
  \qquad\text{for every } j \ne \ell.
\end{equation}
The proportionality conditions of \eqref{eq:bg:pool-rates} are in
force:
\[
  \frac{n_1}{n_j} \to b_j \in (0,\infty),
  \qquad
  \frac{k_1}{k_j} \to c_j \in (0,\infty),
  \qquad j = 1,\ldots,m,
\]
with $b_1 = c_1 = 1$, and the total effective sample size is
$k = \sum_{j = 1}^{m} k_{j}$. Consequently $k_j \asymp k$ uniformly in
$j$ and, with $S \coloneqq \sum_{i = 1}^{m} c_i^{-1}$,
$k/k_j \to c_j S$. Each sample $j$ supplies the two pooling primitives
of \Cref{sec:bg:pooling}: a local Hill estimator
$\hat\gamma_{j}(k_{j})$ as in
\eqref{eq:bg:hill}, and a local Weissman extreme-quantile estimator
$\hat q^{\star}_{j}(\,\cdot\,\mid k_{j})$ as in
\eqref{eq:bg:weissman}. The dot does not represent an additional
communicated function; it denotes the algebraic Weissman rule
recoverable from the communicated Weissman value together with
$(\hat\gamma_{j}(k_{j}), k_j, n_j)$. At any fixed target level this
rule contributes one quantile scalar per sample. These values, together
with their joint asymptotic covariance, constitute the entire
information shared between samples; no raw observation and no
expectile-specific summary statistic crosses between machines.

\paragraph{Heavy-tail regime.}
The common marginal $F$ satisfies the second-order condition
$\mathcal C_{2}(\gamma, \rho, A)$ of \Cref{cond:bg:second-order} with
\begin{equation}
\label{eq:pool-int:moment}
  \gamma \in (0, 1),
  \qquad
  \E X^{-} < \infty,
\end{equation}
the per-sample bias-rate condition
$\sqrt{k_{j}}\, A(n_{j}/k_{j}) \to \lambda_{j} \in \R$ holds for every
$j$, and the moment-negligibility condition
\begin{equation}
\label{eq:pool-int:moment-rate}
  \frac{\sqrt{k}}{\Quant_{X}(\tau_{n})} \;\to\; 0
\end{equation}
is in force. The range $\gamma<1$ and the integrability of $X^{-}$
place the analysis in the finite-mean regime in which expectiles are
well defined and the quantile-based expectile--quantile bridge applies.
The stronger finite-variance restriction $\gamma<1/2$ is needed only
for the direct LAWS estimator reviewed in \Cref{thm:bg:laws-iid}; it is
not an assumption of the pooled QB results below. Condition
\eqref{eq:pool-int:moment-rate} is the multi-sample counterpart of the
$\sqrt{n(1 - \tau_{n})}\,/\Quant_{X}(\tau_{n}) \to 0$ hypothesis of
the QB intermediate theorem and the second-order
expectile--quantile expansion: it ensures that the moment term in the
expectile--quantile expansion \eqref{eq:bg:c-gamma-rho} is
asymptotically negligible, so that the deterministic bias of the
pooled estimator is governed by the single second-order rate carried
by $A$. This is an additional rate restriction linking the target
level and the effective sample size, not an automatic consequence of
heavy-tailedness. Under the equal-fraction scaling
$1 - \tau_n \asymp k/n$, and suppressing slowly varying factors,
$\Quant_X(\tau_n) \asymp (n/k)^\gamma$, so
\eqref{eq:pool-int:moment-rate} requires
$k^{1/2+\gamma}/n^\gamma \to 0$.
The endpoint $\rho=0$ is retained in this intermediate-level chapter:
the expectile--quantile expansion and the QB CLT used below are stated
under the same $\rho\le0$ convention as \Cref{cond:bg:second-order}.
The stricter $\rho<0$ assumption re-enters only for the very-extreme
extrapolation step in \Cref{ch:pool-ext}, through
\Cref{prop:bg:plug-in-clt,thm:bg:pool-weissman}.

\paragraph{The target.}
We target the common population expectile $\expectile{\tau_{n}}$ at
an \emph{intermediate} level $\tau_{n} \to 1$ with
$n(1 - \tau_{n}) \to \infty$, in the regime sense of
\Cref{sec:bg:extreme-expectile}. In addition, the per-sample
log-rate factor
\[
  L_{j}(\tau_{n})
  \;\coloneqq\;
  \log\!\frac{k_{j}}{n_{j}(1 - \tau_{n})}
\]
is assumed to admit a finite limit $L_{j}^{\star} \in \R$ for every
$j$; equivalently, $1 - \tau_{n}$ has the same order as $k_{j}/n_{j}$
for each sample. This condition controls the Hill--Weissman
cross-covariance in \Cref{prop:pool-int:joint-clt} and is the
chapter's link between the chosen target level and the local
effective sample fractions. The very-extreme regime
$n(1 - \tau_{n}^{\prime}) \to c \in [0, \infty)$ is deferred to
\Cref{ch:pool-ext}, where the pooled intermediate-level result proved
here serves as the anchor for Weissman-style extrapolation through
\Cref{prop:bg:plug-in-clt}.

\paragraph{Per-sample QB plug-in: a forced consequence of the pooling protocol.}
Because each sample shares the pooling primitives
$(\hat\gamma_{j}(k_{j}),\,\hat q^{\star}_{j}(\,\cdot\,\mid k_{j}))$
and no expectile-specific summary statistic, the only one of the two
standard intermediate-expectile constructions considered in the thesis
that can be assembled \emph{locally} on each sample, before pooling,
under this communication protocol is the QB plug-in
\eqref{eq:bg:qb} obtained by factoring through the
expectile--quantile equivalence of \Cref{prop:bg:expectile-quantile}:
\[
  \hatexpectile{\tau,\,j}
  \;=\;
  \psifn\!\big(\hat\gamma_{j}(k_{j})\big)\,
  \hat q^{\star}_{j}(\tau \mid k_{j}).
\]
The LAWS direct estimator \eqref{eq:bg:laws} would require sharing
$\tildexpectile{\tau_{n},\,j}$ across samples, because its
asymmetric-least-squares objective is solved as an expectile
estimation problem rather than assembled from the Hill and Weissman
quantile primitives; this lies outside the pooling protocol of
\Cref{sec:bg:pooling}. The comparison of
\Cref{thm:bg:qb-iid,thm:bg:laws-iid} shows that neither LAWS nor QB
uniformly dominates the other on the direct-LAWS range
$\gamma \in (0, 1/2)$; the QB construction is therefore adopted not
because it is uniformly more efficient, but because it is the available
member of this pair under the communicated pooling primitives. The
choice is a derived consequence of the Paper-3 communication protocol,
not a claim that QB is intrinsically optimal among all possible
expectile procedures.

\paragraph{Roadmap.}
\Cref{sec:pool-int:estimator} defines the pooled QB intermediate
expectile estimator from the pooled Hill and pooled Weissman inputs.
\Cref{sec:pool-int:joint-clt} derives the joint $\sqrt{k}$-CLT for
the pooled Hill and log-Weissman inputs, combining the pooled-Hill
limit of \Cref{thm:bg:pool-hill} and the weighted-geometric pooling map
with the local Hill / intermediate-threshold joint CLT, specialised to
the independent-samples case \eqref{eq:pool-int:R-zero}. The output is
the Gaussian input to which the delta method is then applied.
\Cref{sec:pool-int:expectile-clt} applies the delta method through
$\psifn$ and states the main asymptotic-normality theorem of the
chapter. \Cref{sec:pool-int:weights} derives the variance- and
AMSE-optimal pooling weights for the expectile target and compares
them with the corresponding optima for the quantile target.
\Cref{sec:pool-int:CI} provides feasible confidence intervals built
from estimated weights, and \Cref{sec:pool-int:test} adapts the
tail-homogeneity test of \Cref{thm:bg:tail-homogeneity-test} to the
expectile-tail-homogeneity null hypothesis.

\section{The pooled QB expectile estimator}
\label{sec:pool-int:estimator}

The shared primitives $\big(\hat\gamma_{j}(k_{j}),\,
\hat q^{\star}_{j}(\tau_{n} \mid k_{j})\big)_{j = 1, \ldots, m}$
admit two natural \emph{single-weight} constructions of a pooled QB
intermediate-expectile estimator. The first, denoted (A), applies
the QB plug-in once, at the centre, after pooling the Hill and
Weissman primitives separately; the second, denoted (B), computes a
per-sample QB expectile on each machine and then geometric-pools
the $m$ local estimates. At matched single weights --- the only
setting in which the two constructions are directly comparable ---
\Cref{prop:pool-int:A-B-equiv} below shows they share the same
$\sqrt{k}$-asymptotic distribution, so on this common domain the
choice between them is asymptotically inert. We adopt (A) by
parsimony, since it is the direct lift of the iid QB pipeline of
\Cref{thm:bg:qb-iid} --- one tail-index input, one quantile input,
one application of $\psifn$ --- to the pooled setting.

Crucially, (A) exposes an intrinsic \emph{two-weight} structure: the
pooled Hill and pooled Weissman primitives can in principle receive
distinct weight vectors $(\bm\omega_{\gamma}, \bm\omega_{q})$. The
literal local-QB construction (B), understood as first compressing each
sample to one QB scalar and then pooling those scalars with a single
weight vector, does not preserve this separation. One could define a
componentwise two-weight geometric variant by pooling the
$\psifn(\hat\gamma_j)$ factors with $\bm\omega_\gamma$ and the
Weissman factors with $\bm\omega_q$; by the same Taylor argument as
\Cref{prop:pool-int:A-B-equiv}, that variant is again
$\sqrt{k}$-equivalent to (A) for deterministic affine weights. We
therefore adopt (A) because it is algebraically simpler and keeps the
two primitive weight vectors visible.
\Cref{sec:pool-int:weights} will return to the question of whether
this two-weight freedom is ever genuinely exploited, and show that
under the equal-fraction distributed-inference specialisation the
variance-optimal pair lies on the matched-weight diagonal, apart from
the non-identifiability of the Hill weights when $m(\gamma)=0$.

\paragraph{Pooled Hill and pooled Weissman primitives.}
For weight vectors $\bm\omega_{\gamma}, \bm\omega_{q} \in \R^{m}$
satisfying $\bm\omega_{\gamma}^{\top}\bm 1
= \bm\omega_{q}^{\top}\bm 1 = 1$, the linear pooled Hill estimator
\eqref{eq:bg:pooled-hill} and the weighted-geometric pooled
Weissman estimator \eqref{eq:bg:pooled-weissman} read
\begin{align}
\label{eq:pool-int:hill-pool}
  \hat\gamma_{n}(\bm\omega_{\gamma})
    &\;=\;
    \bm\omega_{\gamma}^{\top}\hat{\bm\gamma}_{n}
    \;=\;
    \sum_{j=1}^{m}
      \omega_{\gamma, j}\,\hat\gamma_{j}(k_{j}),
  \\[2pt]
\label{eq:pool-int:weissman-pool}
  \hatWei_{n}(\tau_{n} \mid \bm\omega_{q})
    &\;=\;
    \prod_{j=1}^{m}
      \big[\hat q^{\star}_{j}(\tau_{n} \mid k_{j})\big]^{\omega_{q, j}},
\end{align}
both evaluated at the chapter's intermediate level $\tau_{n}$.
The weights are affine: only the sum-to-one constraint is imposed
here, and the coordinates need not be nonnegative unless a later
special case says so. The geometric products in
\eqref{eq:pool-int:weissman-pool} and below are therefore interpreted
on the log scale and require positive local Weissman factors. Under the
right-heavy-tailed setting with $\gamma > 0$ and $k_j/n_j \to 0$, the
intermediate order statistic $X_{n_j-k_j:n_j,\,j}$ diverges to the
right in probability by regular variation of the tail quantile; hence
$\Prob(\hat q^{\star}_{j}(\tau_{n}\mid k_{j})>0,\ j=1,\ldots,m)\to1$.
All asymptotic log-scale statements below are understood on this
high-probability event, intersected with the event on which the local
and pooled Hill estimators entering $\log\psifn$ lie in $(0,1)$. Arbitrary
definitions on the complementary event do not affect the weak limits.

\paragraph{The two candidate estimators.}
The first construction, which we adopt as canonical, pools the
primitives separately and applies the QB plug-in once at the
centre,
\begin{equation}
\label{eq:pool-int:expectile-A}
  \hatexpectile{\tau_{n}}^{(A)}\!
    (\bm\omega_{\gamma}, \bm\omega_{q})
  \;\coloneqq\;
  \psifn\!\big(\hat\gamma_{n}(\bm\omega_{\gamma})\big)\,
  \hatWei_{n}(\tau_{n} \mid \bm\omega_{q}).
\end{equation}
The second computes a per-sample QB expectile on each machine and
geometric-pools the resulting $m$ local estimates,
\begin{equation}
\label{eq:pool-int:expectile-B}
  \hatexpectile{\tau_{n}}^{(B)}\!(\bm\omega)
  \;\coloneqq\;
  \prod_{j = 1}^{m}\!\big[
    \psifn(\hat\gamma_{j}(k_{j}))\,
    \hat q^{\star}_{j}(\tau_{n} \mid k_{j})
  \big]^{\omega_{j}}
  \;=\;
  \Big(\prod_{j = 1}^{m}\psifn(\hat\gamma_{j}(k_{j}))^{\omega_{j}}\Big)
  \cdot \hatWei_{n}(\tau_{n} \mid \bm\omega).
\end{equation}
When construction (A) is restricted to matched weights
$(\bm\omega_{\gamma},\bm\omega_q) = (\bm\omega,\bm\omega)$, the
Weissman factor is identical in the two formulae; the constructions
differ only in whether the nonlinearity $\psifn$ acts once on the
pooled $\hat\gamma_{n}$ (construction (A)) or componentwise on the
$m$ per-sample $\hat\gamma_{j}$'s followed by weighted-geometric
averaging (construction (B)). The iid QB
plug-in of \Cref{thm:bg:qb-iid} ($m = 1$) reduces both formulae
to $\psifn(\hat\gamma)\,\hat q^{\star}$; construction (A) is the
direct $m \ge 2$ generalisation of that structure, whereas (B)
introduces $m$ separate applications of $\psifn$ that have no
counterpart in the iid baseline.

\begin{proposition}[$\sqrt{k}$-asymptotic equivalence of (A) and (B)]
\label{prop:pool-int:A-B-equiv}
Under the assumptions of \Cref{sec:pool-int:setup}, for fixed
$m \in \mathbb{N}$ and every deterministic weight vector
$\bm\omega \in \R^{m}$ with $\bm\omega^{\top}\bm 1 = 1$,
\[
  \sqrt{k}\,
  \log\!\frac{\hatexpectile{\tau_{n}}^{(A)}(\bm\omega, \bm\omega)}
              {\hatexpectile{\tau_{n}}^{(B)}(\bm\omega)}
  \;\toprob\; 0.
\]
In fact the unscaled log-ratio is itself $O_{p}(k^{-1})$; multiplying
by $\sqrt{k}$ therefore gives $O_p(k^{-1/2}) = o_p(1)$, so the
discrepancy is negligible at the CLT scale. Consequently, when
construction (A) is evaluated at matched weights
$(\bm\omega_{\gamma}, \bm\omega_{q}) = (\bm\omega, \bm\omega)$, the
two estimators share the same $\sqrt{k}$-asymptotic distribution and
\Cref{thm:pool-int:main} applies to either without modification. The
same conclusion extends to estimated weights $\hat{\bm\omega}$
satisfying $\hat{\bm\omega}^{\top}\bm 1 = 1$ exactly and
$\hat{\bm\omega} \toprob \bm\omega$. If exact normalisation is
relaxed to
$\sqrt{k}\abs{\hat{\bm\omega}^{\top}\bm 1 - 1}\toprob 0$, the
$\sqrt{k}$-equivalence still holds, although the unscaled
$O_p(k^{-1})$ refinement need not.
\end{proposition}

\begin{proof}[Sketch.]
A first-order Taylor expansion of $\log\psifn$ around $\gamma$
gives $\log\psifn(\hat\gamma_{j}) - \log\psifn(\gamma) = m(\gamma)
(\hat\gamma_{j} - \gamma) + O_{p}((\hat\gamma_{j} - \gamma)^{2})$,
and analogously for $\hat\gamma_{n}(\bm\omega)$. The linear terms
in $\log\hatexpectile{}^{(A)}$ and $\log\hatexpectile{}^{(B)}$
coincide because
$\sum_{j} \omega_{j}(\hat\gamma_{j} - \gamma) = \hat\gamma_{n}
(\bm\omega) - \gamma$, so the discrepancy is the difference of
quadratic-and-higher remainders, which is $O_{p}(1/k)$. On the
$\sqrt{k}$-scale this is $O_{p}(k^{-1/2}) = o_{p}(1)$. Full proof
in \Cref{app:proofs:A-B-equiv}.
\end{proof}

\paragraph{Canonical choice.}
We adopt
\begin{equation}
\label{eq:pool-int:expectile-pool}
  \hatexpectile{\tau_{n}}^{\mathrm{pool}}\!
    (\bm\omega_{\gamma}, \bm\omega_{q})
  \;\coloneqq\;
  \psifn\!\big(\hat\gamma_{n}(\bm\omega_{\gamma})\big)\,
  \hatWei_{n}(\tau_{n} \mid \bm\omega_{q}),
\end{equation}
identical to \eqref{eq:pool-int:expectile-A}, as the canonical
pooled QB intermediate-expectile estimator for the remainder of
this chapter. The choice is justified by parsimony:
\eqref{eq:pool-int:expectile-A} reproduces the iid QB pipeline of
\Cref{thm:bg:qb-iid} --- one tail-index input, one quantile input,
one application of $\psifn$ --- directly at the pooled level,
whereas \eqref{eq:pool-int:expectile-B} introduces $m$ applications
of $\psifn$ which, by \Cref{prop:pool-int:A-B-equiv}, contribute
no additional asymptotic information.\footnote{At the second order,
the two constructions differ through a Jensen-type term of order
$k^{-1}$ involving $m'(\gamma)$ and the dispersion of the local Hill
estimators across samples; its sign depends on the weight vector
and need not favour either construction in general. This term sits
below the order of the $\sqrt{k}$-asymptotic theory developed here
and below the standard EVI bias control
$\sqrt{k_{j}}\,A(n_{j}/k_{j}) \to \lambda_{j}$ of
\Cref{thm:bg:hill}, so we do not use it.}

\paragraph{Two weight vectors, not one.}
Definition \eqref{eq:pool-int:expectile-pool} carries two free weight
vectors rather than a single one. In the quantile-only analysis of
\eqref{eq:bg:pooled-weissman}, the weighted-geometric quantile pool is
formed with a single $\bm\omega$, and the extreme-quantile limit of
\Cref{thm:bg:pool-weissman} is governed by that same vector. The
variance- and AMSE-optima of \Cref{prop:bg:optimal-weights} therefore
apply to the quantile target through one set of weights. The expectile
target combines the Hill and Weissman primitives
through the nonlinear map $\psifn$, and the asymptotic variance of
$\hatexpectile{\tau_{n}}^{\mathrm{pool}}$ acquires a contribution
from the within-sample cross-covariance of $\hat\gamma_{j}$ and
$\hat q^{\star}_{j}$ that is absent in the quantile-only analysis.
This contribution can in principle pull the variance-optimal
$\bm\omega_{\gamma}$ apart from the variance-optimal
$\bm\omega_{q}$; \Cref{sec:pool-int:weights} identifies the exact
discrepancy and the sufficient conditions under which the two
optima coincide. Keeping the weights separate in the definition
leaves that section free to specify their relationship rather than
imposing it by hand.

\paragraph{Why the pooled Weissman at the intermediate level.}
At the intermediate level $\tau_{n}$, the per-sample Weissman
estimator $\hat q^{\star}_{j}(\tau_{n} \mid k_{j}) = (k_{j}/(n_{j}
(1 - \tau_{n})))^{\hat\gamma_{j}(k_{j})}\, X_{n_{j} - k_{j} : n_{j},\,j}$
is the device that moves sample $j$ from its own local intermediate
level to the common chapter target. When $\tau_{n}$ matches the sample-specific
intermediate level $\tau_{n_{j}} \coloneqq 1 - k_{j}/n_{j}$, the
multiplicative power factor collapses to one and
$\hat q^{\star}_{j}$ reduces to the bare $(k_{j} + 1)$-th upper order
statistic of sample $j$ --- the empirical intermediate quantile.
Otherwise, $\hat q^{\star}_{j}$ interpolates or mildly extrapolates
across intermediate levels, and remains $\sqrt{k_{j}}$-consistent by
the within-sample joint CLT for the Hill estimator and the
intermediate log-order statistic invoked in
\Cref{prop:pool-int:joint-clt} below --- this is the finite-log-rate
analogue of the diverging-log-rate Weissman CLT
\Cref{thm:bg:weissman}, not a direct consequence of it. This bridge
across local intermediate levels is why we use the Weissman primitive
uniformly, instead of switching to bare order statistics at
intermediate levels: it is already the primitive shared by the pooling
protocol of \Cref{sec:bg:pooling}, and it carries the same form to the
extreme regime treated in
\Cref{ch:pool-ext}, keeping the two chapters on a single notational
footing.

\section{Joint asymptotic distribution of the pooled inputs}
\label{sec:pool-int:joint-clt}

This section establishes the joint $\sqrt{k}$-Gaussian limit of
$(\hat\gamma_{n}(\bm\omega_{\gamma}),\,
\log\hatWei_{n}(\tau_{n} \mid \bm\omega_{q}))$ at intermediate
level $\tau_{n}$. The result is the input that
\Cref{sec:pool-int:expectile-clt} feeds through the delta method to
obtain the asymptotic distribution of the pooled QB expectile
estimator \eqref{eq:pool-int:expectile-pool}.

\paragraph{Within-sample decomposition.}
For each sample $j$, the log of the Weissman estimator decomposes
exactly as
\begin{equation}
\label{eq:pool-int:weissman-decomp}
  \log\hat q^{\star}_{j}(\tau_{n} \mid k_{j})
  \;=\;
  \log X_{n_{j} - k_{j} : n_{j},\,j}
  \;+\;
  L_{j}(\tau_{n})\, \hat\gamma_{j}(k_{j}),
  \qquad
  L_{j}(\tau_{n}) \coloneqq
    \log\!\frac{k_{j}}{n_{j}(1 - \tau_{n})}.
\end{equation}
The log-rate factor $L_{j}(\tau_{n})$ records the extent to which
sample $j$ extrapolates from its own intermediate level
$\tau_{n_{j}} \coloneqq 1 - k_{j}/n_{j}$ to the chapter target
$\tau_{n}$: $L_{j} = 0$ when the two coincide, $L_{j} > 0$ for
extrapolation, $L_{j} < 0$ for interpolation. The only stochastic input
to the proof of \Cref{prop:pool-int:joint-clt} is the
within-sample joint asymptotic behaviour of the Hill estimator and
its companion intermediate log-order statistic at effective sample
size $k_{j}$. By the tail-empirical-process representation
\parencite[Theorem~2.4.8 and \S5.1]{deHaanFerreira2006}, the two
statistics are jointly Gaussian on the $\sqrt{k_{j}}$-scale,
\emph{asymptotically independent}, and each has asymptotic variance
$\gamma^{2}$. The zero off-diagonal entry is structural: the Hill
estimator is a smooth functional of the centred log-spacings above
$X_{n_{j} - k_{j} : n_{j},\,j}$, while the intermediate-threshold
fluctuation is the centred $(k_{j}+1)$-th upper log-order statistic
itself, and the two decouple in the tail-empirical-process limit.
This same joint behaviour underlies the iid QB plug-in CLT of
\textcite{DaouiaGirardStupfler2020} reproduced in
\Cref{thm:bg:qb-iid}. Combining it with
\eqref{eq:pool-int:weissman-decomp}, while using the second-order
tail-quantile expansion for the deterministic centering, yields the
within-sample joint central limit theorem
\begin{equation}
\label{eq:pool-int:per-sample-clt}
  \sqrt{k_{j}}\!
    \begin{pmatrix}
      \hat\gamma_{j}(k_{j}) - \gamma \\[2pt]
      \log\hat q^{\star}_{j}(\tau_{n} \mid k_{j})
        - \log \Quant_{X}(\tau_{n})
    \end{pmatrix}
  \;\todist\;
  \mathcal N\!\left(\bm\mu_{j}^{\star},\,
                    \gamma^{2}\!\begin{pmatrix}
                      1 & L_{j}^{\star} \\
                      L_{j}^{\star} & 1 + (L_{j}^{\star})^{2}
                    \end{pmatrix}\right),
\end{equation}
under the common-marginal assumption of \Cref{sec:pool-int:setup}
and the condition $L_{j}(\tau_{n}) \to L_{j}^{\star} \in \R$. With
$\Psi_\rho$ the second-order kernel defined in
\eqref{eq:pool-int:Bcstar-def}, the limiting mean is
\[
  \bm\mu_j^\star
  =
  \begin{pmatrix}
    \lambda_j/(1-\rho) \\[2pt]
    \lambda_j\!\left\{L_j^\star/(1-\rho)
      - \Psi_\rho(e^{L_j^\star})\right\}
  \end{pmatrix}.
\]
The first component is the usual Hill bias, while the second combines
that Hill bias, transported through the log-rate $L_j^\star$, with the
deterministic second-order Weissman remainder. The $2 \times 2$ covariance reveals
the within-sample Hill--Weissman correlation induced by the log-rate
factor: cross-covariance $\gamma^{2} L_{j}^{\star}$, and a Weissman
variance $\gamma^{2}(1 + (L_{j}^{\star})^{2})$ obtained from the
asymptotically independent order-statistic and rescaled-Hill
contributions.

\paragraph{Pooling under independence.}
Independence of the $m$ samples implies that the cross-sample
covariance blocks of the $2m$-vector
$\big(\hat\gamma_{j} - \gamma,\, \log\hat q^{\star}_{j} - \log
\Quant_{X}(\tau_{n})\big)_{j = 1, \ldots, m}$ vanish identically.
Linear pooling with weight vectors
$(\bm\omega_{\gamma}, \bm\omega_{q})$ then collapses this
$2m$-vector onto a $2$-vector whose covariance is assembled
block-by-block from the within-sample blocks of
\eqref{eq:pool-int:per-sample-clt}, each rescaled to the common
$\sqrt{k}$-scale via the proportionality conditions
\eqref{eq:bg:pool-rates}.

\begin{proposition}[Joint pooled CLT at intermediate level]
\label{prop:pool-int:joint-clt}
Under the assumptions of \Cref{sec:pool-int:setup} and the
convergence $L_{j}(\tau_{n}) \to L_{j}^{\star} \in \R$ for every $j$,
for any weight vectors $\bm\omega_{\gamma}, \bm\omega_{q} \in \R^{m}$
satisfying
$\bm\omega_{\gamma}^{\top}\bm 1 = \bm\omega_{q}^{\top}\bm 1 = 1$,
\begin{equation}
\label{eq:pool-int:joint-pooled-clt}
  \sqrt{k}\!
    \begin{pmatrix}
      \hat\gamma_{n}(\bm\omega_{\gamma}) - \gamma \\[2pt]
      \log\hatWei_{n}(\tau_{n} \mid \bm\omega_{q})
        - \log \Quant_{X}(\tau_{n})
    \end{pmatrix}
  \;\todist\;
  \mathcal N\!\big(\bm\mu(\bm\omega_{\gamma},\bm\omega_{q}),\,
                    \bm\Sigma(\bm\omega_{\gamma},\bm\omega_{q})\big),
\end{equation}
with covariance
\begin{equation}
\label{eq:pool-int:Sigma}
  \begin{gathered}
  \bm D^{\star} \coloneqq \mathrm{diag}(\bm L^{\star}),\\[-2pt]
  \bm\Sigma(\bm\omega_{\gamma},\bm\omega_{q})
  \;=\;
  \begin{pmatrix}
    \bm\omega_{\gamma}^{\top}\Vc\bm\omega_{\gamma}
      &
    \bm\omega_{\gamma}^{\top}\bm D^{\star}\Vc\,\bm\omega_{q} \\[2pt]
    \bm\omega_{q}^{\top}\Vc\bm D^{\star}\bm\omega_{\gamma}
      &
    \bm\omega_{q}^{\top}\!\big(\Vc + \bm D^{\star}\Vc\bm D^{\star}
      \big)\,\bm\omega_{q}
  \end{pmatrix}.
  \end{gathered}
\end{equation}
Here $\Vc$ is the diagonal matrix \eqref{eq:bg:V_c-diagonal}, so
$\bm D^{\star}\Vc\bm D^{\star}=(\bm D^{\star})^{2}\Vc$, and
$\bm L^{\star} = (L_{1}^{\star}, \ldots, L_{m}^{\star})^{\top}$. The
mean
\begin{equation}
\label{eq:pool-int:mu-pool}
  \bm\mu(\bm\omega_{\gamma}, \bm\omega_{q})
  \;=\;
  \begin{pmatrix}
    \bm\omega_{\gamma}^{\top}\Bc \\[2pt]
    \bm\omega_{q}^{\top}\Bc^{\star}
  \end{pmatrix},
\end{equation}
where $\Bc$ is the pooled-Hill bias vector of \Cref{thm:bg:pool-hill}
and the limiting pooled-Weissman bias vector at the intermediate
level has entries
\begin{equation}
\label{eq:pool-int:Bcstar-def}
  \big(\Bc^{\star}\big)_{j}
  \;=\;
  \sqrt{c_{j}\textstyle\sum_{i = 1}^{m} c_{i}^{-1}}\;
  \lambda_{j}\Big[\frac{L_{j}^{\star}}{1 - \rho}
       \;-\; \Psi_{\rho}\big(e^{L_{j}^{\star}}\big)\Big],
  \qquad
  \Psi_{\rho}(y) \;=\;
  \begin{cases}
    (y^{\rho} - 1)/\rho, & \rho < 0, \\[2pt]
    \log y, & \rho = 0,
  \end{cases}
\end{equation}
combining the Hill bias $b_{j}^{H} = \lambda_{j}/(1-\rho)$, weighted by
the log-rate $L_{j}^{\star}$, with the limiting second-order Weissman
remainder $-\lambda_{j}\Psi_{\rho}(e^{L_{j}^{\star}})$. The
intermediate log-order statistic contributes no bias term of its own:
centred at $\log\Quant_{X}(\tau_{n_{j}})$ it has zero asymptotic bias on
the $\sqrt{k_{j}}$ scale
\parencite[Theorem~2.4.8]{deHaanFerreira2006}, as established
in \Cref{app:proofs:joint-clt}. The covariance
\eqref{eq:pool-int:Sigma} does not depend on $\Bc^{\star}$.
The proposition is stated for deterministic affine weights. The same
limit holds for estimated weights satisfying
$\hat{\bm\omega}_{\gamma}^{\top}\bm 1 =
\hat{\bm\omega}_{q}^{\top}\bm 1 = 1$ exactly and
$\hat{\bm\omega}_{\gamma} \toprob \bm\omega_{\gamma}$,
$\hat{\bm\omega}_{q} \toprob \bm\omega_{q}$; this estimated-weight lift
is deferred to \Cref{sec:pool-int:CI}.
\end{proposition}

\begin{proof}
See \Cref{app:proofs:joint-clt}.
\end{proof}

\Cref{prop:pool-int:joint-clt} is the joint counterpart, at the
intermediate finite-log-rate level $L_{j}(\tau_{n}) \to L_{j}^{\star}
\in \R$, of the pooled-Hill limit and weighted-geometric Weissman
construction recalled in \Cref{sec:bg:pooling}. The extreme-level,
diverging-log-rate analogue is the subject of
\Cref{ch:pool-ext}. The
new feature for the expectile target is the cross-covariance
$\bm\omega_{\gamma}^{\top}\mathrm{diag}(\bm L^{\star})\Vc\bm\omega_{q}$:
it is the structural reason why the variance-optimal weights
$(\bm\omega_{\gamma}, \bm\omega_{q})$ for the pooled expectile
estimator need not coincide, an observation that
\Cref{sec:pool-int:weights} develops quantitatively.

\section{Asymptotic normality of the pooled expectile estimator}
\label{sec:pool-int:expectile-clt}

The main result of this chapter is the central limit theorem for the
pooled QB intermediate expectile estimator
\eqref{eq:pool-int:expectile-pool}. The result is obtained by feeding
the joint pooled CLT of \Cref{prop:pool-int:joint-clt} through the
smooth log-linear map $(g, x) \mapsto \log\psifn(g) + x$ via a
first-order Taylor expansion, combined with the second-order
expectile--quantile expansion that controls the deterministic gap
between $\log\expectile{\tau_{n}}$ and $\log\psifn(\gamma) +
\log\Quant_{X}(\tau_{n})$.

\paragraph{Derivative of $\log \psifn$.}
A short calculation gives
\begin{equation}
\label{eq:pool-int:m-function}
  \frac{d}{dg} \log \psifn(g)
  \;=\;
  \frac{1}{1 - g} - \log(g^{-1} - 1)
  \;=:\;
  m(g),
\end{equation}
which is the same $m(\gamma)$ that governs the variance of the QB
intermediate-level estimator in the iid baseline
\Cref{thm:bg:qb-iid}. The function $m(\,\cdot\,)$ is continuous on
$(0, 1)$, vanishes at the value $\gamma^{\star} \approx 0.2178$
solving $(1 - \gamma)^{-1} = \log(\gamma^{-1} - 1)$, and is positive
(resp.\ negative) for $\gamma$ above (resp.\ below) $\gamma^{\star}$.

\paragraph{The deterministic expectile--quantile gap.}
The first-order asymptotic equivalence of
\Cref{prop:bg:expectile-quantile},
\[
  \expectile{\tau} / \Quant_{X}(\tau) \to \psifn(\gamma),
  \qquad \tau \uparrow 1,
\]
sharpens, under \Cref{cond:bg:second-order}, through the second-order
expansion \eqref{eq:bg:c-gamma-rho}
\parencite[Proposition~1(i)]{DaouiaGirardStupfler2020},
\begin{equation}
\label{eq:pool-int:xi-Q-gap}
  \begin{aligned}
  \log\expectile{\tau} - \log\psifn(\gamma) - \log\Quant_{X}(\tau)
  &=
  c(\gamma, \rho)\, A\!\big((1 - \tau)^{-1}\big)
  + o\!\big(A((1 - \tau)^{-1})\big)
  \\
  &\quad
  + O\!\big(\Quant_X(\tau)^{-1}\big),
  \qquad \tau \uparrow 1.
  \end{aligned}
\end{equation}
Here the constant $c(\gamma, \rho)$ is given explicitly in
\eqref{eq:bg:c-gamma-rho}. The final term in
\eqref{eq:pool-int:xi-Q-gap} is the first-moment contribution in the
DGS expansion; it is not, in general, an $o(A)$ term. What is needed
for the CLT is only its sequence-level disappearance after
multiplication by $\sqrt{k}$. Along the chapter target $\tau_n$,
\eqref{eq:pool-int:moment-rate} gives
$\sqrt{k}/\Quant_X(\tau_n) \to 0$, while the bias-rate assumptions
below give $\sqrt{k}\,A((1-\tau_n)^{-1}) = O(1)$. Hence
\begin{equation}
\label{eq:pool-int:xi-Q-gap-seq}
  \sqrt{k}\Big[
    \log\expectile{\tau_n}
    - \log\psifn(\gamma)
    - \log\Quant_X(\tau_n)
  \Big]
  =
  c(\gamma,\rho)\sqrt{k}\,A\!\big((1-\tau_n)^{-1}\big)
  + o(1).
\end{equation}
The chapter setup of \Cref{sec:pool-int:setup} already \emph{implies}
a finite limit
\begin{equation}
\label{eq:pool-int:Lambda-bullet}
  \Lambda^{\bullet}
  \;\coloneqq\;
  \lim_{n \to \infty}\sqrt{k}\,A\!\big((1 - \tau_{n})^{-1}\big)
  \;\in\; \R,
\end{equation}
without any additional hypothesis: starting from the per-sample bias
rate $\sqrt{k_{j}}\,A(n_{j}/k_{j}) \to \lambda_{j}$, the log-rate
convergence $L_{j}(\tau_{n}) \to L_{j}^{\star}$, and the signed ratio
convergence for the constant-sign auxiliary function $A$ under
\Cref{cond:bg:second-order}, one obtains
$\sqrt{k_{j}}\,A((1 - \tau_{n})^{-1}) \to \lambda_{j}\,
(e^{L_{j}^{\star}})^{\rho}$, and a further rescaling by
$\sqrt{k/k_{j}} \to \sqrt{c_{j}\sum_{i} c_{i}^{-1}}$ delivers
\eqref{eq:pool-int:Lambda-bullet}. Equivalently, with
$S=\sum_i c_i^{-1}$, the constants already introduced obey the
compatibility identity
\[
  \sqrt{c_jS}\,\lambda_j e^{\rho L_j^\star}
  =
  \sqrt{c_\ell S}\,\lambda_\ell e^{\rho L_\ell^\star}
  \qquad \text{for all } j,\ell,
\]
because both sides are limits of the same sequence
$\sqrt{k}\,A((1-\tau_n)^{-1})$; see \Cref{app:proofs:main-clt} for the
explicit verification.
The finite asymptotic constant
\begin{equation}
\label{eq:pool-int:b-xi-Q-def}
  b^{\xi/\Quant}
  \;\coloneqq\;
  \lim_{n \to \infty}\sqrt{k}\,
  \big[\log\psifn(\gamma) + \log\Quant_{X}(\tau_{n})
       - \log\expectile{\tau_{n}}\big]
  \;=\;
  -\,c(\gamma, \rho)\,\Lambda^{\bullet}
\end{equation}
follows. The sign convention in \eqref{eq:pool-int:b-xi-Q-def} is
chosen so that $b^{\xi/\Quant}$ enters the asymptotic mean of the
pooled estimator with a $+$ sign in \eqref{eq:pool-int:mu-xi} below;
that is, $b^{\xi/\Quant}$ is the deterministic \emph{bias
contribution} of the gap, not the gap itself. The contribution
$b^{\xi/\Quant}$ is \emph{absent} from the analogue quantile-target
pooling theorem of \textcite{DaouiaPadoanStupfler2021}: it is the deterministic
correction that arises when the quantile target is replaced by the
expectile target through the QB plug-in.

\begin{theorem}[Pooled QB intermediate expectile CLT]
\label{thm:pool-int:main}
Under the setup and heavy-tail regime of \Cref{sec:pool-int:setup}
(including \Cref{eq:pool-int:moment,eq:pool-int:moment-rate}), the
positivity/log-domain convention of \Cref{sec:pool-int:estimator}, the
assumptions of \Cref{prop:pool-int:joint-clt}, and the finite
expectile--quantile bias limit \eqref{eq:pool-int:Lambda-bullet} implied
by these conditions, for any weight vectors
$\bm\omega_{\gamma}, \bm\omega_{q} \in \R^{m}$ satisfying
$\bm\omega_{\gamma}^{\top}\bm 1 = \bm\omega_{q}^{\top}\bm 1 = 1$,
\begin{equation}
\label{eq:pool-int:main-clt}
  \sqrt{k}\!\left(
    \frac{\hatexpectile{\tau_{n}}^{\mathrm{pool}}\!
          (\bm\omega_{\gamma}, \bm\omega_{q})}
         {\expectile{\tau_{n}}}
    - 1\right)
  \;\todist\;
  \mathcal N\!\big(\mu^{\xi}(\bm\omega_{\gamma},\bm\omega_{q}),\,
                    \sigma^{\xi}(\bm\omega_{\gamma},\bm\omega_{q})^{2}\big),
\end{equation}
with asymptotic variance
\begin{equation}
\label{eq:pool-int:sigma-xi}
  \sigma^{\xi}(\bm\omega_{\gamma},\bm\omega_{q})^{2}
  \;=\;
  m(\gamma)^{2}\,\bm\omega_{\gamma}^{\top}\Vc\bm\omega_{\gamma}
  \;+\;
  2\, m(\gamma)\,
    \bm\omega_{\gamma}^{\top}\mathrm{diag}(\bm L^{\star})\Vc\bm\omega_{q}
  \;+\;
  \bm\omega_{q}^{\top}\!\big(\Vc + \mathrm{diag}(\bm L^{\star})^{2}
    \Vc\big)\bm\omega_{q},
\end{equation}
and asymptotic bias
\begin{equation}
\label{eq:pool-int:mu-xi}
  \mu^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})
  \;=\;
  m(\gamma)\,\bm\omega_{\gamma}^{\top}\Bc
  \;+\;
  \bm\omega_{q}^{\top}\Bc^{\star}
  \;+\;
  b^{\xi/\Quant},
\end{equation}
where $\Bc$ and $\Bc^{\star}$ are as in
\Cref{prop:pool-int:joint-clt} (the pooled-Hill and limiting
intermediate-level pooled-Weissman bias vectors) and $b^{\xi/\Quant}$
is the asymptotic expectile--quantile bias contribution of
\eqref{eq:pool-int:b-xi-Q-def}.
\end{theorem}

\begin{proof}
See \Cref{app:proofs:main-clt}.
\end{proof}

Writing $\bm D^{\star}=\mathrm{diag}(\bm L^\star)$, the variance in
\eqref{eq:pool-int:sigma-xi} factorises as
\[
  \sigma^{\xi}(\bm\omega_{\gamma},\bm\omega_{q})^{2}
  =
  \big(m(\gamma)\bm\omega_{\gamma}+\bm D^{\star}\bm\omega_q\big)^{\top}
  \Vc\big(m(\gamma)\bm\omega_{\gamma}+\bm D^{\star}\bm\omega_q\big)
  + \bm\omega_q^{\top}\Vc\bm\omega_q.
\]
This form makes the non-negativity immediate and shows how the log-rate
vector $\bm L^\star$ couples the Hill and Weissman inputs. The
weight-optimisation section exploits exactly this factorisation.

\paragraph{Reduction to the iid baseline.}
\Cref{thm:pool-int:main} contains \Cref{thm:bg:qb-iid} as the
special case $m = 1$ with the chapter target aligned to the single
sample's intermediate level. The weight vectors collapse to scalars
$\omega_{\gamma} = \omega_{q} = 1$; the alignment forces
$L_{1}^{\star} = 0$, $\Psi_{\rho}(e^{L_{1}^{\star}}) = 0$, and
$\Vc = \gamma^{2}$. The asymptotic variance \eqref{eq:pool-int:sigma-xi}
collapses to $\gamma^{2}(1 + m(\gamma)^{2})$, in exact agreement with
\Cref{thm:bg:qb-iid}. The bias \eqref{eq:pool-int:mu-xi} collapses
likewise. The pooled-Weissman bias vanishes, $\Bc^{\star} = 0$: at
$L_{1}^{\star} = 0$ the bracket in \eqref{eq:pool-int:Bcstar-def}
reduces to $L_{1}^{\star}/(1 - \rho) - \Psi_{\rho}(1) = 0$, since the
intermediate order statistic has zero asymptotic bias on the
$\sqrt{k}$ scale. Hence
\begin{equation}
\label{eq:pool-int:iid-bias}
  \mu^{\xi}
  \;=\;
  m(\gamma)\,\Bc + \Bc^{\star} + b^{\xi/\Quant}
  \;=\;
  \frac{m(\gamma)}{1 - \rho}\,\lambda \;-\; c(\gamma, \rho)\,\lambda
  \;=\;
  \Big(\frac{m(\gamma)}{1 - \rho} - c(\gamma, \rho)\Big)\lambda,
\end{equation}
the Hill bias $\lambda/(1 - \rho)$ propagated through $m(\gamma)$,
minus the deterministic expectile--quantile gap $c(\gamma, \rho)\lambda$
of \eqref{eq:bg:c-gamma-rho}. This is precisely the bias of
\Cref{thm:bg:qb-iid}: there is no cancellation between an
order-statistic bias and the gap, because --- as
\Cref{app:proofs:joint-clt} establishes --- the intermediate order
statistic has zero asymptotic bias on the $\sqrt{k}$ scale and
contributes no bias term at all.

\paragraph{Equal-fraction distributed simplification.}
Under the equal-effective-fraction distributed specialisation,
$k_{j}/n_{j} = (k/n)(1 + o(1))$ for every $j$, the log-rate factor
$L_{j}(\tau_{n}) = \log(k/(n(1 - \tau_{n}))) + o(1)$ becomes a
sample-free quantity. Writing $L^{\star}$ for its common limit,
\eqref{eq:pool-int:sigma-xi} simplifies to
\begin{equation}
\label{eq:pool-int:sigma-xi-distrib}
  \sigma^{\xi}(\bm\omega_{\gamma},\bm\omega_{q})^{2}
  \;=\;
  m(\gamma)^{2}\,\bm\omega_{\gamma}^{\top}\Vc\bm\omega_{\gamma}
  \;+\;
  2\,m(\gamma)\, L^{\star}\,
    \bm\omega_{\gamma}^{\top}\Vc\bm\omega_{q}
  \;+\;
  \big(1 + (L^{\star})^{2}\big)\,
    \bm\omega_{q}^{\top}\Vc\bm\omega_{q}.
\end{equation}
With matched weights $\bm\omega_{\gamma} = \bm\omega_{q} = \bm\omega$
and the chapter target aligned to the common per-sample intermediate
level ($L^{\star} = 0$), \eqref{eq:pool-int:sigma-xi-distrib}
collapses to $(1 + m(\gamma)^{2})\,\bm\omega^{\top}\Vc\bm\omega$ ---
the natural pooled analogue of the iid intermediate QB variance
$\gamma^{2}(1 + m(\gamma)^{2})$ of \Cref{thm:bg:qb-iid}.

\paragraph{Why two weight vectors.}
The separate weight pair $(\bm\omega_{\gamma}, \bm\omega_{q})$ is
needed to expose the full variance structure of the pooled QB
expectile estimator: the cross term in \eqref{eq:pool-int:sigma-xi}
couples the two pooled inputs through $\mathrm{diag}(\bm L^{\star})$
and is a genuine structural feature of the expectile target. As
\Cref{sec:pool-int:weights} shows, however, the additional freedom is
not exploited by the variance optimum in the equal-fraction
distributed case: under the additional equal-fraction
specialisation of \Cref{sec:pool-int:setup}, the variance-optimal
quantile weights remain
$\bm\omega^{\xi,\mathrm{Var}}_{q}=\bm\omega^{\mathrm{Var}}$
(\Cref{prop:pool-int:var-opt-distrib}). When $m(\gamma)\ne0$ the unique
variance-optimal pair is matched,
$\bm\omega^{\xi,\mathrm{Var}}_{\gamma}
= \bm\omega^{\xi,\mathrm{Var}}_{q} = \bm\omega^{\mathrm{Var}}$; when
$m(\gamma)=0$, the $\gamma$-weight is variance-unidentified and this
matched vector is only a canonical representative.

\section{Variance- and AMSE-optimal weights}
\label{sec:pool-int:weights}

We now identify the weight vectors $(\bm\omega_{\gamma},
\bm\omega_{q})$ that minimise the asymptotic variance
\eqref{eq:pool-int:sigma-xi} and the AMSE
$(\mu^{\xi})^{2} + \sigma^{\xi, 2}$ of the pooled QB expectile
estimator, subject to the affine sum-to-one constraints
$\bm\omega_{\gamma}^{\top}\bm 1 = \bm\omega_{q}^{\top}\bm 1 = 1$.
The qualitative question driving this section is whether the
expectile-optimal weights coincide with those for the quantile target
of \Cref{prop:bg:optimal-weights}, or whether the expectile target
requires its own weight scheme.

\subsection*{Variance-optimal weights}

The variance \eqref{eq:pool-int:sigma-xi} admits the identity
\begin{equation}
\label{eq:pool-int:variance-factorisation}
  \sigma^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})^{2}
  \;=\;
  \bm u^{\top}\Vc\bm u
  \;+\;
  \bm\omega_{q}^{\top}\Vc\bm\omega_{q},
  \qquad
  \bm u \;\coloneqq\; m(\gamma)\,\bm\omega_{\gamma}
       + \mathrm{diag}(\bm L^{\star})\,\bm\omega_{q},
\end{equation}
which one verifies by direct expansion using the diagonality of
$\Vc$ (and hence its commutativity with $\mathrm{diag}(\bm L^{\star})$).
For a quadratic form $\bm x^{\top}\Vc\bm x$ under a constraint
$\bm x^{\top}\bm 1=d$, the minimiser is proportional to
$\Vc^{-1}\bm 1$. Whether both terms in
\eqref{eq:pool-int:variance-factorisation} can be minimised
simultaneously by one pair
$(\bm\omega_{\gamma}, \bm\omega_{q})$ depends on whether the linear
constraint induced on $\bm u$ by the sum-to-one constraints is
sample-free. This happens when $\bm L^{\star}$ is a scalar multiple of
$\bm 1$; in the independent common-distribution distributed
specialisation of \Cref{sec:pool-int:setup}, this algebraic collapse is
induced by the equal-fraction condition
$k_{j}/n_{j} = (k/n)(1 + o(1))$.

\begin{proposition}[Variance-optimal weights, equal-fraction distributed inference]
\label{prop:pool-int:var-opt-distrib}
Under the assumptions of \Cref{thm:pool-int:main} and the
equal-effective-sample-fraction condition
$k_{j}/n_{j} = (k/n)(1 + o(1))$ described in
\Cref{sec:pool-int:setup}, the variance-optimal weights for
the pooled QB expectile estimator satisfy
\begin{equation}
\label{eq:pool-int:var-opt-distrib-coincide}
  \bm\omega^{\xi,\mathrm{Var}}_{q}
  \;=\;
  \bm\omega^{\mathrm{Var}}
  \;=\;
  \frac{\Vc^{-1}\bm 1}{\bm 1^{\top}\Vc^{-1}\bm 1},
\end{equation}
with corresponding asymptotic variance
\begin{equation}
\label{eq:pool-int:var-opt-distrib-value}
  \sigma^{\xi}\big(\bm\omega^{\xi,\mathrm{Var}}_{\gamma},
                    \bm\omega^{\xi,\mathrm{Var}}_{q}\big)^{2}
  \;=\;
  \big[(m(\gamma) + L^{\star})^{2} + 1\big]\,
  \big(\bm 1^{\top}\Vc^{-1}\bm 1\big)^{-1}.
\end{equation}
If $m(\gamma) \ne 0$, the variance-optimal pair is unique and
$\bm\omega^{\xi,\mathrm{Var}}_{\gamma} = \bm\omega^{\mathrm{Var}}$
as well, so both pooled inputs receive the quantile-pooling weights of
\textcite{DaouiaPadoanStupfler2021}. If $m(\gamma) = 0$, the
asymptotic variance does not depend
on $\bm\omega_{\gamma}$: any admissible vector is then
variance-optimal, and the matched representative
$\bm\omega^{\xi,\mathrm{Var}}_{\gamma} = \bm\omega^{\mathrm{Var}}$ is
one canonical choice.
\end{proposition}

\begin{proof}
The proposition is the specialisation $\bm L^{\star} = L^{\star}\bm
1$ of \Cref{prop:pool-int:var-opt-general}; see
\Cref{app:proofs:var-opt-general}, which discharges both the
$m(\gamma) \ne 0$ closed form and the reduced $m(\gamma) = 0$ case.
\end{proof}

\Cref{prop:pool-int:var-opt-distrib} is the central
variance-efficiency statement of this chapter: under equal-fraction
distributed inference, \emph{the variance-optimal pooling weights are
invariant under the change of target from extreme quantile to
extreme expectile.} The two targets share the same data-fusion
recipe; the only change in the asymptotic variance is the
multiplicative factor $(m(\gamma) + L^{\star})^{2} + 1$ replacing the
corresponding Hill/extreme-quantile factor of $1$ under the corresponding
normalisation. Under the stronger oracle-rate equal-fraction condition
of \Cref{cor:bg:pool-distrib-oracle}, this identifies the same pooling
rule that would enter the corresponding unfeasible pooled-sample
comparison for an expectile target; the present chapter uses only the
variance-optimality statement proved above, rather than a separate
oracle-equivalence theorem for a newly defined pooled-sample expectile
benchmark. The AMSE criterion requires separate bookkeeping because the
bias term is target-specific; as the second subsection below explains,
however, the equal-fraction distributed setting is also a collapse
regime for AMSE, while genuinely target-dependent AMSE weights can
appear once the effective sample fractions are unequal.

\paragraph{Beyond the equal-fraction setting.}
When the effective sample fractions $k_{j}/n_{j}$ vary across $j$,
$\mathrm{diag}(\bm L^{\star})$ ceases to be a scalar multiple of the
identity, the constraint
$\bm u^{\top}\bm 1
= m(\gamma) + \bm\omega_{q}^{\top}\bm L^{\star}$ acquires explicit
dependence on $\bm\omega_{q}$, and the decoupling argument used in
\Cref{prop:pool-int:var-opt-distrib} breaks. The variance-optimal
pair then lifts off the matched-weight diagonal and admits the
explicit closed form recorded below. Set
$\bm D \coloneqq \mathrm{diag}(\bm L^{\star})$ and abbreviate the
three scalar invariants of $(\Vc, \bm L^{\star})$:
\begin{equation}
\label{eq:pool-int:var-opt-abcDelta}
  a \;\coloneqq\; \bm 1^{\top}\Vc^{-1}\bm 1,
  \qquad
  b \;\coloneqq\; \bm 1^{\top}\Vc^{-1}\bm L^{\star},
  \qquad
  c \;\coloneqq\; (\bm L^{\star})^{\top}\Vc^{-1}\bm L^{\star},
  \qquad
  \Delta \;\coloneqq\; a(a + c) - b^{2}.
\end{equation}

\begin{proposition}[Variance-optimal weights, general case]
\label{prop:pool-int:var-opt-general}
Under the assumptions of \Cref{thm:pool-int:main} and $m(\gamma) \ne
0$, the Cauchy--Schwarz inequality in the $\Vc^{-1}$-inner product
gives $ac - b^{2} \ge 0$, with equality iff $\bm L^{\star}$ is a
scalar multiple of $\bm 1$. This is the limiting collapse induced by
the equal-effective-sample-fraction condition in the
distributed specialisation. Consequently
\[
  \Delta \;=\; a^{2} + (ac - b^{2}) \;\ge\; a^{2} \;>\; 0,
\]
with $\Delta = a^{2}$ exactly under the equal-fraction
distributed specialisation of
\Cref{prop:pool-int:var-opt-distrib}. The variance-optimal pooling
weights for the pooled QB intermediate expectile estimator are unique
and given by
\begin{align}
\label{eq:pool-int:var-opt-general-q}
  \bm\omega^{\xi,\mathrm{Var}}_{q}
    &\;=\;
    \frac{(a + c + m(\gamma)\,b)\,\Vc^{-1}\bm 1
          \;-\;(m(\gamma)\,a + b)\,\Vc^{-1}\bm L^{\star}}
         {\Delta},
  \\[2pt]
\label{eq:pool-int:var-opt-general-g}
  \bm\omega^{\xi,\mathrm{Var}}_{\gamma}
    &\;=\;
    \frac{(m(\gamma)\,a + b)\,\Vc^{-1}(\bm 1 + \bm D^{2}\bm 1)
          \;-\;(a + c + m(\gamma)\,b)\,\Vc^{-1}\bm L^{\star}}
         {m(\gamma)\,\Delta},
\end{align}
and the associated minimum asymptotic variance is
\begin{equation}
\label{eq:pool-int:var-opt-general-value}
  \sigma^{\xi}\!\big(\bm\omega^{\xi,\mathrm{Var}}_{\gamma},
                      \bm\omega^{\xi,\mathrm{Var}}_{q}\big)^{\!2}
  \;=\;
  \frac{(m(\gamma)^{2} + 1)\,a \;+\; 2\,m(\gamma)\,b \;+\; c}{\Delta}.
\end{equation}
Under the equal-effective-fraction condition
$\bm L^{\star} = L^{\star}\bm 1$
one has $b = L^{\star}a$, $c = (L^{\star})^{2}a$, $\Delta = a^{2}$;
formulas
\eqref{eq:pool-int:var-opt-general-q}--\eqref{eq:pool-int:var-opt-general-value}
then collapse to
\eqref{eq:pool-int:var-opt-distrib-coincide} and
\eqref{eq:pool-int:var-opt-distrib-value}, recovering
\Cref{prop:pool-int:var-opt-distrib}.
\end{proposition}

\begin{proof}
See \Cref{app:proofs:var-opt-general}, where the change of variables
$(\bm\omega_{\gamma}, \bm\omega_{q}) \mapsto (\bm u, \bm\omega_{q})$
with $\bm u = m(\gamma)\bm\omega_{\gamma} + \bm D\bm\omega_{q}$,
positivity of $\Delta$ from Cauchy--Schwarz, and the explicit
Lagrangian dual system are spelled out step-by-step.
\end{proof}

The assumption $m(\gamma) \ne 0$ is essential for the two-vector
closed form: it is the condition that makes the change of variables
$(\bm\omega_{\gamma}, \bm\omega_{q}) \mapsto (\bm u, \bm\omega_{q})$ a
bijection, and it appears in the denominator of
\eqref{eq:pool-int:var-opt-general-g}. When $m(\gamma) = 0$ the
asymptotic variance does not identify $\bm\omega_{\gamma}$, and the
optimisation collapses to the single-vector problem in
$\bm\omega_{q}$ recorded in \Cref{app:proofs:var-opt-general}.

\paragraph{Departure from the matched-weight diagonal.}
The discrepancy between $(\bm\omega^{\xi,\mathrm{Var}}_{\gamma},
\bm\omega^{\xi,\mathrm{Var}}_{q})$ and the Daouia--Padoan--Stupfler
matched-weight
pair $(\bm\omega^{\mathrm{Var}}, \bm\omega^{\mathrm{Var}})$ is
controlled by the $\Vc^{-1}$-Cauchy--Schwarz defect $ac - b^{2} \ge
0$, which vanishes exactly when $\bm L^{\star}$ is a scalar multiple
of $\bm 1$. Writing $\bar L^{\star} \coloneqq b/a$ for the
$\Vc^{-1}\bm 1$-weighted mean of $\bm L^{\star}$, a direct expansion
of \eqref{eq:pool-int:var-opt-general-q} delivers the closed-form
discrepancy
\begin{equation}
\label{eq:pool-int:var-opt-general-departure}
  \bm\omega^{\xi,\mathrm{Var}}_{q} - \bm\omega^{\mathrm{Var}}
  \;=\;
  -\,\frac{m(\gamma)\,a + b}{\Delta}\,
   \Vc^{-1}\big(\bm L^{\star} - \bar L^{\star}\bm 1\big),
\end{equation}
verified in \Cref{app:proofs:var-opt-general} (Step (v)), with an
analogous identity for $\bm\omega^{\xi,\mathrm{Var}}_{\gamma}$
obtained by inversion of the $\bm u$-definition. The correction lies
entirely along the $\Vc^{-1}$-centred direction
$\Vc^{-1}(\bm L^{\star} - \bar L^{\star}\bm 1)$. Writing $\Delta =
a^{2} + (ac - b^{2})$ exhibits its scalar prefactor as
\[
  \frac{m(\gamma)\,a + b}{\Delta}
  \;=\;
  \frac{m(\gamma) + \bar L^{\star}}{a}
  \cdot
  \left[1 \;+\; O\!\left(\frac{ac - b^{2}}{a^{2}}\right)\right]
\]
in the small Cauchy--Schwarz defect regime $ac - b^{2} \ll a^{2}$.
The quantile-pooling weights therefore remain near-optimal whenever the
log-rate vector is nearly sample-free ($ac - b^{2} \approx 0$, i.e.\
$\bm L^{\star}$ nearly constant in the $\Vc^{-1}\bm 1$-weighted
sense). The quantile-weight component
$\bm\omega^{\xi,\mathrm{Var}}_{q}$ also remains close to
$\bm\omega^{\mathrm{Var}}$ when
$m(\gamma) + \bar L^{\star} \approx 0$, which combines the
expectile-specific zero $\gamma^{\star} \approx 0.218$ of
$m(\,\cdot\,)$ with the log-rate weighted mean; this second condition
by itself controls the displayed $\bm\omega_q$ correction, not the full
two-vector pair. The degenerate case $m(\gamma) = 0$
leaves $\bm\omega_{\gamma}$ indeterminate and reduces the
optimisation to the single-vector problem $\min \bm\omega_{q}^{\top}
\Vc(\bm I + \bm D^{2})\bm\omega_{q}$ subject to
$\bm\omega_{q}^{\top}\bm 1 = 1$, whose solution is recorded in
\Cref{app:proofs:var-opt-general}; it coincides with
$\bm\omega^{\mathrm{Var}}$ in the equal-fraction setting and
deviates from it otherwise.

\subsection*{AMSE-optimal weights}

For the closed-form stacked optimisation in this subsection, assume
$m(\gamma)\ne0$. The degenerate case $m(\gamma)=0$ requires a reduced
optimisation because the variance term no longer identifies the Hill
weights; we return to that boundary case whenever it affects a stated
collapse result.

The AMSE-optimal pair
$(\bm\omega^{\xi,\mathrm{AMSE}}_{\gamma},
\bm\omega^{\xi,\mathrm{AMSE}}_{q})$ minimises the asymptotic mean
squared error
\begin{equation}
\label{eq:pool-int:amse-def}
  \mathrm{AMSE}^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})
  \;\coloneqq\;
  \big(\mu^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})\big)^{2}
  + \sigma^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})^{2}
\end{equation}
subject to the affine sum-to-one constraints
$\bm\omega_{\gamma}^{\top}\bm 1 = \bm\omega_{q}^{\top}\bm 1 = 1$, with
$\mu^{\xi}$ and $\sigma^{\xi}$ as in \eqref{eq:pool-int:mu-xi} and
\eqref{eq:pool-int:sigma-xi}. The objective parallels
\Cref{prop:bg:optimal-weights}(ii), but the expectile target reshapes
the bias landscape away from the equal-fraction collapse regime. In the
two-weight problem the bias is a linear functional of the stacked vector
$(\bm\omega_{\gamma}^{\top}, \bm\omega_{q}^{\top})^{\top}$ with
\emph{stacked gradient} $(m(\gamma)\Bc, \Bc^{\star})^{\top}$: the
pooled-Hill bias enters rescaled by the QB Jacobian $m(\gamma)$, while
the pooled-Weissman bias enters unrescaled. Only after restricting to
the matched-weight diagonal $\bm\omega_{\gamma} = \bm\omega_{q}$ does
one see the composite direction $m(\gamma)\Bc + \Bc^{\star}$, which is
the expectile analogue of the single quantile-pooling bias vector
$\Bc$. In addition, $\mu^{\xi}$ carries the additive offset
$b^{\xi/\Quant}$ of
\eqref{eq:pool-int:b-xi-Q-def}. This offset has no effect on the
variance problem, but it enters the AMSE first-order conditions whenever
the remaining bias part depends on the weights.

\paragraph{Stacked notation.}
To state the optimum compactly, stack the two pooling weight vectors
into $\bm w \coloneqq (\bm\omega_{\gamma}^{\top},
\bm\omega_{q}^{\top})^{\top} \in \R^{2m}$ and assemble the bias
gradient and variance Hessian implied by
\eqref{eq:pool-int:mu-xi}--\eqref{eq:pool-int:sigma-xi}:
\begin{equation}
\label{eq:pool-int:amse-stacked}
  \bm h \;\coloneqq\;
  \begin{pmatrix} m(\gamma)\,\Bc \\[2pt] \Bc^{\star} \end{pmatrix}
  \in \R^{2m},
  \qquad
  \bm Q \;\coloneqq\;
  \begin{pmatrix}
    m(\gamma)^{2}\,\Vc & m(\gamma)\,\bm D\,\Vc \\[2pt]
    m(\gamma)\,\bm D\,\Vc & \Vc + \bm D^{2}\Vc
  \end{pmatrix}
  \in \R^{2m \times 2m},
\end{equation}
so that
$\mu^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q}) = \bm h^{\top}\bm w +
b^{\xi/\Quant}$ and
$\sigma^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})^{2} = \bm w^{\top}
\bm Q\bm w$. Write $\bm A \in \R^{2m \times 2}$ for the matrix with
columns $\bm 1_{g} \coloneqq (\bm 1^{\top}, \bm 0^{\top})^{\top}$ and
$\bm 1_{q} \coloneqq (\bm 0^{\top}, \bm 1^{\top})^{\top}$, so that the
sum-to-one constraints read $\bm A^{\top}\bm w = \bm 1_{2}$.

\begin{proposition}[AMSE-optimal weights]
\label{prop:pool-int:amse-opt}
Under the assumptions of \Cref{thm:pool-int:main} and $m(\gamma) \ne
0$, the matrix $\bm M \coloneqq \bm Q + \bm h\bm h^{\top}$ is
positive definite, and the AMSE objective \eqref{eq:pool-int:amse-def}
is strictly convex on the affine constraint set
$\{\bm w \in \R^{2m} : \bm A^{\top}\bm w = \bm 1_{2}\}$. The unique
minimiser admits the closed form
\begin{equation}
\label{eq:pool-int:amse-opt-formula}
  \bm w^{\xi,\mathrm{AMSE}}
  \;=\;
  \bm M^{-1}\!\left[\bm A\,(\bm A^{\top}\bm M^{-1}\bm A)^{-1}\!
                     \big(\bm 1_{2} \;+\; b^{\xi/\Quant}\,
                          \bm A^{\top}\bm M^{-1}\bm h\big)
                    \;-\; b^{\xi/\Quant}\,\bm h\right],
\end{equation}
and is equivalently characterised by the first-order conditions
\begin{equation}
\label{eq:pool-int:amse-foc}
\begin{aligned}
  m(\gamma)^{2}\,\Vc\,\bm\omega_{\gamma}
  + m(\gamma)\,\bm D\,\Vc\,\bm\omega_{q}
  &\;=\; \tfrac{1}{2}\lambda_{\gamma}\bm 1
       \;-\; m(\gamma)\,\mu^{\xi}\,\Bc, \\
  m(\gamma)\,\bm D\,\Vc\,\bm\omega_{\gamma}
  + \big(\Vc + \bm D^{2}\Vc\big)\,\bm\omega_{q}
  &\;=\; \tfrac{1}{2}\lambda_{q}\bm 1
       \;-\; \mu^{\xi}\,\Bc^{\star},
\end{aligned}
\end{equation}
together with the sum-to-one constraints
$\bm\omega_{\gamma}^{\top}\bm 1 = \bm\omega_{q}^{\top}\bm 1 = 1$ and
the self-referential identity
$\mu^{\xi} = m(\gamma)\,\bm\omega_{\gamma}^{\top}\Bc +
\bm\omega_{q}^{\top}\Bc^{\star} + b^{\xi/\Quant}$. The Lagrange
multipliers $\lambda_{\gamma}, \lambda_{q} \in \R$ are uniquely
determined by these constraints.
\end{proposition}

\begin{proof}
See \Cref{app:proofs:amse-opt}, where positive-definiteness of $\bm
Q$ (and hence of $\bm M = \bm Q + \bm h\bm h^{\top}$) is verified in
Step (ii) via the variance factorisation, the Lagrangian and
explicit FOC normalisation are recorded in Step (iii), and the block
decomposition matching \eqref{eq:pool-int:amse-foc} is in Step (iv).
\end{proof}

\paragraph{Equal-fraction AMSE collapse.}
Under the equal-effective-sample-fraction condition
$k_j/n_j=(k/n)(1+o(1))$, the log-rate vector is
$\bm L^{\star}=L^{\star}\bm 1$. The compatibility identity following
\eqref{eq:pool-int:Lambda-bullet} then gives
\[
  \sqrt{c_jS}\,\lambda_j
  =
  \Lambda^{\bullet}(e^{L^{\star}})^{-\rho},
  \qquad S=\sum_{i=1}^{m}c_i^{-1},
\]
which is sample-free. Combining this with
\eqref{eq:pool-int:Bcstar-def} yields
\[
  \Bc=\beta_H\bm 1,
  \qquad
  \Bc^{\star}=\beta_W\bm 1
\]
for constants $\beta_H,\beta_W$ depending on
$(\gamma,\rho,L^\star,\Lambda^\bullet)$ but not on $j$. Hence, for all
admissible weight pairs,
\[
  \mu^\xi(\bm\omega_\gamma,\bm\omega_q)
  =
  m(\gamma)\beta_H+\beta_W+b^{\xi/\Quant},
\]
which is constant on the constraint set. In this regime the AMSE
criterion is therefore the variance criterion plus a constant. If
$m(\gamma)\ne0$, the unique AMSE-optimal pair is the same matched pair
as in \Cref{prop:pool-int:var-opt-distrib},
\[
  \bm\omega_\gamma^{\xi,\mathrm{AMSE}}
  =
  \bm\omega_q^{\xi,\mathrm{AMSE}}
  =
  \bm\omega^{\mathrm{Var}},
\]
and when $m(\gamma)=0$ the same indeterminacy of
$\bm\omega_\gamma$ described in
\Cref{prop:pool-int:var-opt-distrib} remains. This is exactly parallel
to the equal-fraction distributed setting of
\textcite{DaouiaPadoanStupfler2021}: there too the quantile-pooling
bias vector $\Bc$ is collinear with $\bm 1$, so the AMSE criterion
collapses to the variance criterion under the affine constraint.

\paragraph{Comparison with the Daouia--Padoan--Stupfler AMSE optimum away from equal fractions.}
Outside the equal-fraction collapse regime, the expectile
AMSE-optimum generally does \emph{not} reduce to the
Daouia--Padoan--Stupfler
AMSE-optimum of \Cref{prop:bg:optimal-weights}(ii). Three structural
differences are responsible:
\begin{enumerate}[label=(\roman*)]
  \item the bias gradient is the stacked vector
        $\bm h = (m(\gamma)\Bc, \Bc^{\star})^{\top}$ rather than the
        single-block $\Bc$ of \Cref{prop:bg:optimal-weights};
  \item $\mu^{\xi}$ carries the additive offset
        $b^{\xi/\Quant}$, which enters the AMSE first-order conditions
        through the self-consistent mean $\mu^\xi$;
  \item the expectile minimiser is genuinely a pair
        $(\bm\omega_{\gamma}, \bm\omega_{q})$, and the diagonal
        $\bm\omega_{\gamma} = \bm\omega_{q}$ is generically not a
        solution to the FOC \eqref{eq:pool-int:amse-foc} because the
        Hill and Weissman blocks of $\bm h$ enter asymmetrically
        through $m(\gamma)\Bc$ and $\Bc^{\star}$.
\end{enumerate}
Exact coincidence with the Daouia--Padoan--Stupfler AMSE weights can
occur, but it is a
compatibility condition on the full FOC system rather than a simple
proportionality of $\Bc$ and $\Bc^\star$. One would have to insert the
Daouia--Padoan--Stupfler candidate weight into
\eqref{eq:pool-int:amse-foc}, impose the two affine constraints, and
solve the resulting block equations in the quantities
$(\Vc,\bm D,\Bc,\Bc^{\star},b^{\xi/\Quant})$. We do not characterise
that exceptional locus here. The robust qualitative point is narrower
and sharper: the equal-fraction distributed specialisation preserves
the Daouia--Padoan--Stupfler weights for both variance and AMSE, whereas unequal
effective sample fractions generically make the expectile AMSE
calibration target-dependent.

\section{Estimated weights and confidence intervals}
\label{sec:pool-int:CI}

The optimal weights of
\Cref{prop:pool-int:var-opt-distrib,prop:pool-int:var-opt-general,prop:pool-int:amse-opt}
are unknown in practice: they depend on the population covariance
$\Vc$, on the log-rate vector $\bm L^{\star}$, on $m(\gamma)$, and ---
for the AMSE optimum --- on the bias quantities $\Bc, \Bc^{\star}$
and the expectile--quantile offset $b^{\xi/\Quant}$. We follow the
plug-in scheme of \textcite[\S2.2]{DaouiaPadoanStupfler2021} recalled
in \Cref{sec:bg:pooling}: each population quantity is replaced by a
consistent empirical analogue, and the resulting plug-in weights are
substituted into the pooled estimator
\eqref{eq:pool-int:expectile-pool}. This section organises the
plug-in inputs by what each optimum needs, states the resulting
confidence interval under undersmoothing (\Cref{cor:pool-int:CI})
with the variance-optimal plug-in as the default implementation, and
contrasts it with the bias-corrected variant required in the
AMSE-optimal regime (\Cref{rem:pool-int:CI-bias-correction}).

\subsection*{Feasible plug-in weights}

The three optima of \Cref{sec:pool-int:weights} have different
plug-in demands. Listing them in order of increasing complexity:

\begin{enumerate}[label=(\alph*), leftmargin=*, itemsep=2pt]
\item \emph{Equal-fraction variance-optimal weights}
(\Cref{prop:pool-int:var-opt-distrib}).
Under $k_{j}/n_{j} = (k/n)(1 + o(1))$ the optimum collapses to
$\bm\omega^{\mathrm{Var}} = \Vc^{-1}\bm 1 /
(\bm 1^{\top}\Vc^{-1}\bm 1)$. Since the independent-samples covariance
\eqref{eq:bg:V_c-diagonal} is diagonal with a common multiplicative
factor, this population vector has coordinates
$\omega_{j}^{\mathrm{Var}} \propto c_{j}^{-1}$. The canonical
distributed plug-in is therefore the effective-sample-size weight used
by \textcite[\S3.2]{DaouiaPadoanStupfler2021},
\begin{equation}
\label{eq:pool-int:plugin-eqfrac}
  \hat\omega_{j}^{\mathrm{Var}}
  \;=\;
  \frac{k_{j}}{k},
  \qquad j = 1,\ldots,m.
\end{equation}
Equivalently, any consistent diagonal covariance plug-in $\hatVc
\toprob\Vc$ may be inserted into the generic formula of
\textcite{DaouiaPadoanStupfler2021},
$\hatVc^{-1}\bm 1/(\bm 1^\top\hatVc^{-1}\bm 1)$; it has the same limit.
These are two asymptotically equivalent implementations, not the same
finite-sample rule unless $\hatVc$ has the distributed diagonal shape
proportional to $\mathrm{diag}(k/k_1,\ldots,k/k_m)$. Below the default
interval uses the non-random vector \eqref{eq:pool-int:plugin-eqfrac}
and evaluates the variance by direct substitution into
\eqref{eq:pool-int:sigma-xi}; if the inverse-$\hatVc$ rule is used
instead, the denominator form
$(\bm 1^\top\hatVc^{-1}\bm 1)^{-1}$ is exact for that alternative
plug-in.

\item \emph{General variance-optimal weights}
(\Cref{prop:pool-int:var-opt-general}).
When $k_{j}/n_{j}$ varies across $j$, the closed forms
\eqref{eq:pool-int:var-opt-general-q}--\eqref{eq:pool-int:var-opt-general-g}
also depend on $m(\gamma)$ and on the log-rate vector $\bm L^{\star}$.
The plug-in therefore additionally requires
$\hat m \toprob m(\gamma)$ and $L_{j,n} \to L_{j}^{\star}$;
the latter is deterministic ($L_{j,n} \coloneqq
L_{j}(\tau_{n}) = \log(k_{j}/(n_{j}(1 - \tau_{n})))$ converges by the
chapter setup of \Cref{sec:pool-int:setup}), so the only random
plug-in beyond $\hatVc$ is $\hat m$. To avoid circularity --- the
general formula for $\hat{\bm\omega}_{\gamma}$ would otherwise
depend on $\hat m$, which would itself depend on
$\hat{\bm\omega}_{\gamma}$ --- we seed $\hat m$ from a preliminary
consistent pooled Hill estimator that needs no weight optimisation,
for instance the effective-sample-size pool
\begin{equation}
\label{eq:pool-int:plugin-m}
  \hat m^{(0)}
  \;\coloneqq\;
  m\!\Big(\sum_{j=1}^{m}\!(k_{j}/k)\,\hat\gamma_{j}(k_{j})\Big),
  \qquad
  \hat m^{(0)} \toprob m(\gamma),
\end{equation}
which is consistent under \Cref{thm:bg:hill} and avoids the
self-reference. One may optionally iterate after computing the
weights --- updating $\hat m \leftarrow m(\hat\gamma_{n}
(\hat{\bm\omega}_{\gamma}))$ once the weights have been formed --- but
this is asymptotically immaterial. The closed forms
\eqref{eq:pool-int:var-opt-general-q}--\eqref{eq:pool-int:var-opt-general-g}
are used only away from the interior root of $m(\,\cdot\,)$: formally,
one may assume $|m(\gamma)|>\varepsilon$ and compute the plug-in on the
event $|\hat m|>\varepsilon/2$, whose probability tends to one. If
$|\hat m|$ is numerically close to zero, the variance problem should be
treated through the $m(\gamma)=0$ single-vector solution displayed in
\eqref{eq:app:varopt:m-zero}, with $\bm\omega_{\gamma}$ supplied by a
fixed canonical affine choice such as \eqref{eq:pool-int:plugin-eqfrac}.

\item \emph{AMSE-optimal weights} (\Cref{prop:pool-int:amse-opt}).
The closed form \eqref{eq:pool-int:amse-opt-formula} depends on all
of $\Vc, \bm L^{\star}, m(\gamma), \Bc, \Bc^{\star}, b^{\xi/\Quant}$
through the stacked vector $\bm h$ and Hessian $\bm Q$ of
\eqref{eq:pool-int:amse-stacked}. The plug-in therefore requires,
on top of $\hatVc$ and $\hat m$, consistent estimators of the
second-order quantities
\begin{equation}
\label{eq:pool-int:plugin-amse}
  \hatBc \toprob \Bc,
  \qquad
  \hatBc^{\star} \toprob \Bc^{\star},
  \qquad
  \hat b^{\xi/\Quant} \toprob b^{\xi/\Quant}.
\end{equation}
These convergences are not automatic consequences of the abstract
second-order condition $\mathcal C_{2}(\gamma,\rho,A)$: that condition
specifies the limiting auxiliary function $A$, not a directly
estimable finite-dimensional parameter. Operationally one needs an
additional second-order working model, such as the Hall-type
parametrisation $A(t)=\gamma\beta t^{\rho}$ used by
\textcite[\S2.2]{DaouiaPadoanStupfler2021}. Under such a parametrisation,
machinewise estimators $\hat\beta_j,\hat\rho_j$ of the common
second-order parameters can be inserted into
\eqref{eq:pool-int:Bcstar-def}--\eqref{eq:pool-int:b-xi-Q-def} and into
the formula for $\Bc$ from \Cref{thm:bg:pool-hill}. Without this extra
model, \eqref{eq:pool-int:plugin-amse} should be read as a high-level
assumption that consistent estimators of the limiting bias constants are
available, not as a consequence of $\mathcal C_{2}(\gamma,\rho,A)$
alone.
In the absence of such a model and of estimators satisfying
\eqref{eq:pool-int:plugin-amse}, the AMSE-optimal formula is best read
as an oracle or sensitivity-analysis device rather than as a fully
feasible procedure. Thus a practical plug-in AMSE implementation is
exploratory unless the application supplies, and defends, a stable
second-order calibration. The closed form in \Cref{prop:pool-int:amse-opt} is
also not claimed at the root $m(\gamma)=0$; confidence intervals at
that root require an externally supplied consistent affine weight pair.
\end{enumerate}

The consistency argument is a local one. At the population value,
$\bm 1^{\top}\Vc^{-1}\bm 1>0$; in the general variance formula
$\Delta>0$ and, on the region where it is used, $m(\gamma)\ne0$; in the
AMSE formula $\bm M$ is positive definite and
$\det(\bm A^{\top}\bm M^{-1}\bm A)>0$. Continuity therefore gives a
neighbourhood of the population arguments on which the relevant
denominators remain non-zero. The assumed plug-in convergences place
the estimated arguments in this neighbourhood with probability tending
one, and the continuous-mapping theorem then transfers the convergences
above into consistency
$\hat{\bm\omega}_{\gamma} \toprob \bm\omega_{\gamma},
\hat{\bm\omega}_{q} \toprob \bm\omega_{q}$ of the optimal-weight
plug-ins, and from there into consistency of the plug-in variance
$\hat\sigma^{\xi}(\hat{\bm\omega}_{\gamma}, \hat{\bm\omega}_{q})
\toprob \sigma^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})$ via
\eqref{eq:pool-int:sigma-xi}. By construction the closed forms
\eqref{eq:pool-int:plugin-eqfrac},
\eqref{eq:pool-int:var-opt-general-q}--\eqref{eq:pool-int:var-opt-general-g}
and \eqref{eq:pool-int:amse-opt-formula} are normalised against the
affine sum-to-one constraints, so $\hat{\bm\omega}_{\gamma}^{\top}\bm 1
= \hat{\bm\omega}_{q}^{\top}\bm 1 = 1$ holds \emph{exactly}, not just
in probability --- a property that the estimated-weight lift in
\Cref{cor:pool-int:CI} below relies on.

\paragraph{Practical recommendation.}
The complexity gradient (a) $\to$ (b) $\to$ (c) above is also a
finite-sample stability gradient. Estimators of the second-order
parameter $\rho$ and the rate function $A(\,\cdot\,)$ that
\eqref{eq:pool-int:plugin-amse} relies on are notoriously delicate in
the heavy-tailed regime; see the bias-correction discussion of
\textcite[\S3.1.1]{DaouiaGirardStupfler2018} and
\textcite[Remark~4]{DaouiaPadoanStupfler2021}. Because such
second-order estimators are often unstable in finite samples, the
variance-optimal plug-ins (a)--(b) are the conservative default unless
the application supplies reliable second-order calibration for
$\rho$ and $A$.

\subsection*{Confidence interval under undersmoothing}

The asymptotic bias \eqref{eq:pool-int:mu-xi} is the linear
combination
\[
  \mu^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})
  \;=\;
  m(\gamma)\,\bm\omega_{\gamma}^{\top}\Bc
  \;+\; \bm\omega_{q}^{\top}\Bc^{\star}
  \;+\; b^{\xi/\Quant}.
\]
The formulas \eqref{eq:pool-int:Bcstar-def} and
\Cref{thm:bg:pool-hill} show that $\Bc$ and $\Bc^\star$ are continuous
linear combinations of the per-sample bias rates
$\lambda_{j} \coloneqq \lim \sqrt{k_{j}}\,A(n_{j}/k_{j})$. The offset
$b^{\xi/\Quant}$ is controlled by the compatibility identity following
\eqref{eq:pool-int:Lambda-bullet},
\[
  \Lambda^{\bullet}
  =
  \sqrt{c_jS}\,\lambda_j e^{\rho L_j^\star},
  \qquad S=\sum_{i=1}^{m}c_i^{-1},
  \qquad j=1,\ldots,m,
\]
which identifies the common limit
$\sqrt{k}A((1-\tau_n)^{-1})$ through any sample $j$.
The \emph{undersmoothing calibration}
\begin{equation}
\label{eq:pool-int:undersmoothing}
  \sqrt{k_{j}}\,A(n_{j}/k_{j}) \;\to\; 0
  \qquad\text{for every } j = 1, \ldots, m,
\end{equation}
therefore forces $\Bc = \bm 0$, $\Bc^{\star} = \bm 0$,
$\Lambda^{\bullet} = 0$, $b^{\xi/\Quant} = 0$, and consequently
$\mu^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q}) = 0$ for every
admissible pair $(\bm\omega_{\gamma}, \bm\omega_{q})$. This is the
hypothesis under which the variance-only confidence interval below
is asymptotically valid. Without \eqref{eq:pool-int:undersmoothing}
the limit-bias is generically non-zero and the interval must be
re-centred on a bias-corrected point estimator
(\Cref{rem:pool-int:CI-bias-correction}).

\begin{corollary}[Asymptotic confidence interval for $\expectile{\tau_{n}}$]
\label{cor:pool-int:CI}
Assume the setting of \Cref{thm:pool-int:main} together with the
undersmoothing condition \eqref{eq:pool-int:undersmoothing}, and let
$(\hat{\bm\omega}_{\gamma}, \hat{\bm\omega}_{q})$ be estimated weight
pairs satisfying the affine sum-to-one constraints
$\hat{\bm\omega}_{\gamma}^{\top}\bm 1
= \hat{\bm\omega}_{q}^{\top}\bm 1 = 1$ exactly and the consistency
conditions
\begin{equation}
\label{eq:pool-int:CI-consistency}
  \hat{\bm\omega}_{\gamma} \toprob \bm\omega_{\gamma},
  \qquad
  \hat{\bm\omega}_{q} \toprob \bm\omega_{q},
  \qquad
  \hatVc \toprob \Vc,
  \qquad
  \hat m \toprob m(\gamma),
\end{equation}
for some deterministic limit pair $(\bm\omega_{\gamma},
\bm\omega_{q})$ on the affine constraint set. By the right-heavy-tailed
setup and the expectile--quantile equivalence,
$\expectile{\tau_n}>0$ eventually; the local Weissman factors, and
hence the pooled estimator, are positive with probability tending one.
Thus the logarithms below are well defined on an event whose probability
tends to one. Then, for every $\alpha \in (0, 1)$, the random log-scale
interval
\begin{equation}
\label{eq:pool-int:CI-log}
  I_{n}^{\log}
  \;\coloneqq\;
  \log \hatexpectile{\tau_{n}}^{\mathrm{pool}}\!
       (\hat{\bm\omega}_{\gamma}, \hat{\bm\omega}_{q})
  \;\pm\;
  z_{1 - \alpha/2}\,
  \frac{\hat\sigma^{\xi}(\hat{\bm\omega}_{\gamma},
                         \hat{\bm\omega}_{q})}{\sqrt{k}}
\end{equation}
and the associated original-scale interval
\[
  I_{n} \coloneqq \exp(I_{n}^{\log})
  = [\exp(a_n), \exp(b_n)]
  \quad\text{when } I_{n}^{\log}=[a_n,b_n],
\]
are asymptotic $(1 - \alpha)$ confidence intervals
for $\log \expectile{\tau_{n}}$ and $\expectile{\tau_{n}}$
respectively:
\[
  \Prob\!\big(\log \expectile{\tau_{n}} \in I_{n}^{\log}\big)
  \;\to\; 1 - \alpha,
  \qquad
  \Prob\!\big(\expectile{\tau_{n}} \in I_{n}\big)
  \;\to\; 1 - \alpha,
\]
as $n \to \infty$, where $z_{1 - \alpha/2}$ is the standard-normal
quantile of order $1 - \alpha/2$ and $\hat\sigma^{\xi}$ is the
plug-in version of \eqref{eq:pool-int:sigma-xi}.
\end{corollary}

\begin{proof}
See \Cref{app:proofs:CI}, where the four-step argument is spelled
out: (i) the log-scale recasting of \Cref{thm:pool-int:main}; (ii)
the lift to estimated weights via the exact sum-to-one constraint and
Slutsky; (iii) self-normalisation through the plug-in variance under
undersmoothing; (iv) transport of coverage to the original scale by
strict monotonicity of $\exp$.
\end{proof}

\paragraph{Practical remarks.}
First, the log-scale interval \eqref{eq:pool-int:CI-log} is symmetric
in $\log \expectile{\tau_{n}}$ while its exponential is not; the
latter is the natural form for applications in which
$\expectile{\tau_{n}}$ is interpreted as a monetary risk measure.
Second, under the equal-fraction distributed specialisation of
\Cref{prop:pool-int:var-opt-distrib} the variance-optimal weights are
identical across the two pooled inputs. With the canonical
effective-sample-size weights
$\hat{\bm\omega}^{k}=(k_1/k,\ldots,k_m/k)^\top$ of
\eqref{eq:pool-int:plugin-eqfrac}, the finite-sample plug-in variance
is the direct substitution
\[
  \hat\sigma^{\xi}(\hat{\bm\omega}^{k},
                    \hat{\bm\omega}^{k})^{2}
  \;=\;
  \big[(\hat m + L_{n})^{2} + 1\big]\,
  (\hat{\bm\omega}^{k})^{\top}\hatVc\,\hat{\bm\omega}^{k},
  \qquad
  L_{n} \;=\; \log\!\big(k/(n(1 - \tau_{n}))\big).
\]
If one instead uses the inverse-covariance plug-in
$\hat{\bm\omega}^{V}=\hatVc^{-1}\bm 1/
(\bm 1^\top\hatVc^{-1}\bm 1)$, the corresponding variance simplifies
exactly to
\[
  \hat\sigma^{\xi}(\hat{\bm\omega}^{V},
                    \hat{\bm\omega}^{V})^{2}
  \;=\;
  \big[(\hat m + L_{n})^{2} + 1\big]
  \big/ \big(\bm 1^{\top}\hatVc^{-1}\bm 1\big),
\]
the direct plug-in image of the limit
\eqref{eq:pool-int:var-opt-distrib-value}. The two variance estimates
have the same limit, and they agree exactly when $\hatVc$ has the
distributed diagonal structure described above.

\paragraph{CI validity at $m(\gamma) = 0$.}
The hypotheses of \Cref{cor:pool-int:CI} do \emph{not} require
$m(\gamma) \ne 0$: the proof of \Cref{app:proofs:CI} uses only
(i) consistency of the estimated weights toward a deterministic
affine weight pair, (ii) consistency of the plug-in variance, and (iii)
positivity of the limit variance $\sigma^{\xi}(\bm\omega_{\gamma},
\bm\omega_{q})^{2}$. The restriction $m(\gamma) \ne 0$ enters
elsewhere: it is needed for the variance-optimal closed form of
\Cref{prop:pool-int:var-opt-general} (because the change of
variables $(\bm\omega_{\gamma}, \bm\omega_{q}) \mapsto
(\bm u, \bm\omega_{q})$ ceases to be a bijection at $m(\gamma) = 0$)
and for the AMSE positive-definiteness of
\Cref{prop:pool-int:amse-opt}. At the root $\gamma = \gamma^{\star}
\approx 0.2178$, the equal-fraction variance-optimal
$\hat{\bm\omega}^{k}$ of \eqref{eq:pool-int:plugin-eqfrac}
remains well defined and consistent, and the CI
\eqref{eq:pool-int:CI-log} retains its $(1 - \alpha)$ asymptotic
coverage; the variance simply reduces to
$\hat\sigma^{\xi,\,2} = [L_{n}^{2} + 1]\,
(\hat{\bm\omega}^{k})^\top\hatVc\,\hat{\bm\omega}^{k}$. For the
inverse-covariance plug-in $\hat{\bm\omega}^{V}$, this same statement
uses $[L_{n}^{2} + 1]/(\bm 1^\top\hatVc^{-1}\bm 1)$.

\begin{remark}[Bias-corrected confidence interval]
\label{rem:pool-int:CI-bias-correction}
The undersmoothing condition \eqref{eq:pool-int:undersmoothing} is
the natural companion of the variance-only interval
\eqref{eq:pool-int:CI-log}. It is a different asymptotic calibration
from the second-order regime in which the bias component of
\eqref{eq:pool-int:mu-xi} is retained and the AMSE weight criterion of
\Cref{prop:pool-int:amse-opt} is meaningful. In that regime the
limit-bias $\mu^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})$ need not
vanish. When
$\mu^{\xi} \ne 0$, \eqref{eq:pool-int:CI-log} does not have nominal
asymptotic coverage. The standard remedy is to use the log-scale
bias-corrected interval
\[
  I_{n,\mathrm{bc}}^{\log}
  \;\coloneqq\;
  \log \hatexpectile{\tau_{n}}^{\mathrm{pool}}\!
       (\hat{\bm\omega}_{\gamma}, \hat{\bm\omega}_{q})
  \;-\;
  \frac{\hat\mu^{\xi}(\hat{\bm\omega}_{\gamma},
                     \hat{\bm\omega}_{q})}{\sqrt{k}}
  \;\pm\;
  z_{1-\alpha/2}
  \frac{\hat\sigma^{\xi}(\hat{\bm\omega}_{\gamma},
                         \hat{\bm\omega}_{q})}{\sqrt{k}},
\]
where $\hat\mu^{\xi}$ substitutes $\hatBc, \hatBc^{\star},
\hat b^{\xi/\Quant}$ and $\hat m$ into \eqref{eq:pool-int:mu-xi}. This
interval has the same asymptotic coverage argument as
\Cref{cor:pool-int:CI} provided all ingredients used there remain
valid and, in addition,
\[
  \hat\mu^{\xi}(\hat{\bm\omega}_{\gamma}, \hat{\bm\omega}_{q})
  \toprob \mu^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q}),\qquad
  \hat\sigma^{\xi}(\hat{\bm\omega}_{\gamma}, \hat{\bm\omega}_{q})
  \toprob \sigma^{\xi}(\bm\omega_{\gamma}, \bm\omega_{q})>0,
\]
with
$\hat{\bm\omega}_{\gamma}\toprob\bm\omega_{\gamma}$ and
$\hat{\bm\omega}_{q}\toprob\bm\omega_{q}$ under exact affine
normalisation and the same high-probability well-definedness
conditions for logarithms and square roots. If the interval is paired
with AMSE-optimal weights, these requirements include consistency of
the AMSE plug-in weights and the second-order bias estimators in
\eqref{eq:pool-int:plugin-amse}, a hypothesis substantially more
demanding than \eqref{eq:pool-int:CI-consistency} because it requires
consistent estimation of the second-order parameter $\rho$ and the rate
function $A$ \parencite[\S5.2]{deHaanFerreira2006}.
\end{remark}

Standard practice in the extreme-value literature accordingly favours
the variance-only interval \eqref{eq:pool-int:CI-log} under
\eqref{eq:pool-int:undersmoothing} over the bias-corrected variant of
\Cref{rem:pool-int:CI-bias-correction}: undersmoothing trades a
small amount of asymptotic efficiency for the absence of a
hard-to-estimate correction and places inference in a
variance-dominated regime. This trade-off is the operational reason why
\Cref{prop:pool-int:var-opt-distrib} --- not
\Cref{prop:pool-int:amse-opt} --- is the natural default companion to
\Cref{cor:pool-int:CI} in applications.

\section{A root-\texorpdfstring{$k$}{k} diagnostic for high-expectile homogeneity}
\label{sec:pool-int:test}

The variance-optimal pooling of \Cref{prop:pool-int:var-opt-distrib}
rests on the common-$\gamma$ assumption $(H_{\gamma})$ and, in the
distributed-inference interpretation, on a common high-expectile target
across the $m$ samples. Although the chapter assumes a common marginal
$F$ throughout, no finite-dimensional statistic can validate equality
of the entire marginal distribution. The statistic below should
therefore be read more narrowly: it is a root-$k$ diagnostic for
equality of the high-expectile targets being pooled. It can reveal
tail-index or tail-scale incompatibilities that affect the expectile
level, but failure to reject does not validate the full common-marginal
model. This mirrors the chi-square test of
\Cref{thm:bg:tail-homogeneity-test}, which diagnoses tail-index
homogeneity rather than equality of complete distributions.

Let
\begin{equation}
\label{eq:pool-int:test-per-sample}
  \hatexpectile{\tau_{n},\,j}
  \;\coloneqq\;
  \psifn\!\big(\hat\gamma_{j}(k_{j})\big)\,
  \hat q^{\star}_{j}(\tau_{n} \mid k_{j}),
  \qquad j = 1, \ldots, m,
\end{equation}
denote the per-sample QB intermediate-expectile estimators of
\Cref{sec:pool-int:setup} --- one per machine, each assembled from
that machine's own pooling primitives
$(\hat\gamma_{j}(k_{j}),\,\hat q^{\star}_{j})$ --- and collect their
logarithms into the vector
\[
  \hat{\bm\Xi}_{n}
  \;\coloneqq\;
  \big(\log\hatexpectile{\tau_{n},\,1}, \ldots,
       \log\hatexpectile{\tau_{n},\,m}\big)^{\!\top}.
\]
As in \Cref{sec:pool-int:estimator}, all logarithms are understood on
the high-probability event on which the local Weissman factors are
positive and the local Hill estimators entering $\log\psifn$ lie in
$(0,1)$; arbitrary definitions on the complementary event do not change
the weak limits.

\paragraph{Joint limit of the per-sample log-expectiles.}
Applying the expectile-specific Jacobian row $(m(\gamma),\,1)$ ---
the same linearisation of $\log\psifn$ that produced
\Cref{thm:pool-int:main} --- to the within-sample $2$-vector
\eqref{eq:pool-int:per-sample-clt} of each sample, and aggregating
across the $m$ independent samples on the common $\sqrt{k}$ scale,
yields the joint central limit theorem
\begin{equation}
\label{eq:pool-int:test-clt}
  \sqrt{k}\,\big(\hat{\bm\Xi}_{n}
                 - \log\expectile{\tau_{n}}\,\bm 1\big)
  \;\todist\;
  \mathcal N\!\big(\bm\beta,\, \bm\Sigma_{\Xi}\big),
\end{equation}
with diagonal covariance matrix
\begin{equation}
\label{eq:pool-int:test-Sigma-Xi}
  \bm\Sigma_{\Xi}
  \;=\;
  \mathrm{diag}\!\big((m(\gamma) + L_{j}^{\star})^{2} + 1\big)_{j = 1}^{m}\,
  \Vc,
  \qquad
  (\bm\Sigma_{\Xi})_{j, j}
  \;=\;
  \big[(m(\gamma) + L_{j}^{\star})^{2} + 1\big]\,(\Vc)_{j, j},
\end{equation}
and asymptotic mean vector
\begin{equation}
\label{eq:pool-int:test-beta}
  \bm\beta
  \;=\;
  m(\gamma)\,\Bc \;+\; \Bc^{\star} \;+\; b^{\xi/\Quant}\,\bm 1,
\end{equation}
where $\Vc$, $\Bc$, $\Bc^{\star}$, $b^{\xi/\Quant}$ and the log-rate
limits $L_{j}^{\star}$ are exactly the quantities of
\Cref{prop:pool-int:joint-clt,thm:pool-int:main}. The covariance
\eqref{eq:pool-int:test-Sigma-Xi} is diagonal because the $m$ samples
are independent \eqref{eq:pool-int:R-zero}; its $j$th diagonal entry
is the $\sqrt{k}$-rescaled per-sample log-expectile variance
$\gamma^{2}\big[(m(\gamma) + L_{j}^{\star})^{2} + 1\big]\,(k/k_{j})$,
read off from \eqref{eq:pool-int:per-sample-clt} through the row
$(m(\gamma), 1)$ and recognising $\gamma^{2}\,(k/k_{j}) \to
(\Vc)_{j, j}$. The mean \eqref{eq:pool-int:test-beta} is the
per-sample analogue of the pooled bias \eqref{eq:pool-int:mu-xi},
the deterministic expectile--quantile offset $b^{\xi/\Quant}$ of
\eqref{eq:pool-int:b-xi-Q-def} entering every coordinate equally (it
is sample-free, hence collinear with $\bm 1$). The derivation of
\eqref{eq:pool-int:test-clt} is recorded in
\Cref{app:proofs:hom-test}.

\paragraph{The GLS deviance statistic.}
Under the common-marginal assumption all $\hatexpectile{\tau_{n},\,j}$
estimate the same population $\expectile{\tau_{n}}$, and their
log-scale dispersion about the generalised-least-squares (GLS)
weighted mean is captured by the deviance statistic
\begin{equation}
\label{eq:pool-int:test-stat}
  \Lambda^{\xi}_{n}
  \;\coloneqq\;
  k\,\big(\hat{\bm\Xi}_{n} - \hat\nu_{n}\bm 1\big)^{\!\top}
    \hat{\bm\Sigma}_{\Xi,n}^{-1}
    \big(\hat{\bm\Xi}_{n} - \hat\nu_{n}\bm 1\big),
  \qquad
  \hat\nu_{n}
  \;=\;
  \frac{\bm 1^{\top}\hat{\bm\Sigma}_{\Xi,n}^{-1}\hat{\bm\Xi}_{n}}
       {\bm 1^{\top}\hat{\bm\Sigma}_{\Xi,n}^{-1}\bm 1},
\end{equation}
in which $\hat\nu_{n}$ is the GLS estimator of the common
log-expectile under \eqref{eq:pool-int:test-clt} and
$\hat{\bm\Sigma}_{\Xi,n}$ is the plug-in image of
\eqref{eq:pool-int:test-Sigma-Xi} assembled from the consistent
estimators of \Cref{sec:pool-int:CI},
\begin{equation}
\label{eq:pool-int:test-Sigma-Xi-hat}
  \hat{\bm\Sigma}_{\Xi,n}
  \;=\;
  \mathrm{diag}\!\big((\hat m + L_{j,n})^{2} + 1\big)_{j = 1}^{m}\,
  \hatVc,
  \qquad
  L_{j,n}
  \;=\;
  \log\!\big(k_{j}/(n_{j}(1 - \tau_{n}))\big),
\end{equation}
with $\hat m = \hat m^{(0)}$ the preliminary effective-sample-size
seed \eqref{eq:pool-int:plugin-m} and $\hatVc$ any consistent diagonal
covariance plug-in for \eqref{eq:bg:V_c-diagonal}, as in
\Cref{sec:pool-int:CI}. The residual
$\hat{\bm\Xi}_{n} - \hat\nu_{n}\bm 1$ is the vector of GLS residuals
about $\hat\nu_{n}$; by construction it annihilates the direction
$\bm 1$, which is the structural reason the statistic is insensitive
both to the common centring $\log\expectile{\tau_{n}}$ and --- as the
theorem makes precise --- to any asymptotic bias collinear with
$\bm 1$.

Because \eqref{eq:pool-int:test-stat} is scaled by $k$, the formal
no-separation null behind the chi-square calibration is a
\emph{root-$k$} log-expectile homogeneity condition. If
\[
  \bm\Xi_{n}^{\star}
  \;\coloneqq\;
  \big(\log\expectile{\tau_{n},\,1}, \ldots,
       \log\expectile{\tau_{n},\,m}\big)^{\top}
\]
denotes the vector of population per-sample log-expectiles in a
diagnostic reading beyond the common-marginal model, this condition is
\[
  H_{0,k}^{\xi}\colon\quad
  \sqrt{k}\,\min_{c \in \R}
  \big\|\bm\Xi_{n}^{\star} - c\,\bm 1\big\| \;\to\; 0.
\]
It is automatic under the chapter's standing common-marginal
assumption, for which
$\bm\Xi_{n}^{\star}=\log\expectile{\tau_{n}}\bm 1$. The weaker
first-order condition
$\expectile{\tau_n,j}/\expectile{\tau_n,\ell}\to1$ is useful for
interpreting the target of the diagnostic, but by itself it is not
strong enough to justify a root-$k$ chi-square calibration.

It is useful to keep three objects separate. The \emph{model null} is
the common-marginal/common-tail setup under which
\eqref{eq:pool-int:test-clt} supplies the covariance
$\bm\Sigma_{\Xi}$ and bias vector $\bm\beta$. The
\emph{target-homogeneity null} is the root-$k$ condition
$H_{0,k}^{\xi}$ above. The \emph{central-calibration null} requires, in
addition, that the asymptotic bias be invisible to the GLS residual
projection, namely $\bm\beta\in\mathrm{span}\{\bm 1\}$. The theorem
gives the general noncentral limit; the following corollary extracts
the ordinary central chi-square rejection rule.

\begin{theorem}[Noncentral high-expectile homogeneity diagnostic]
\label{thm:pool-int:hom-test}
Adopt the assumptions of \Cref{thm:pool-int:main}, assume $m\ge2$, and
work on the high-probability logarithmic domain described above. Suppose
$\hat{\bm\Sigma}_{\Xi,n}$ is symmetric positive definite with
probability tending one and
$\hat{\bm\Sigma}_{\Xi,n} \toprob \bm\Sigma_{\Xi}$, with
$\bm\Sigma_{\Xi}$ positive definite. Under the chapter's
common-marginal null model, which implies the common-target centring
$\bm\Xi_{n}^{\star}=\log\expectile{\tau_n}\bm 1$,
the statistic \eqref{eq:pool-int:test-stat} converges to a
noncentral chi-square,
\begin{equation}
\label{eq:pool-int:test-noncentral}
  \Lambda^{\xi}_{n} \;\todist\; \chi^{2}_{m - 1}(\delta),
  \qquad
  \delta
  \;=\;
  \bm\beta^{\top}\bm\Sigma_{\Xi}^{-1}\bm\beta
  \;-\;
  \frac{\big(\bm 1^{\top}\bm\Sigma_{\Xi}^{-1}\bm\beta\big)^{2}}
       {\bm 1^{\top}\bm\Sigma_{\Xi}^{-1}\bm 1}
  \;\ge\; 0,
\end{equation}
with $\bm\beta$, $\bm\Sigma_{\Xi}$ as in
\eqref{eq:pool-int:test-Sigma-Xi}--\eqref{eq:pool-int:test-beta}. The
noncentrality $\delta$ is the squared $\bm\Sigma_{\Xi}^{-1}$-norm of
$\bm\beta$ projected off the direction $\bm 1$, and vanishes ---
so that the limit is the \emph{central} $\chi^{2}_{m - 1}$ ---
exactly when $\bm\beta \in \mathrm{span}\{\bm 1\}$.
Moreover, if the statistic is used diagnostically outside the
common-marginal model in a setting where
$\hat{\bm\Xi}_{n}-\bm\Xi_{n}^{\star}=O_{p}(k^{-1/2})$ and the plug-in
precision matrices $\hat{\bm\Sigma}_{\Xi,n}^{-1}$ remain bounded and
uniformly positive definite in probability, then under any separated
alternative for which the population log-expectiles are separated
faster than $k^{-1/2}$ --- formally,
$\sqrt{k}\,\min_{c \in \R}\big\|(\log\expectile{\tau_{n},\,j})_{j=1}^{m}
- c\,\bm 1\big\| \to \infty$ --- the statistic diverges,
$\Lambda^{\xi}_{n} \toprob \infty$. This alternative-side statement is
only a divergence result. It does not supply a central null calibration
for heterogeneous margins satisfying root-$k$ target equality; such a
calibration would require the corresponding heterogeneous covariance
and bias structure.
\end{theorem}

\begin{proof}
See \Cref{app:proofs:hom-test}.
\end{proof}

\begin{corollary}[Central chi-square calibration]
\label{cor:pool-int:hom-test-central}
Under the assumptions of \Cref{thm:pool-int:hom-test}, the central
chi-square rejection rule
\[
  \Lambda^{\xi}_{n} > \chi^{2}_{m - 1,\, 1 - \alpha},
\]
with $\chi^{2}_{m - 1,\, 1 - \alpha}$ the $(1-\alpha)$-quantile of the
central $\chi^{2}_{m-1}$ law, has asymptotic size $\alpha$ whenever
$\bm\beta\in\mathrm{span}\{\bm 1\}$. In particular this holds in two
standing regimes of the chapter:
\begin{enumerate}[label=(\roman*)]
\item the undersmoothing calibration \eqref{eq:pool-int:undersmoothing},
      under which $\Bc = \Bc^{\star} = \bm 0$ and $b^{\xi/\Quant} = 0$,
      hence $\bm\beta = \bm 0$;
\item the equal-fraction distributed specialisation
      $k_{j}/n_{j} = (k/n)(1 + o(1))$ described in
      \Cref{sec:pool-int:setup}, under which every coordinate of
      $\bm\beta$ collapses to one common value.
\end{enumerate}
If $\delta>0$ under the common-marginal model, the same central
critical value has limiting rejection probability
$\Prob\{\chi^{2}_{m-1}(\delta)>\chi^{2}_{m - 1,\,1-\alpha}\}$, which is
larger than $\alpha$.
\end{corollary}

\begin{proof}
The sufficient conditions in (i)--(ii) are verified in
\Cref{app:proofs:hom-test}. The size statement follows immediately
from \Cref{thm:pool-int:hom-test} and the continuity of the central
chi-square distribution function at its quantiles.
\end{proof}

\paragraph{Discussion.}
Three comments situate the test. First, under the chapter's standing
assumption of a common marginal $F$ the root-$k$ null
$H_{0,k}^{\xi}$ holds automatically, so the test is a specification
check on the pooling hypothesis rather than a test of a substantive
scientific null. A non-rejection is consistent with using the pooling
recipes of \Cref{sec:pool-int:weights,sec:pool-int:CI} at the tested
level, while a rejection warns that the samples should not be pooled
without modelling heterogeneity. Second, the central calibration in
\Cref{cor:pool-int:hom-test-central} is the same variance-dominated
calibration that underlies the confidence interval of
\Cref{cor:pool-int:CI}: case (i) is exactly the undersmoothing
condition \eqref{eq:pool-int:undersmoothing} adopted there, and case
(ii) is the equal-fraction distributed regime in which the
variance-optimal weights of \Cref{prop:pool-int:var-opt-distrib} are
used. Outside both regimes the asymptotic bias $\bm\beta$ is
generically not collinear with $\bm 1$, the limit is the noncentral
$\chi^{2}_{m - 1}(\delta)$ of \eqref{eq:pool-int:test-noncentral},
and the central critical value over-rejects by an amount governed by
$\delta$; this is the expectile-level counterpart of the identical
bias caveat carried, but left implicit, by the tail-index test of
\Cref{thm:bg:tail-homogeneity-test}.

Third, the first-order expectile null
$H_{0,0}^{\xi}\colon
\expectile{\tau_n,j}/\expectile{\tau_n,\ell}\to1$ and the tail-index
null $H_{0}\colon \gamma_{1} = \cdots = \gamma_{m}$ of
\Cref{thm:bg:tail-homogeneity-test} are related but not identical.
Per-sample expectile--quantile equivalence
(\Cref{prop:bg:expectile-quantile}) gives
$\expectile{\tau_{n},\,j} \sim \psifn(\gamma_{j})\,
\Quant_{X_{j}}(\tau_{n})$, so $H_{0,0}^{\xi}$ requires the products
$\psifn(\gamma_{j})\,\Quant_{X_{j}}(\tau_{n})$ to agree across
samples in the limit. The non-injectivity of $\psifn$ on $(0,1)$
should not be read in isolation: under regular variation, different
tail indices force the quantile ratio
$\Quant_{X_j}(\tau)/\Quant_{X_\ell}(\tau)$ to tend to $0$ or
$\infty$, regardless of possible coincidences among the scalar values
$\psifn(\gamma_j)$. Thus fixed alternatives with unequal tail indices
are not hidden by equal $\psifn$ values once the quantile factor is
included. The complementary strength of $\Lambda^{\xi}_{n}$ is instead
scale sensitivity: when the tail indices are equal but the
tail-homoskedasticity condition $(\mathcal H)$ of
\Cref{cond:bg:tail-homoskedasticity} fails, the tail-index test can
fail to reject $H_{0}$ while $H_{0,0}^{\xi}$ fails. In that case the
expectile test registers departures in the actual inferential target
--- the expectile --- to which a test of the tail index alone is by
construction blind. The two diagnostics are therefore complementary:
one targets tail-index equality, the other targets equality of the
high-expectile levels being pooled.
