% !TEX root = ../main.tex
\chapter{Extrapolation to very-extreme expectile levels}
\label{ch:pool-ext}

\section*{Overview}

This chapter extrapolates the pooled intermediate-level expectile
estimator of \Cref{ch:pool-int} to very-extreme levels
$\tau_n^\prime \uparrow 1$ with
$n(1-\tau_n^\prime) \to c \in [0, \infty)$. The argument is deliberately
short: the pooled intermediate result of \Cref{thm:pool-int:main} is
fed into the plug-in extrapolation lemma
\Cref{prop:bg:plug-in-clt}, which is the thesis-level adaptation of
\textcite[Supplement, Proposition~C.6]{DavisonPadoanStupfler2023} proved in
\Cref{app:proofs:plugin-clt}. In the
independent common-margin distributed setting of \Cref{ch:pool-int},
the resulting limit is the first-order expectile analogue of the
pooled Weissman extreme-quantile limit in \Cref{thm:bg:pool-weissman}:
after division by the diverging log-rate, only the pooled Hill
estimator remains visible at first order.

\section{Motivation}
\label{sec:pool-ext:motivation}

The intermediate-level theory of \Cref{ch:pool-int} estimates
$\expectile{\tau_n}$ at levels for which the expected number of
upper-tail observations diverges. This chapter moves to the
very-extreme regime
\[
  \tau_n^\prime \uparrow 1,
  \qquad
  n(1-\tau_n^\prime)\to c\in[0,\infty),
\]
where only finitely many observations are expected beyond the target
level, so the direct intermediate-level asymptotics of \Cref{ch:pool-int}
no longer apply. As recalled in
\Cref{sec:bg:weissman-expectile}, the iid extreme-expectile theory of
\textcite{DaouiaGirardStupfler2018} handles this regime by
extrapolating an intermediate-level expectile estimator with a
Weissman-type power factor. Direct LAWS or direct QB estimation at
$\tau_n^\prime$ is therefore not the object of the chapter: the stable
statistical anchor remains the intermediate level $\tau_n$, and the
very-extreme target is reached by extrapolation.

The pooling protocol is unchanged. Each sample still communicates only
the two primitives inherited from \textcite{DaouiaPadoanStupfler2021}:
the local Hill estimator $\hat\gamma_j(k_j)$ and the local Weissman
quantile input $\hat q_j^\star(\,\cdot\,\mid k_j)$. The asymptotic
covariance matrix, or its plug-in estimate in feasible procedures, is
used to choose weights and standard errors; it is not an additional
expectile-specific primitive. No raw
observations and no expectile-specific scalar, such as a local LAWS
estimate, are introduced at this stage. Consequently the
intermediate-level anchor available for extrapolation is precisely the
canonical pooled QB estimator
\[
  \hatexpectile{\tau_n}^{\mathrm{pool}}
    (\bm\omega_\gamma,\bm\omega_q)
  =
  \psifn\!\big(\hat\gamma_n(\bm\omega_\gamma)\big)
  \hatWei_n(\tau_n\mid\bm\omega_q)
\]
from \Cref{eq:pool-int:expectile-pool}.
This is a consequence of the Paper-3 communication protocol, not a
claim that QB extrapolation is intrinsically preferable to every
possible expectile-specific procedure.

The role of \Cref{prop:bg:plug-in-clt}, the thesis-level plug-in
lemma adapted from \textcite[Supplement, Proposition~C.6]{DavisonPadoanStupfler2023},
is exactly to make this transfer modular. Once an intermediate estimator and a
tail-index estimator satisfy a joint root-rate limit, their
Weissman-extrapolated expectile will have an extreme-level limit
governed only by the tail-index component. Chapter~3 supplies that
joint limit for
$\big(\hat\gamma_n(\bm\omega_\gamma),
\hatexpectile{\tau_n}^{\mathrm{pool}}(\bm\omega_\gamma,\bm\omega_q)\big)$.
The remaining sections of the present chapter therefore do not develop
a new expectile primitive; they specify the extrapolation regime,
define the pooled Weissman-type lift, and read the limiting
distribution and optimal weights from the pooled Hill component.

\section{Extrapolation regime and input limit}
\label{sec:pool-ext:regime}

Throughout this chapter we keep the setup of
\Cref{sec:pool-int:setup}. The number of samples $m$ is fixed; the
samples are independent and have common marginal distribution function
$F$; $n=\sum_{j=1}^{m}n_j$ and $k=\sum_{j=1}^{m}k_j$ with
$n_j\to\infty$, $k_j\to\infty$ and $k_j/n_j\to0$; and the
proportionality conditions \eqref{eq:bg:pool-rates} hold. The common
margin satisfies $\mathcal C_2(\gamma,\rho,A)$, with
$\gamma\in(0,1)$, and obeys the finite first-moment condition
\[
  \E X^{-}<\infty .
\]
The per-sample bias-rate conditions
$\sqrt{k_j}A(n_j/k_j)\to\lambda_j\in\R$ are in force. The
intermediate anchor level satisfies
\[
  L_j(\tau_n)
  =
  \log\!\frac{k_j}{n_j(1-\tau_n)}
  \;\longrightarrow\;
  L_j^\star\in\R,
  \qquad j=1,\ldots,m,
\]
and the moment-negligibility condition is
\begin{equation}
\label{eq:pool-ext:moment-rate}
  \frac{\sqrt{k}}{\Quant_X(\tau_n)}\;\longrightarrow\;0.
\end{equation}
All logarithms of Hill, Weissman and QB quantities are understood on
the high-probability positivity event described in
\Cref{sec:pool-int:estimator}; arbitrary definitions on its complement
do not affect the weak limits. The only strengthening relative to the
intermediate chapter is that the second-order index is now assumed to
satisfy
\begin{equation}
\label{eq:pool-ext:rho-strict}
  \rho<0.
\end{equation}
This strict version is part of the extrapolation theory in
\textcite[Supplement, Proposition~C.6]{DavisonPadoanStupfler2023}, adapted here
as \Cref{prop:bg:plug-in-clt}, and is the same strict second-order
condition that appears in the pooled Weissman theorem
\Cref{thm:bg:pool-weissman}.
All results below are therefore statements in the independent
common-margin distributed specialisation. References to
\Cref{thm:bg:pool-weissman} are analogies within this setting, not
extensions to the full tail-copula framework of \Cref{sec:bg:pooling}.

The new target level is a very-extreme sequence
\begin{equation}
\label{eq:pool-ext:very-extreme-level}
  \tau_n^\prime\uparrow1,
  \qquad
  n(1-\tau_n^\prime)\to c\in[0,\infty),
  \qquad
  \frac{1-\tau_n^\prime}{1-\tau_n}\to0.
\end{equation}
Write the associated Weissman log-rate as
\begin{equation}
\label{eq:pool-ext:ell-def}
  \ell_n
  \;\coloneqq\;
  \log\!\left(\frac{1-\tau_n}{1-\tau_n^\prime}\right).
\end{equation}
Then $\ell_n\to\infty$, and the extrapolation is not allowed to be so
far into the tail that the Weissman normalisation overtakes the
available root-$k$ information:
\begin{equation}
\label{eq:pool-ext:weissman-rate}
  \frac{\sqrt{k}}{\ell_n}\;\longrightarrow\;\infty.
\end{equation}
The finite-log condition $L_j(\tau_n)\to L_j^\star$ already implies
$n(1-\tau_n)\to\infty$, because
$n_j(1-\tau_n)$ has the same order as $k_j$ for every $j$. Hence
\eqref{eq:pool-ext:very-extreme-level} is exactly the pair of regimes
required by \Cref{prop:bg:plug-in-clt}: $\tau_n$ is intermediate and
$\tau_n^\prime$ is very-extreme.
Moreover, if $\alpha_j=\lim n_j/n\in(0,1)$, then the proportionality
conditions give
\[
  \alpha_j
  =
  \frac{b_j^{-1}}{\sum_{i=1}^{m}b_i^{-1}},
  \qquad
  n_j(1-\tau_n^\prime)\to \alpha_jc.
\]
The very-extreme level is therefore global, but it has the expected
local interpretation under the proportional-growth assumptions.

The logarithmic rate used here is also aligned with the pooled
Weissman rate in \Cref{thm:bg:pool-weissman}. The finite-log condition
and proportionality assumptions imply
\[
  \log\!\frac{k}{n(1-\tau_n)} = O(1),
\]
and hence
\begin{equation}
\label{eq:pool-ext:rate-alignment}
  \log\!\frac{k}{n(1-\tau_n^\prime)}
  =
  \ell_n
  +
  \log\!\frac{k}{n(1-\tau_n)}
  =
  \ell_n + O(1).
\end{equation}
Since $\ell_n\to\infty$, the two logarithmic normalisers are
asymptotically equivalent on the very-extreme scale.

For the application of \Cref{prop:bg:plug-in-clt} we take
\[
  a_n=\sqrt{k},\qquad
  \hat\gamma_n=\hat\gamma_n(\bm\omega_\gamma),\qquad
  \widehat\xi_{\tau_n}^{\,I}
  =
  \hatexpectile{\tau_n}^{\mathrm{pool}}
    (\bm\omega_\gamma,\bm\omega_q).
\]
The non-Gaussian hypotheses of the proposition are then verified as
follows. Since $k\to\infty$, $a_n\to\infty$. The preceding paragraph
gives $n(1-\tau_n)\to\infty$, while
\eqref{eq:pool-ext:very-extreme-level} gives
$n(1-\tau_n^\prime)\to c$ and
$(1-\tau_n^\prime)/(1-\tau_n)\to0$. The Weissman-rate condition is
\eqref{eq:pool-ext:weissman-rate}. The second-order limit
\[
  a_n A((1-\tau_n)^{-1})
  =
  \sqrt{k}\,A((1-\tau_n)^{-1})
  \;\longrightarrow\;
  \Lambda^\bullet\in\R
\]
is \Cref{eq:pool-int:Lambda-bullet}, and
\eqref{eq:pool-ext:moment-rate} gives
$a_n/\Quant_X(\tau_n)\to0$. Finally,
$\Prob(\widehat\xi_{\tau_n}^{\,I}>0)\to1$ because
$\hatWei_n(\tau_n\mid\bm\omega_q)>0$ on the local Weissman positivity
event and $\psifn(\hat\gamma_n(\bm\omega_\gamma))>0$ on the event on
which the pooled Hill estimator lies in $(0,1)$.

\begin{lemma}[Deterministic very-extreme ratio control]
\label{lem:pool-ext:det-ratio}
Under the tail and rate assumptions of this section,
\[
  \frac{\sqrt{k}}{\ell_n}
  \left\{
    \log\expectile{\tau_n^\prime}
    - \log\expectile{\tau_n}
    - \gamma\ell_n
  \right\}
  \;\longrightarrow\;0.
\]
This lemma records, in the present notation, the deterministic
population step behind the finite-mean plug-in extrapolation result of
\textcite[Supplement, Proposition~C.6]{DavisonPadoanStupfler2023}. The proof of
that proposition reduces the argument to the same log-decomposition as
\textcite[Theorem~5]{DaouiaGirardStupfler2020}; the control is used
here as a source-paper ingredient, not as a new contribution of the
thesis.
\end{lemma}

\begin{proof}
See \Cref{app:proofs:pool-ext-det-ratio}.
\end{proof}

It remains to record the exact intermediate Gaussian input that will
be fed into the plug-in extrapolation lemma. Put
\[
  \bm D^\star \coloneqq \mathrm{diag}(\bm L^\star),
  \qquad
  \bm L^\star=(L_1^\star,\ldots,L_m^\star)^\top .
\]

\begin{proposition}[Joint input for the extrapolation step]
\label{prop:pool-ext:input-clt}
Under the assumptions stated above, for any deterministic affine weight vectors
$\bm\omega_\gamma,\bm\omega_q\in\R^m$ satisfying
$\bm\omega_\gamma^\top\bm 1=\bm\omega_q^\top\bm 1=1$,
\begin{equation}
\label{eq:pool-ext:input-clt}
  \sqrt{k}
  \begin{pmatrix}
    \hat\gamma_n(\bm\omega_\gamma)-\gamma \\[2pt]
    \hatexpectile{\tau_n}^{\mathrm{pool}}
      (\bm\omega_\gamma,\bm\omega_q)/\expectile{\tau_n}-1
  \end{pmatrix}
  \;\todist\;
  \mathcal N\!\left(
  \begin{pmatrix}
    \bm\omega_\gamma^\top\Bc \\[2pt]
    \mu^\xi(\bm\omega_\gamma,\bm\omega_q)
  \end{pmatrix},
  \begin{pmatrix}
    \bm\omega_\gamma^\top\Vc\bm\omega_\gamma
      & s_{\gamma,\xi}(\bm\omega_\gamma,\bm\omega_q) \\[2pt]
    s_{\gamma,\xi}(\bm\omega_\gamma,\bm\omega_q)
      & \sigma^\xi(\bm\omega_\gamma,\bm\omega_q)^2
  \end{pmatrix}
  \right),
\end{equation}
where $\mu^\xi$ and $\sigma^\xi$ are given in
\eqref{eq:pool-int:mu-xi} and \eqref{eq:pool-int:sigma-xi}, and
\begin{equation}
\label{eq:pool-ext:gamma-xi-cross}
  s_{\gamma,\xi}(\bm\omega_\gamma,\bm\omega_q)
  =
  m(\gamma)\,\bm\omega_\gamma^\top\Vc\bm\omega_\gamma
  +
  \bm\omega_\gamma^\top\bm D^\star\Vc\bm\omega_q .
\end{equation}
The same convergence holds with random weights
$(\hat{\bm\omega}_\gamma,\hat{\bm\omega}_q)$ satisfying
\[
  \hat{\bm\omega}_\gamma^\top\bm 1
  =
  \hat{\bm\omega}_q^\top\bm 1
  =
  1
  \quad\text{exactly,}
  \qquad
  \hat{\bm\omega}_\gamma\toprob\bm\omega_\gamma,
  \qquad
  \hat{\bm\omega}_q\toprob\bm\omega_q .
\]
In this estimated-weight version, the mean and covariance on the
right-hand side of \eqref{eq:pool-ext:input-clt} are evaluated at the
deterministic probability limits
$(\bm\omega_\gamma,\bm\omega_q)$.
\end{proposition}

\begin{proof}
See \Cref{app:proofs:pool-ext-input-clt}.
\end{proof}

\Cref{prop:pool-ext:input-clt} is more information than the final
very-extreme theorem will visibly retain. The extrapolated limit uses
only the first component, the pooled Hill estimator, because
\Cref{prop:bg:plug-in-clt} divides by the diverging log-rate
$\ell_n$. The joint statement is nevertheless the right object to
state here: it is exactly the hypothesis of the plug-in theorem, with
\[
  \Gamma\sim
  \mathcal N\!\big(\bm\omega_\gamma^\top\Bc,\,
                   \bm\omega_\gamma^\top\Vc\bm\omega_\gamma\big),
  \qquad
  \Delta
  \text{ equal to the second component of }
  \eqref{eq:pool-ext:input-clt}.
\]
Together with the hypothesis verification above, this is precisely the
input required by \Cref{prop:bg:plug-in-clt}. For clarity, we collect
the full hypothesis map before defining the very-extreme estimator.

\paragraph{Checklist for the plug-in extrapolation step.}
In the application of \Cref{prop:bg:plug-in-clt} to
\Cref{thm:pool-ext:main}, the generic quantities are
\[
  a_n=\sqrt{k},\qquad
  \hat\gamma_n=\hat\gamma_n(\bm\omega_\gamma),\qquad
  \widehat\xi_{\tau_n}^{\,I}
  =
  \hatexpectile{\tau_n}^{\mathrm{pool}}
    (\bm\omega_\gamma,\bm\omega_q).
\]
The required conditions are satisfied as follows.
\begin{enumerate}
\item \emph{Intermediate level.}
The finite-log condition
$L_j(\tau_n)\to L_j^\star$ and the proportional-growth assumptions
imply $n(1-\tau_n)\to\infty$, so $\tau_n$ is intermediate at the
global scale required by \Cref{prop:bg:plug-in-clt}.
\item \emph{Extreme level.}
\Cref{eq:pool-ext:very-extreme-level} gives
$\tau_n^\prime\uparrow1$, $n(1-\tau_n^\prime)\to c\in[0,\infty)$ and
$(1-\tau_n^\prime)/(1-\tau_n)\to0$; hence
$\ell_n=\log\{(1-\tau_n)/(1-\tau_n^\prime)\}\to\infty$.
\item \emph{Log-ratio condition.}
The rate condition \eqref{eq:pool-ext:weissman-rate} is exactly
$a_n/\ell_n\to\infty$. The deterministic population remainder
\[
  \frac{a_n}{\ell_n}
  \{\log\expectile{\tau_n^\prime}
     -\log\expectile{\tau_n}-\gamma\ell_n\}
  \to0
\]
is \Cref{lem:pool-ext:det-ratio}.
\item \emph{Second-order condition.}
The margin satisfies $\mathcal C_2(\gamma,\rho,A)$ and
\eqref{eq:pool-ext:rho-strict} imposes $\rho<0$, the strict
second-order condition used by the source extrapolation argument.
Moreover \Cref{eq:pool-int:Lambda-bullet} gives the finite
normalisation
$a_nA((1-\tau_n)^{-1})=\sqrt{k}A((1-\tau_n)^{-1})
\to\Lambda^\bullet\in\R$.
\item \emph{Finite-mean condition.}
The chapter setup imposes $\gamma\in(0,1)$ and
$\E X^{-}<\infty$, so the QB expectile--quantile bridge is used only
in the finite-mean range.
\item \emph{Input Gaussian limit.}
\Cref{prop:pool-ext:input-clt} is exactly the joint input
\eqref{eq:bg:joint-input}. Its first component has limiting variable
\[
  \Gamma\sim
  \mathcal N\!\big(\bm\omega_\gamma^\top\Bc,\,
                   \bm\omega_\gamma^\top\Vc\bm\omega_\gamma\big),
\]
and its second component is the corresponding intermediate pooled-QB
limit.
\item \emph{Bias normalisation.}
\Cref{eq:pool-ext:moment-rate} gives
$a_n/\Quant_X(\tau_n)=\sqrt{k}/\Quant_X(\tau_n)\to0$, so the
lower-tail moment term in the expectile--quantile expansion does not
create an additional $\lambda_2$ bias. The remaining intermediate
expectile bias is finite on the $\sqrt{k}$ scale through
\Cref{eq:pool-int:b-xi-Q-def,eq:pool-int:mu-xi}; after the plug-in
normalisation by $\sqrt{k}/\ell_n$, this second-coordinate bias is
divided by the diverging $\ell_n$ and disappears. The only bias left
in \Cref{thm:pool-ext:main} is therefore the pooled-Hill bias
$\bm\omega_\gamma^\top\Bc$ carried by $\Gamma$.
\end{enumerate}

\section{The pooled extrapolated expectile estimator}
\label{sec:pool-ext:estimator}

The extrapolated estimator is obtained by applying the Weissman-type
lift of \Cref{eq:bg:weissman-expectile} to the pooled intermediate
anchor of \Cref{eq:pool-int:expectile-pool}. For affine weights
$\bm\omega_{\gamma},\bm\omega_q\in\R^m$ satisfying
$\bm\omega_\gamma^\top\bm 1=\bm\omega_q^\top\bm 1=1$, define
\begin{equation}
\label{eq:pool-ext:estimator}
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
    (\bm\omega_\gamma,\bm\omega_q)
  \;\coloneqq\;
  \left(\frac{1-\tau_n^\prime}{1-\tau_n}\right)^{
    -\hat\gamma_n(\bm\omega_\gamma)}
  \hatexpectile{\tau_n}^{\mathrm{pool}}
    (\bm\omega_\gamma,\bm\omega_q).
\end{equation}
Although $\bm\omega_q$ affects the finite-sample anchor and the
intermediate-level component of the joint input limit, it disappears
from the first-order very-extreme CLT after normalisation by
$\sqrt{k}/\ell_n$. It remains part of the definition because different
choices of $\bm\omega_q$ give different finite-sample estimators and
higher-order terms.
Equivalently, using \Cref{eq:pool-int:expectile-pool},
\[
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
    (\bm\omega_\gamma,\bm\omega_q)
  =
  \exp\{\ell_n\hat\gamma_n(\bm\omega_\gamma)\}\,
  \psifn\!\big(\hat\gamma_n(\bm\omega_\gamma)\big)\,
  \hatWei_n(\tau_n\mid\bm\omega_q),
\]
where $\ell_n$ is the log-rate in \eqref{eq:pool-ext:ell-def}. Thus
the same two weight vectors used at the intermediate level remain
visible: $\bm\omega_\gamma$ acts on the pooled Hill input, both in the
QB factor $\psifn(\hat\gamma_n)$ and in the extrapolation power, while
$\bm\omega_q$ acts only on the weighted-geometric quantile input that
forms the intermediate anchor. As in \Cref{sec:pool-int:estimator},
all log-scale and real-power expressions are read on the
high-probability positivity event for the local Weissman factors and
on the event where the pooled Hill estimator entering $\psifn$ lies in
$(0,1)$. This event also gives
$\Prob\{\hatexpectile{\tau_n}^{\mathrm{pool}}
(\bm\omega_\gamma,\bm\omega_q)>0\}\to1$, the positivity hypothesis in
\Cref{prop:bg:plug-in-clt}; arbitrary definitions on the complementary
event do not alter the weak limits.

The construction targets the same expectile level $\tau_n^\prime$ that
appears in the population quantity $\expectile{\tau_n^\prime}$. No
adjusted quantile level is introduced. The reason is that the
expectile--quantile bridge of \Cref{prop:bg:expectile-quantile} is a
same-level equivalence,
$\expectile{\tau}/\Quant_X(\tau)\to\psifn(\gamma)$, and the
Weissman step in \Cref{prop:bg:plug-in-clt} extrapolates the
intermediate expectile itself from $\tau_n$ to $\tau_n^\prime$.

At matched weights, \eqref{eq:pool-ext:estimator} is exactly the
QB plug-in applied to the pooled Weissman quantile at the
very-extreme level. Indeed, by the local Weissman identity,
\[
  \hat q_j^\star(\tau_n^\prime\mid k_j)
  =
  \left(\frac{1-\tau_n^\prime}{1-\tau_n}\right)^{
    -\hat\gamma_j(k_j)}
  \hat q_j^\star(\tau_n\mid k_j),
\]
and therefore, for every affine $\bm\omega$,
\begin{equation}
\label{eq:pool-ext:matched-weissman-identity}
  \hatWei_n(\tau_n^\prime\mid\bm\omega)
  =
  \left(\frac{1-\tau_n^\prime}{1-\tau_n}\right)^{
    -\hat\gamma_n(\bm\omega)}
  \hatWei_n(\tau_n\mid\bm\omega).
\end{equation}
Combining \eqref{eq:pool-ext:matched-weissman-identity} with
\eqref{eq:pool-ext:estimator} gives the exact matched-weight identity
\begin{equation}
\label{eq:pool-ext:matched-expectile-identity}
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
    (\bm\omega,\bm\omega)
  =
  \psifn\!\big(\hat\gamma_n(\bm\omega)\big)\,
  \hatWei_n(\tau_n^\prime\mid\bm\omega).
\end{equation}
Thus, on the matched diagonal, the Chapter~4 estimator is not a new
third primitive: it is the same pooled-Hill / pooled-Weissman QB
pipeline as in Chapter~3, evaluated after the Weissman quantile input
has been extrapolated to $\tau_n^\prime$.

\begin{remark}[Matched local-QB equivalence]
\label{rem:pool-ext:matched-local-qb}
There is also a local-QB analogue of the construction, parallel to
\Cref{eq:pool-int:expectile-B}. Define the marginal extrapolated QB
expectile on sample $j$ by
\[
  \hatexpectile{\tau_n^\prime,j}^{\star}
  \;\coloneqq\;
  \left(\frac{1-\tau_n^\prime}{1-\tau_n}\right)^{
    -\hat\gamma_j(k_j)}
  \psifn\!\big(\hat\gamma_j(k_j)\big)\,
  \hat q_j^\star(\tau_n\mid k_j)
  =
  \psifn\!\big(\hat\gamma_j(k_j)\big)\,
  \hat q_j^\star(\tau_n^\prime\mid k_j),
\]
and geometrically pool these $m$ scalars:
\begin{equation}
\label{eq:pool-ext:local-qb-geometric}
  \hatexpectile{\tau_n^\prime}^{\mathrm{loc},\star}(\bm\omega)
  \;\coloneqq\;
  \prod_{j=1}^{m}
    \big[\hatexpectile{\tau_n^\prime,j}^{\star}\big]^{\omega_j}.
\end{equation}
The comparison with \eqref{eq:pool-ext:matched-expectile-identity} is
identical to \Cref{prop:pool-int:A-B-equiv}, because the
very-extreme Weissman factor is common to both estimators. More
precisely,
\[
  \log\!\frac{
    \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
      (\bm\omega,\bm\omega)}
    {\hatexpectile{\tau_n^\prime}^{\mathrm{loc},\star}(\bm\omega)}
  =
  \log\psifn\!\big(\hat\gamma_n(\bm\omega)\big)
  -
  \sum_{j=1}^{m}\omega_j
    \log\psifn\!\big(\hat\gamma_j(k_j)\big),
\]
so the unscaled log-ratio is $O_p(k^{-1})$ and in particular is
negligible both on the intermediate $\sqrt{k}$ scale and on the
extreme $\sqrt{k}/\ell_n$ scale. The canonical estimator
\eqref{eq:pool-ext:estimator} is therefore consistent with the
local-QB interpretation whenever the weights are matched.
\end{remark}

This matched-weight qualification is important. If
$\bm\omega_\gamma\ne\bm\omega_q$, the alternative estimator
\[
  \psifn\!\big(\hat\gamma_n(\bm\omega_\gamma)\big)\,
  \hatWei_n(\tau_n^\prime\mid\bm\omega_q)
\]
is generally first-order different from the object analysed in this
chapter; it is not an equivalent rewrite of
\eqref{eq:pool-ext:estimator}. By
\eqref{eq:pool-ext:matched-weissman-identity}, it replaces the
extrapolation power $\hat\gamma_n(\bm\omega_\gamma)$ by
$\hat\gamma_n(\bm\omega_q)$, and their log-ratio is
\[
  \ell_n\{
    \hat\gamma_n(\bm\omega_q)-\hat\gamma_n(\bm\omega_\gamma)
  \}.
\]
After multiplication by the extreme normalisation
$\sqrt{k}/\ell_n$, this term contributes
$\sqrt{k}\{\hat\gamma_n(\bm\omega_q)-\hat\gamma_n(\bm\omega_\gamma)\}$,
which is generally non-negligible. Hence the two-weight estimator in
\eqref{eq:pool-ext:estimator} is the object to which the plug-in CLT
will be applied; replacing its extrapolation exponent changes the
first-order extreme limit unless the two Hill weightings agree
asymptotically.

\section{Main extrapolation theorem}
\label{sec:pool-ext:CLT}

We can now state the chapter's main limit theorem. It is a direct
application of the plug-in extrapolation result
\Cref{prop:bg:plug-in-clt} to the pooled intermediate input of
\Cref{prop:pool-ext:input-clt}. The theorem is intentionally
short: all expectile-specific work has already been done at the
intermediate level, and the very-extreme scale keeps only the
pooled-Hill component.

\begin{theorem}[Pooled Weissman extrapolation for extreme expectiles]
\label{thm:pool-ext:main}
Assume the setup and rate conditions of \Cref{sec:pool-ext:regime};
in particular, $\rho<0$, $\sqrt{k}/\Quant_X(\tau_n)\to0$,
$\ell_n\to\infty$ and $\sqrt{k}/\ell_n\to\infty$. Use the
positivity and log-domain convention of
\Cref{sec:pool-ext:estimator}, and the deterministic ratio control
verified in \Cref{lem:pool-ext:det-ratio}.
Let $\bm\omega_\gamma,\bm\omega_q\in\R^m$ be deterministic affine
weight vectors satisfying
$\bm\omega_\gamma^\top\bm 1=\bm\omega_q^\top\bm 1=1$, and define
$\hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
(\bm\omega_\gamma,\bm\omega_q)$ by
\eqref{eq:pool-ext:estimator}. Then
\begin{equation}
\label{eq:pool-ext:main-clt}
  \frac{\sqrt{k}}{\ell_n}
  \left(
    \frac{\hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
      (\bm\omega_\gamma,\bm\omega_q)}
         {\expectile{\tau_n^\prime}}
    - 1
  \right)
  \;\todist\;
  \mathcal N\!\left(
    \bm\omega_\gamma^\top\Bc,\;
    \bm\omega_\gamma^\top\Vc\bm\omega_\gamma
  \right),
\end{equation}
where $\ell_n=\log\{(1-\tau_n)/(1-\tau_n^\prime)\}$.

The same conclusion holds with estimated weights
$(\hat{\bm\omega}_\gamma,\hat{\bm\omega}_q)$, provided
\[
  \hat{\bm\omega}_\gamma^\top\bm 1
  =
  \hat{\bm\omega}_q^\top\bm 1
  =
  1
  \quad\text{exactly,}
  \qquad
  \hat{\bm\omega}_\gamma\toprob\bm\omega_\gamma,
  \qquad
  \hat{\bm\omega}_q\toprob\bm\omega_q .
\]
\end{theorem}

\begin{proof}
See \Cref{app:proofs:pool-ext-main}.
\end{proof}

\Cref{thm:pool-ext:main} is the first-order expectile analogue of
\Cref{thm:bg:pool-weissman} in the independent common-margin
specialisation used in the contribution chapters. In both cases,
extrapolation to a very-extreme level multiplies the tail-index
estimation error by a diverging logarithmic distance, while the
intermediate scale estimate is too small to survive the normalisation.
The only role of the expectile construction is to provide a valid
positive intermediate anchor and to verify the deterministic
expectile-ratio control needed by \Cref{prop:bg:plug-in-clt}. The QB
derivative $m(\gamma)$ and the intermediate expectile-bias terms
$\Bc^\star$ and $b^{\xi/\Quant}$ are not zero in general; they are
lower order on the very-extreme normalisation because they enter only
through the intermediate anchor.

\section{First-order optimal Hill weights for extrapolation}
\label{sec:pool-ext:weights}

The limiting law in \Cref{thm:pool-ext:main} has the same
optimal-weight structure as the pooled Hill and pooled Weissman limits of
\Cref{thm:bg:pool-hill,thm:bg:pool-weissman}. Indeed, after the
normalisation by $\sqrt{k}/\ell_n$, the intermediate anchor
\[
  \hatexpectile{\tau_n}^{\mathrm{pool}}
    (\bm\omega_\gamma,\bm\omega_q)
\]
contributes only an $o_p(1)$ term. The first-order mean and variance
are therefore
\begin{equation}
\label{eq:pool-ext:mean-var}
  \mu^{\xi,\star}(\bm\omega_\gamma)
  =
  \bm\omega_\gamma^\top\Bc,
  \qquad
  \sigma^{\xi,\star}(\bm\omega_\gamma)^2
  =
  \bm\omega_\gamma^\top\Vc\bm\omega_\gamma .
\end{equation}
This is the same first-order quadratic structure as in
\Cref{prop:bg:optimal-weights}, and it is also the structure that
\textcite[Theorem~2]{DaouiaPadoanStupfler2021} obtain for the
weighted-geometric pooled Weissman quantile. Within the independent
common-margin setting of this chapter, the only difference is
interpretive: here the extrapolated target is the common expectile
$\expectile{\tau_n^\prime}$ rather than the common extreme quantile.
The first-order criterion therefore identifies only the extrapolation
exponent, that is, the Hill-weight vector $\bm\omega_\gamma$; it does
not identify the anchor weight $\bm\omega_q$.

As in \Cref{prop:bg:optimal-weights}, weights are real affine weights;
no non-negativity constraint is imposed. If convex weights are required,
the solution is the corresponding constrained quadratic programme and
need not equal the closed forms below.

\begin{proposition}[First-order optimal Hill weights for extrapolated expectiles]
\label{prop:pool-ext:optimal-weights}
Assume the hypotheses of \Cref{thm:pool-ext:main}, and suppose that
$\Vc$ is positive definite.
Among affine Hill weights $\bm\omega_\gamma^\top\bm 1=1$, the
variance-optimal extrapolation weight is
\begin{equation}
\label{eq:pool-ext:var-opt}
  \bm\omega_\gamma^{\xi,\mathrm{Var},\star}
  =
  \bm\omega^{\mathrm{Var}}
  =
  \frac{\Vc^{-1}\bm 1}{\bm 1^\top\Vc^{-1}\bm 1},
\end{equation}
and the minimum first-order variance is
\begin{equation}
\label{eq:pool-ext:var-opt-value}
  \sigma^{\xi,\star}
    (\bm\omega_\gamma^{\xi,\mathrm{Var},\star})^2
  =
  \frac{1}{\bm 1^\top\Vc^{-1}\bm 1}.
\end{equation}
If the limiting bias is retained, the first-order AMSE criterion is
\begin{equation}
\label{eq:pool-ext:amse-def}
  \mathrm{AMSE}^{\xi,\star}(\bm\omega_\gamma)
  =
  \big(\bm\omega_\gamma^\top\Bc\big)^2
  +
  \bm\omega_\gamma^\top\Vc\bm\omega_\gamma,
  \qquad
  \bm\omega_\gamma^\top\bm 1=1,
\end{equation}
whose unique minimiser is
\begin{equation}
\label{eq:pool-ext:amse-opt}
  \bm\omega_\gamma^{\xi,\mathrm{AMSE},\star}
  =
  \bm\omega^{\mathrm{AMSE}}
  =
  \frac{(1 + \Bc^\top\Vc^{-1}\Bc)\,\Vc^{-1}\bm 1
        - (\bm 1^\top\Vc^{-1}\Bc)\,\Vc^{-1}\Bc}
       {(1 + \Bc^\top\Vc^{-1}\Bc)(\bm 1^\top\Vc^{-1}\bm 1)
        - (\bm 1^\top\Vc^{-1}\Bc)^2}.
\end{equation}
The corresponding minimum AMSE is
\begin{equation}
\label{eq:pool-ext:amse-opt-value}
  \mathrm{AMSE}^{\xi,\star}
    (\bm\omega_\gamma^{\xi,\mathrm{AMSE},\star})
  =
  \frac{1+\Bc^\top\Vc^{-1}\Bc}
       {(1 + \Bc^\top\Vc^{-1}\Bc)(\bm 1^\top\Vc^{-1}\bm 1)
        - (\bm 1^\top\Vc^{-1}\Bc)^2}.
\end{equation}
The denominator in \eqref{eq:pool-ext:amse-opt}--\eqref{eq:pool-ext:amse-opt-value}
is strictly positive.
The quantile-anchor weight $\bm\omega_q$ is not identified by either
first-order criterion: for the variance objective, every pair
$(\bm\omega^{\mathrm{Var}},\bm\nu)$ with
$\bm\nu^\top\bm 1=1$ has the same first-order variance, and for the
AMSE objective the same is true of every pair
$(\bm\omega^{\mathrm{AMSE}},\bm\nu)$.
\end{proposition}

\begin{proof}
See \Cref{app:proofs:pool-ext-optimal-weights}.
\end{proof}

In the independent common-distribution distributed setting used
throughout the contribution chapters, $\Rjl\equiv0$ for $j\ne\ell$
and \Cref{eq:bg:V_c-diagonal} gives a diagonal covariance matrix.
Consequently \eqref{eq:pool-ext:var-opt} reduces to
\begin{equation}
\label{eq:pool-ext:var-opt-distrib}
  \omega_{\gamma,j}^{\xi,\mathrm{Var},\star}
  =
  \left(\sum_{i=1}^{m}c_i^{-1}\right)^{-1}c_j^{-1}
  \;\sim\;
  \frac{k_j}{k},
  \qquad j=1,\ldots,m.
\end{equation}
Here $c_j=\lim k_1/k_j$, so $c_j^{-1}$ is proportional to the
limiting effective sample size $k_j$. Thus, at very-extreme expectile
levels, the variance-optimal distributed rule is again
effective-sample-size weighting. This is the same simplification as in
the pooled extreme-quantile case of \Cref{eq:bg:V-opt-distrib}. If the
second-order bias vector is
collinear with $\bm 1$ --- in particular under the undersmoothing
calibration that makes $\Bc=\bm0$, or under the
equal-effective-fraction common-distribution regime, where the
aggregate-scale entries
$\sqrt{c_jS}\lambda_j/(1-\rho)$ of $\Bc$, with
$S=\sum_{i=1}^{m}c_i^{-1}$, are sample-free --- then
$\bm\omega_\gamma^\top\Bc$ is constant on the affine constraint set.
In that case the AMSE criterion
\eqref{eq:pool-ext:amse-def} differs from the variance criterion only
by an additive constant, so
$\bm\omega_\gamma^{\xi,\mathrm{AMSE},\star}
=\bm\omega_\gamma^{\xi,\mathrm{Var},\star}$.

\Cref{prop:pool-ext:optimal-weights} also explains why the
very-extreme chapter differs from \Cref{sec:pool-int:weights}. At
intermediate levels the delta method through the QB map retains the
full pair of primitive contributions, and the expectile variance and
AMSE involve
$m(\gamma)$, $\bm L^\star$, $\Bc^\star$ and $b^{\xi/\Quant}$; see
\eqref{eq:pool-int:sigma-xi}--\eqref{eq:pool-int:mu-xi}. At
very-extreme levels those quantities live in the intermediate
component $\Delta$ of \Cref{prop:bg:plug-in-clt}. The division by the
diverging extrapolation log-rate $\ell_n$ removes that component from
the first-order law, leaving only the pooled-Hill mean and variance in
\eqref{eq:pool-ext:mean-var}. Hence, outside the collapse regimes in
which both AMSE criteria reduce to the variance criterion, the
Chapter~4 AMSE weights need not be the Chapter~3 expectile AMSE
weights; they are the Daouia--Padoan--Stupfler Hill, and therefore
extreme-quantile, weights read through the expectile extrapolation
theorem.

The non-identification of $\bm\omega_q$ is only a first-order
statement. The anchor weight still changes the finite-sample point
estimator in \eqref{eq:pool-ext:estimator}, and it may affect
higher-order accuracy through the intermediate estimator. A natural
canonical representative is therefore the matched choice
\[
  \bm\omega_q=\bm\omega_\gamma,
\]
with $\bm\omega_\gamma$ chosen as either
\eqref{eq:pool-ext:var-opt} or \eqref{eq:pool-ext:amse-opt}. This
is an implementation convention rather than a consequence of the
first-order optimality criterion; a higher-order or finite-sample
criterion would be needed to rank different affine choices of
$\bm\omega_q$. The matched convention keeps the extrapolated expectile
in the same weighted-geometric Weissman form as
\Cref{eq:pool-ext:matched-expectile-identity}: for matched weights,
the estimator is exactly
$\psifn(\hat\gamma_n(\bm\omega_\gamma))\,
\hatWei_n(\tau_n^\prime\mid\bm\omega_\gamma)$. It also gives a single
default implementation for the confidence intervals of
\Cref{sec:pool-ext:CI}, whose first-order standard errors depend only
on $\bm\omega_\gamma$.

\section{Bias-reduction}
\label{sec:pool-ext:bias-reduction}

The limit in \Cref{thm:pool-ext:main} retains, on the
$\sqrt{k}/\ell_n$ scale, the same asymptotic bias as the pooled Hill
estimator:
\[
  \mu^{\xi,\star}(\bm\omega_\gamma)
  =
  \bm\omega_\gamma^\top\Bc .
\]
This is the only first-order bias visible after extrapolation to
$\tau_n^\prime$. The intermediate QB bias terms that were central in
\Cref{ch:pool-int} enter the second coordinate of the joint input in
\Cref{prop:pool-ext:input-clt}; \Cref{prop:bg:plug-in-clt} removes
that coordinate by dividing by the diverging log-rate $\ell_n$.
Consequently, bias reduction at the very-extreme level is inherited
directly from the tail-index correction of
\textcite[\S2.2]{DaouiaPadoanStupfler2021}.

Assume that a consistent estimator of the limiting Hill-bias vector is
available,
\begin{equation}
\label{eq:pool-ext:Bc-hat-consistency}
  \hatBc \toprob \Bc .
\end{equation}
The following result is therefore conditional on a feasible
second-order bias-estimation step; under the abstract condition
$\mathcal C_2(\gamma,\rho,A)$ alone,
\eqref{eq:pool-ext:Bc-hat-consistency} is not automatic.
For any affine Hill-weight vector $\bm\omega_\gamma^\top\bm 1=1$,
define the bias-reduced pooled tail-index estimator
\begin{equation}
\label{eq:pool-ext:bias-reduced-gamma}
  \overline\gamma_n(\bm\omega_\gamma)
  \;\coloneqq\;
  \hat\gamma_n(\bm\omega_\gamma)
  -
  \frac{\bm\omega_\gamma^\top\hatBc}{\sqrt{k}} .
\end{equation}
The associated bias-reduced extrapolated expectile keeps the same
intermediate anchor as \eqref{eq:pool-ext:estimator}, but replaces the
extrapolation exponent by \eqref{eq:pool-ext:bias-reduced-gamma}:
\begin{equation}
\label{eq:pool-ext:bias-reduced-estimator}
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star,\mathrm{bc}}
    (\bm\omega_\gamma,\bm\omega_q)
  \;\coloneqq\;
  \left(\frac{1-\tau_n^\prime}{1-\tau_n}\right)^{
    -\overline\gamma_n(\bm\omega_\gamma)}
  \hatexpectile{\tau_n}^{\mathrm{pool}}
    (\bm\omega_\gamma,\bm\omega_q).
\end{equation}
For feasible estimated weights satisfying the exact affine constraints
we use the corresponding plug-in convention
\[
  \overline\gamma_n(\hat{\bm\omega}_\gamma)
  =
  \hat\gamma_n(\hat{\bm\omega}_\gamma)
  -
  \frac{\hat{\bm\omega}_\gamma^\top\hatBc}{\sqrt{k}},
\]
and define
\[
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star,\mathrm{bc}}
    (\hat{\bm\omega}_\gamma,\hat{\bm\omega}_q)
  =
  \left(\frac{1-\tau_n^\prime}{1-\tau_n}\right)^{
    -\overline\gamma_n(\hat{\bm\omega}_\gamma)}
  \hatexpectile{\tau_n}^{\mathrm{pool}}
    (\hat{\bm\omega}_\gamma,\hat{\bm\omega}_q).
\]
Only the exponent is corrected because only the exponent contributes at
the $\sqrt{k}/\ell_n$ scale. Replacing the occurrence of
$\hat\gamma_n(\bm\omega_\gamma)$ inside the intermediate QB factor
$\psifn(\hat\gamma_n(\bm\omega_\gamma))$ by
$\overline\gamma_n(\bm\omega_\gamma)$, on the same high-probability
compact event used for Taylor expansion of $\log\psifn$, would change
\eqref{eq:pool-ext:bias-reduced-estimator} by a log-ratio of order
$O_p(k^{-1/2})$, hence by $o_p(1)$ after multiplication by
$\sqrt{k}/\ell_n$. The minimal correction above is therefore enough
for the first-order very-extreme theorem.
The correction introduces no new log-domain restriction: the
extrapolation base is deterministic and positive, so the only random
positivity requirement remains that of the intermediate anchor.

\begin{corollary}[Bias-reduced extrapolated expectile CLT]
\label{cor:pool-ext:bias-reduced-clt}
Assume the hypotheses of \Cref{thm:pool-ext:main}, and assume
\eqref{eq:pool-ext:Bc-hat-consistency}; in particular
$\ell_n\to\infty$ and $\sqrt{k}/\ell_n\to\infty$, and the positivity
convention of \Cref{sec:pool-ext:estimator} is in force. Let
$\bm\omega_\gamma,\bm\omega_q\in\R^m$ be deterministic affine weight
vectors satisfying
$\bm\omega_\gamma^\top\bm 1=\bm\omega_q^\top\bm 1=1$.
Then the estimator
\eqref{eq:pool-ext:bias-reduced-estimator} satisfies
\begin{equation}
\label{eq:pool-ext:bias-reduced-clt}
  \frac{\sqrt{k}}{\ell_n}
  \left(
    \frac{
      \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star,\mathrm{bc}}
        (\bm\omega_\gamma,\bm\omega_q)}
         {\expectile{\tau_n^\prime}}
    -1
  \right)
  \;\todist\;
  \mathcal N\!\left(
    0,\;
    \bm\omega_\gamma^\top\Vc\bm\omega_\gamma
  \right).
\end{equation}
The same conclusion holds with estimated weights
$(\hat{\bm\omega}_\gamma,\hat{\bm\omega}_q)$ satisfying the exact
affine constraints
$\hat{\bm\omega}_\gamma^\top\bm 1
=\hat{\bm\omega}_q^\top\bm 1=1$ and
$\hat{\bm\omega}_\gamma\toprob\bm\omega_\gamma$,
$\hat{\bm\omega}_q\toprob\bm\omega_q$.
\end{corollary}

\begin{proof}
See \Cref{app:proofs:pool-ext-bias-reduced-clt}.
\end{proof}

\Cref{cor:pool-ext:bias-reduced-clt} is deliberately a first-order
statement. It should not be read as saying that the intermediate
expectile-bias terms $\Bc^\star$ and $b^{\xi/\Quant}$ disappear from
finite-sample behaviour; they disappear only from the leading
very-extreme normalisation. A higher-order expansion would again have
to keep track of the anchor and of the choice of $\bm\omega_q$.

The practical cost is the same one already present in
\textcite{DaouiaPadoanStupfler2021}: \eqref{eq:pool-ext:Bc-hat-consistency}
requires information about the second-order component of the tail. In
their feasible implementation this is obtained under a Hall-type
working model, $A(t)=\gamma\beta t^\rho$, by estimating the
second-order parameters and substituting them into the formula for
$\Bc$; their supplementary calculations collect standard estimators of
$\rho$ and $\beta$. This extra layer is not a consequence of the
abstract condition $\mathcal C_2(\gamma,\rho,A)$ alone. It is an
additional modelling and estimation step, and it inherits the usual
variance cost and finite-sample instability of second-order bias
estimation noted by \textcite[Remark~4]{DaouiaPadoanStupfler2021}
and, for expectile bias reduction, by
\textcite[\S3.1.1]{DaouiaGirardStupfler2018}.

Once \eqref{eq:pool-ext:bias-reduced-estimator} is used, the
first-order objective is variance rather than AMSE: the limiting mean
in \eqref{eq:pool-ext:bias-reduced-clt} is zero, and the remaining
variance is $\bm\omega_\gamma^\top\Vc\bm\omega_\gamma$. Thus the
natural companion weight is the variance-optimal
$\bm\omega^{\mathrm{Var}}$ of \eqref{eq:pool-ext:var-opt}. Using the
AMSE weight from \eqref{eq:pool-ext:amse-opt} is still asymptotically
valid, but after bias correction it is no longer first-order
variance-optimal unless it coincides with
$\bm\omega^{\mathrm{Var}}$.
No additional first-order variance term appears from $\hatBc$, because
its estimation error enters the normalised log ratio as
$-\bm\omega_\gamma^\top(\hatBc-\Bc)=o_p(1)$.

\section{Asymptotic confidence intervals}
\label{sec:pool-ext:CI}

The limiting theorem above is most naturally used on the logarithmic
scale. This is the same scale used by
\textcite[Corollary~4]{DaouiaPadoanStupfler2021} for pooled
Weissman quantiles. For expectiles,
\textcite[Theorem~3.5 and Eq.~(6)]{DavisonPadoanStupfler2023}
formulate extreme-level intervals on the same log-exponential scale,
and their supplement explicitly motivates the log-scale Gaussian
approximation for expectile intervals
\parencite[\S B.1]{DavisonPadoanStupfler2023}. In the present chapter
the standard error has a particularly simple form: the intermediate
anchor is first-order invisible, and the standard error is the
pooled-Hill standard error multiplied by the extrapolation distance
$\ell_n$. Equivalently, the log-scale variance is scaled by
$\ell_n^2/k$.

For estimated Hill weights define
\begin{equation}
\label{eq:pool-ext:se-def}
  \hat\sigma_{\gamma,n}^{\star\,2}
  \;\coloneqq\;
  \hat{\bm\omega}_{\gamma}^{\top}\hatVc
  \hat{\bm\omega}_{\gamma},
  \qquad
  \ell_n
  =
  \log\!\left(\frac{1-\tau_n}{1-\tau_n^\prime}\right).
\end{equation}
The anchor weight $\hat{\bm\omega}_q$ enters the point estimator but
not the first-order standard error.

\begin{corollary}[Variance-only confidence interval for
                  $\expectile{\tau_n^\prime}$]
\label{cor:pool-ext:CI}
Assume the setting of \Cref{thm:pool-ext:main}. Let
$(\hat{\bm\omega}_{\gamma},\hat{\bm\omega}_{q})$ be estimated weights
satisfying the affine constraints
\[
  \hat{\bm\omega}_{\gamma}^{\top}\bm 1
  =
  \hat{\bm\omega}_{q}^{\top}\bm 1
  =
  1
  \quad\text{exactly,}
  \qquad
  \hat{\bm\omega}_{\gamma}\toprob\bm\omega_{\gamma},
  \qquad
  \hat{\bm\omega}_{q}\toprob\bm\omega_{q},
\]
for deterministic affine limits
$(\bm\omega_{\gamma},\bm\omega_{q})$. Assume also that
\[
  \hatVc \toprob \Vc,
  \qquad
  \bm\omega_{\gamma}^{\top}\Vc\bm\omega_{\gamma}>0,
  \qquad
  \hat\sigma_{\gamma,n}^{\star\,2}>0
  \text{ with probability tending to one.}
\]
Let $\widehat\zeta_{\tau_n^\prime}$ denote either
\[
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
    (\hat{\bm\omega}_{\gamma},\hat{\bm\omega}_{q})
\]
under the centring condition
$\bm\omega_{\gamma}^{\top}\Bc=0$ --- in particular under the
undersmoothing calibration \eqref{eq:pool-int:undersmoothing} --- or
\[
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star,\mathrm{bc}}
    (\hat{\bm\omega}_{\gamma},\hat{\bm\omega}_{q})
\]
under the bias-reduction assumptions of
\Cref{cor:pool-ext:bias-reduced-clt}. Under the positivity convention
of \Cref{sec:pool-ext:estimator} and the expectile--quantile
equivalence, $\widehat\zeta_{\tau_n^\prime}$ is positive with
probability tending to one and $\expectile{\tau_n^\prime}$ is positive
eventually. Then, for every $\alpha\in(0,1)$, with
\begin{equation}
\label{eq:pool-ext:CI-log}
  r_{n,\star}
  \;\coloneqq\;
  z_{1-\alpha/2}\,
  \ell_n
  \sqrt{\frac{\hat\sigma_{\gamma,n}^{\star\,2}}{k}},
  \qquad
  I_{n,\star}^{\log}
  \;\coloneqq\;
  \big[
    \log\widehat\zeta_{\tau_n^\prime}-r_{n,\star},\,
    \log\widehat\zeta_{\tau_n^\prime}+r_{n,\star}
  \big],
\end{equation}
the log-scale interval $I_{n,\star}^{\log}$ and its exponential image
\[
  I_{n,\star}
  \;\coloneqq\;
  \exp(I_{n,\star}^{\log})
  =
  [\exp(L_{n,-}),\exp(L_{n,+})]
  \quad\text{when } I_{n,\star}^{\log}=[L_{n,-},L_{n,+}],
\]
are asymptotic $(1-\alpha)$ confidence intervals for
$\log\expectile{\tau_n^\prime}$ and
$\expectile{\tau_n^\prime}$ respectively:
\[
  \Prob\!\big(\log\expectile{\tau_n^\prime}
        \in I_{n,\star}^{\log}\big)
  \;\to\;
  1-\alpha,
  \qquad
  \Prob\!\big(\expectile{\tau_n^\prime}
        \in I_{n,\star}\big)
  \;\to\;
  1-\alpha.
\]
\end{corollary}

\begin{proof}
See \Cref{app:proofs:pool-ext-CI}.
\end{proof}

\begin{remark}[Bias-centred interval]
\label{rem:pool-ext:CI-bias-centred}
If the asymptotic bias in \Cref{thm:pool-ext:main} is retained rather
than removed, the variance-only interval
\eqref{eq:pool-ext:CI-log} is not centred correctly unless
$\bm\omega_{\gamma}^{\top}\Bc=0$. Suppose instead that
\[
  \hatBc \toprob \Bc,
  \qquad
  \hat\mu_{\gamma,n}^{\star}
  \;\coloneqq\;
  \hat{\bm\omega}_{\gamma}^{\top}\hatBc
  \toprob
  \bm\omega_{\gamma}^{\top}\Bc .
\]
The subscript $\gamma$ emphasises that, on the very-extreme scale, the
surviving first-order bias is the Hill-bias component. With the same
weights and the same $\hatBc$, this centering is algebraically
identical to using the minimal exponent-corrected estimator in
\eqref{eq:pool-ext:bias-reduced-estimator}, since
\[
  \log
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star,\mathrm{bc}}
    (\hat{\bm\omega}_{\gamma},\hat{\bm\omega}_{q})
  =
  \log
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
    (\hat{\bm\omega}_{\gamma},\hat{\bm\omega}_{q})
  -
  \hat\mu_{\gamma,n}^{\star}\frac{\ell_n}{\sqrt{k}}.
\]
Then, for every $\alpha\in(0,1)$, using the radius $r_{n,\star}$ from
\eqref{eq:pool-ext:CI-log}, the log-scale bias-centred interval
\begin{equation}
\label{eq:pool-ext:CI-log-bias-centred}
  C_{n,\star,\mathrm{bcen}}
  \;\coloneqq\;
  \log
  \hatexpectile{\tau_n^\prime}^{\mathrm{pool},\star}
    (\hat{\bm\omega}_{\gamma},\hat{\bm\omega}_{q})
  -
  \hat\mu_{\gamma,n}^{\star}\frac{\ell_n}{\sqrt{k}},
  \qquad
  I_{n,\star,\mathrm{bcen}}^{\log}
  \;\coloneqq\;
  \big[
    C_{n,\star,\mathrm{bcen}}-r_{n,\star},\,
    C_{n,\star,\mathrm{bcen}}+r_{n,\star}
  \big]
\end{equation}
has the same asymptotic coverage as
\eqref{eq:pool-ext:CI-log}, and so does its exponential image. This is
only a feasible procedure when the Hill-bias vector can be estimated
consistently. As in \textcite{DaouiaPadoanStupfler2021}, such a step
typically requires an additional second-order working model, for
example a Hall-type form for $A$. The same practical caveat appears in
the iid expectile discussion of
\textcite[\S3.1.1]{DaouiaGirardStupfler2018}: bias reduction is
available in principle, but it imports the variability and modelling
choices associated with estimating second-order tail parameters.
\end{remark}

The width of \eqref{eq:pool-ext:CI-log} is
\[
  2z_{1-\alpha/2}\,
  \ell_n
  \sqrt{\hat{\bm\omega}_{\gamma}^{\top}\hatVc
        \hat{\bm\omega}_{\gamma}/k}.
\]
Thus the interval expands linearly with the extrapolation distance
from $\tau_n$ to $\tau_n^\prime$ and contracts with the effective
root-$k$ information in the pooled Hill estimator. The standard error
does not involve $m(\gamma)$, $\bm L^\star$, $\Bc^\star$ or
$b^{\xi/\Quant}$; those quantities affect the intermediate anchor but
not the first-order very-extreme normalisation. For the same reason
there is no special restriction at the root $m(\gamma)=0$ in this
section.

In the distributed common-distribution case, the natural
variance-dominated choice is to pair \Cref{cor:pool-ext:CI} with the
variance-optimal effective-sample-size weights of
\eqref{eq:pool-ext:var-opt-distrib}, using the matched convention
$\hat{\bm\omega}_q=\hat{\bm\omega}_\gamma$ for the point estimator.
Under undersmoothing, or more generally whenever
$\bm\omega_\gamma^\top\Bc=0$, this yields the variance-only interval
\eqref{eq:pool-ext:CI-log} without estimating $\Bc$. Outside such a
centred calibration, the same variance-only interval is not
asymptotically centred; valid inference then requires either the
bias-reduced estimator of \eqref{eq:pool-ext:bias-reduced-estimator}
or, equivalently, the bias-centred log interval
\eqref{eq:pool-ext:CI-log-bias-centred}. Both require a consistent
estimator of the Hill-bias vector and therefore belong to the
second-order-calibrated regime.
