% !TEX root = ../main.tex
\chapter{Conclusion}
\label{ch:conclusion}

This work has studied inference for extreme expectile-based risk measures in
a distributed heavy-tailed setting. The construction builds on the
optimal-pooling framework of \textcite{DaouiaPadoanStupfler2021}: it first
forms a geometrically pooled Weissman estimator of an extreme quantile and only
then applies the high-expectile/high-quantile bridge. Looking back, the main
value of the two-weight formulation is that it separates two roles
that are easy to conflate: \(\bm\omega\) controls the pooled-Weissman
component, while \(\bm\nu\) enters only through the Hill estimator used in the
outer bridge. This separation makes clear which part of the estimator is
governed by the source pooling theory.

\section{Main findings}
\label{sec:conclusion:findings}

The theoretical argument is useful precisely because it is explicit about what
has to be controlled. The log error splits into the pooled-Weissman error, the
plug-in error from evaluating \(\psifn\) at a pooled Hill estimator, and the
population error in replacing the expectile by the asymptotic quantile bridge.
In the iid common-marginal distributed regime, under the extreme-Weissman
conditions and the bridge-rate condition, the Chapter~3 asymptotic argument
treats the two bridge terms as lower order on the pooled-Weissman scale. The
resulting conjectural limiting law therefore depends on \(\bm\omega\), but not
on \(\bm\nu\).

This is the main conjectural message of \Cref{ch:pooled-extreme}: first-order
variance and AMSE optimisation transfer to \(\bm\omega\) from the theory of
\textcite{DaouiaPadoanStupfler2021}, while \(\bm\nu\) is first-order
unidentified in this regime. The simulations then adopt the fixed
threshold-allocation bridge vector used in the plug-in route.

The work then formulates a conditional transfer to source-admissible estimated
\(\bm\omega\) weights. This transfer is deliberately conditional: when the
source construction supplies consistent estimated weights and consistent
plug-in centring and variance objects, the expectile estimator is expected to
inherit the same first-order studentised limit. The resulting
confidence interval is obtained by inverting the transferred log-scale
expectile statistic. It is not a direct reuse of a quantile interval formula,
and it does not introduce an expectile-specific plug-in bias or variance
estimator.

\section{Finite-sample lessons}
\label{sec:conclusion:finite-sample}

The simulation study in \Cref{ch:finite-sample} is matched to the theoretical
scope of this work. It uses the practical bridge-Hill convention and varies
only the pooled-Weissman weights. Within this design, the variance weights of
\textcite{DaouiaPadoanStupfler2021} provide a stable practical baseline. In
the representative imbalanced grid, they make the distributed estimator track
the centralised benchmark closely for point estimation, whereas equal
pooled-Weissman weights are less efficient under strong imbalance.

The finite-sample role of AMSE weights is more conditional. In the main
equal-threshold grid, oracle and plug-in AMSE weights are nearly
indistinguishable from variance weights. Under heterogeneous local threshold
fractions, AMSE weights can matter more, and in some rows they improve either
coverage or log RMSE. They are not uniformly superior, however, and the
diagnostics show that signed or unstable weights can appear. Thus the
simulation evidence supports reporting AMSE weights as a meaningful
source-based alternative, not as a uniformly preferable default.

The interval results are more cautionary. The first-order log-scale intervals
under-cover in several finite-sample designs, and the centralised oracle
benchmark also under-covers. The decomposition diagnostics explain why this is
not only a distributed-splitting issue. At the simulated sample sizes, the
terms \(B_n(\bm\nu)\) and \(C_n\), although asymptotically negligible under
the conjectured first-order scaling, can still be visible. In particular,
finite-sample bridge error can affect calibration even when point estimation is
competitive.
The interval study should therefore be read as a diagnostic of the plug-in
route rather than as evidence of strong finite-sample nominal coverage.

\section{Limitations and future work}
\label{sec:conclusion:future}

The limitations of this work are mainly in its scope. The analysis
is carried out in the finite-source iid common-marginal distributed setting.
This restriction is useful: it keeps the target expectile unambiguous and lets
the pooled-Weissman theory of \textcite{DaouiaPadoanStupfler2021} be combined
with the high-expectile/high-quantile bridge without adding a separate
target-selection problem. A natural next step is to move beyond common
margins, for example to tail-homoskedastic or otherwise heterogeneous sources.
That extension would first have to specify the common object being estimated
and then revisit the decomposition terms source by source, because the
population bridge and the centring objects would no longer be identical across
machines.

A second limitation is that the present argument is first-order. Under the
bridge-rate condition used in \Cref{ch:pooled-extreme}, the terms coming from
the plug-in \(\psifn\) factor and from the population expectile--quantile
bridge are lower order on the pooled-Weissman scale. This is precisely what
makes the DPS weight theory transfer to \(\bm\omega\), but it also means that
the bridge-Hill vector \(\bm\nu\) is not selected by the first-order criterion.
A refined analysis retaining lower-order bridge terms would be needed before
any performance-based data-adaptive rule for \(\bm\nu\) could be justified.

The same boundary applies to estimated weights and plug-in inference. The
estimated-weight transfer in this work is an \(\bm\omega\)-transfer: when
the DPS source construction supplies consistent estimated weights and
consistent plug-in centring and variance objects, the pooled extreme expectile
estimator inherits the corresponding first-order log-scale limit. The
bridge-Hill convention remains deterministic in this route. Allowing a random
or fully data-adaptive bridge-Hill vector would require checking that the
compact-event Taylor argument for the \(\psifn\) term still gives a negligible
contribution at the relevant scale, and, if lower-order accuracy is desired,
that the retained bridge terms are correctly represented.

The finite-sample interval results point to a related direction rather than to
a purely numerical gap. The simulations show that first-order log-scale
plug-in intervals can under-cover, including for the centralised benchmark.
This suggests that calibration is affected by threshold choice, second-order
tail bias and finite-sample bridge effects that are deliberately suppressed by
the first-order theory. More systematic calibration work would therefore be
most informative after the relevant lower-order terms or broader source
settings have been theoretically isolated. Possible routes include
bias-corrected centring, alternative studentisation, threshold diagnostics and
simulation-guided sensitivity analysis, but each would need to be tied to a
specific extension of the asymptotic argument.

The simulation study is narrow for the same reason. It is designed to test the
promoted estimator, the DPS variance and AMSE choices for \(\bm\omega\), the
fixed bridge-Hill convention, and the associated plug-in inference route in a
controlled iid common-marginal setting. Broader simulation grids, real-data
examples, time-series dependence, non-identical sources, growing numbers of
machines and more systematic threshold-selection rules are all natural next
analyses. Their interpretation, however, depends on which theoretical
extension is being assessed. Simulating such designs before the corresponding
target and decomposition have been fixed would risk turning numerical
evidence into a substitute for the missing mathematical layer.

Finally, this work follows the geometrically pooled Weissman route because
it is the route directly supported by the DPS framework and by the log-scale
decomposition used in \Cref{ch:pooled-extreme}. Other architectures remain
possible. One could transmit and pool local quantile-based or LAWS expectile
estimators, compare geometric and arithmetic pooling of expectile estimates,
or change the communication protocol to send additional tail summaries. These
are not merely cosmetic variants of the estimator studied here. They would
define different statistical objects, and should therefore be analysed through
their own exact decompositions before being compared theoretically or in
finite samples.
