% !TEX root = ../main.tex
\chapter{Discussion and future work}
\label{ch:discussion}

\section{Summary of contributions}
\label{sec:discussion:summary}

\todo{Restate the contributions of \Cref{ch:pool-int,ch:pool-ext} in
the language of \Cref{sec:intro:contributions}. Emphasise the
modular architecture: the pooled inputs $(\hat\gamma_n,\hatWei_n)$ of
\textcite{DaouiaPadoanStupfler2021} compose cleanly with the
expectile--quantile asymptotic equivalence via the plug-in CLT of
\textcite[Prop.~C.6]{DavisonPadoanStupfler2023}, so the only
expectile-specific computation is the delta-method calculus around
$\psifn(\gamma) = (\gamma^{-1}-1)^{-\gamma}$.}

\section{Connections to Expected Shortfall}
\label{sec:discussion:ES}

\todo{For a heavy-tailed loss with index $\gamma\in(0,1)$ there is an
asymptotic relation $\expectile{\tau}\sim \mathrm{ES}_{\tau'}$ for a
specific matching level $\tau'=\tau'(\tau,\gamma)$ that can be expressed
in closed form (cf.\ \textcite{DaouiaGirardStupfler2018}). Thus pooled
expectile inference yields, as an immediate corollary, pooled inference
on the Expected Shortfall, sidestepping the non-elicitability of ES.
Sketch the level-matching identity and the corresponding plug-in
estimator.}

\section{Limitations}
\label{sec:discussion:limitations}

\todo{(i) Tail-homogeneity across samples is assumed when pooling
extreme quantiles geometrically (cf.\ assumption $(\mathcal H)$ in
\textcite[\S2.3]{DaouiaPadoanStupfler2021}) and is also needed for
pooling expectiles. The test of \Cref{sec:pool-int:test} should be
applied first.\\
(ii) The choice of effective sample sizes $k_j$ is data-driven and
introduces variability that the asymptotic theory does not account for.\\
(iii) The quantile-based route remains theoretically available for
finite-mean tails $\gamma\in[1/2,1)$, but standard direct-LAWS
finite-variance CLTs no longer apply there; simulation evidence in that
range should therefore be interpreted as QB stress testing rather than
LAWS comparison.\\
(iv) The framework is iid within and across samples. Time-series
dependence within a sample and dependence across samples are excluded.}

\section{Future work}
\label{sec:discussion:future-work}

\todo{(i) \emph{Time-series dependence within each sample.} Extending
the framework to $\beta$-mixing observations within a sample, in the
spirit of \textcite{DavisonPadoanStupfler2023}, would require inflating
the within-sample asymptotic variance by a serial-dependence factor.
The pooled inference architecture itself is preserved.\\
(ii) \emph{Multivariate observations.} Each sample could be
$d$-dimensional, with the goal of pooling marginal-component-wise
expectiles. The joint expectile machinery of
\textcite{PadoanStupfler2022} provides the within-sample tools; the
cross-sample step is identical to the univariate case.\\
(iii) \emph{Tail-heteroscedasticity.} A simple parametric model for
$q_j(\tau)/q_\ell(\tau) \to \kappa_{j,\ell}$ would allow pooling even
when assumption $(\mathcal H)$ fails.\\
(iv) \emph{Empirical applications.} Distributed-inference settings in
operational risk (cross-bank loss data), insurance (multi-line claims),
and climate (multi-station extremes) are natural use cases.\\
(v) \emph{Bias-robust pooled estimators.} Use bias-reduced tail-index
estimators (e.g.\ corrected Hill) in place of the plain Hill estimator
in the pooled construction.}
