% Master's thesis: Improved Statistical Inference for Expectile-Based Risk Measures
% Author: Filippo Gombac
% Compile with: latexmk -pdf main.tex

\input{preamble}

\title{Improved Statistical Inference for Expectile-Based Risk Measures\\[0.5em]
       \large An Optimal Pooling Approach in the Heavy-Tailed Regime}
\author{Filippo Gombac\\
        Supervisor: Simone Padoan\\
        Department of Decision Sciences\\
        Bocconi University}
\date{\today}

\begin{document}

\frontmatter
\hypersetup{pageanchor=false}
\maketitle
\cleardoublepage
\hypersetup{pageanchor=true}

\chapter*{Abstract}
This thesis studies statistical inference for extreme expectile-based risk
measures in a distributed heavy-tailed setting. It extends optimal
weighted-pooling methods for extreme quantiles to a pooled extreme-expectile
estimator built from a geometrically pooled Weissman quantile estimator and an
expectile--quantile bridge evaluated with a pooled Hill tail-index estimator.
The estimator is deliberately written with two weight vectors:
\(\bm\omega\) weights the pooled-Weissman component, while a separate
deterministic vector \(\bm\nu\) is used only in the outer bridge-Hill
estimator.

An exact log decomposition separates the pooled-Weissman error, the
bridge-Hill plug-in error and the population expectile--quantile bridge error.
Under the adopted iid common-marginal regime and an explicit bridge-rate
condition, the first-order limit is inherited from the pooled extreme-quantile
component. It depends on \(\bm\omega\), but not on \(\bm\nu\), so first-order
variance and AMSE optimisation apply only to the pooled-Weissman weights.
Estimated \(\bm\omega\) weights and plug-in centring and variance objects
transfer from the source pooling theory when its consistency conditions hold,
yielding a studentised statistic and a multiplicative log-scale interval. The
finite-sample study finds variance weights to be a stable baseline; AMSE
weights are mostly close in the equal-threshold grid but can matter under
heterogeneous thresholds. The plug-in intervals are reported as diagnostics,
since undercoverage remains visible in several designs.

\tableofcontents

\mainmatter

\include{chapters/01_introduction}
\include{chapters/02_background}
\include{chapters/03_pooled_extreme_expectiles}
\include{chapters/04_finite_sample_study}
\include{chapters/05_discussion}

\backmatter
\printbibliography[heading=bibintoc]

\end{document}
